---
title: "Hypoglycemia Analysis - CFRD REMOVED"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    fig_height: 8
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nlme)
library(emmeans)
library(patchwork)
library(ggsignif)
library(arsenal)
library(zoo)
library(vcd)
library(kableExtra)
```

```{r}
getAUCs <- function(data, subject, time, vars) {
  all_return <- NULL
  
  for (var in vars) {
    subs <- unique(data[[subject]])
    return <- NULL
    for (sub in subs) {
      try({
        df <- data[data[[subject]] == sub, ]
        ogtt_0 <- df[df[[time]] == 0, var]
        ogtt_30 <- df[df[[time]] == 30, var]
        ogtt_60 <- df[df[[time]] == 60, var]
        ogtt_90 <- df[df[[time]] == 90, var]
        ogtt_120 <- df[df[[time]] == 120, var]
        ogtt_auc <- (ogtt_30[[var]] + ogtt_60[[var]] + ogtt_90[[var]]) / 
          2 + (ogtt_0[[var]] + ogtt_120[[var]]) / 4
        if (length(ogtt_auc) > 0) {
          row <- c(sub, ogtt_auc)
          return <- rbind(return, row)
        }
      })
    }
    return <- as.data.frame(return)
    names(return) <- c(subject, paste0(var, "_auc"))
    if (is.null(all_return)) {
      all_return <- return
    } else {
      all_return <- merge(all_return, return, all = T)
    }
  }
  all_return
}
```

```{r, warning = F, message = F}
dat <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Hypoglycemia/Background/Previous Analyses/FsOGTT/data/CGMInHealthyControls_DATA_2021-05-25_0918.csv", na = "-999")

dat$group_category[dat$group_category == "1"] <- "Healthy"
dat$group_category[dat$group_category == "2"] <- "CF-NGT"
dat$group_category[dat$group_category == "3"] <- "CF-AG"
dat$group_category[dat$group_category == "4"] <- "CFRD"
dat$group_category <- factor(dat$group_category, levels = c("Healthy", "CF-NGT", "CF-AG", "CFRD"))

dat$minute_60_glucose <- dat$lab_ogtt_1_hour_glucose
dat$minute_120_glucose <- dat$lab_ogtt_2_hour_glucose
dat$minute_0_glucose <- dat$lab_ogtt_fasting

dat$weight[is.na(dat$weight)] <- dat$weight_visit_2[is.na(dat$weight)]
dat$height[is.na(dat$height)] <- dat$height_visit_2[is.na(dat$height)]
dat$bmi[is.na(dat$bmi)] <- dat$bmi_visit_2[is.na(dat$bmi)]
dat$bmi_z_score[is.na(dat$bmi_z_score)] <- dat$bmi_z_score_visit_2[is.na(dat$bmi_z_score)]

dat <- dat[dat$date_of_visit != "", ]

dat$subject_id <- gsub("*--1", "", dat$subject_id)
dat$subject_id <- gsub("*--2", "", dat$subject_id)
dat$subject_id <- gsub("*\\.1", "", dat$subject_id)
dat$subject_id <- gsub("*\\.2", "", dat$subject_id)
dat <- unique(dat)

dat$age <- as.numeric(dat$age)

dat$gender[dat$gender == 1] <- "Male"
dat$gender[dat$gender == 2] <- "Female"

dat$race_ethnicity[dat$race_ethnicity == 1] <- "White"
dat$race_ethnicity[dat$race_ethnicity == 2] <- "Black/African American"
dat$race_ethnicity[dat$race_ethnicity == 3] <- "Asian"
dat$race_ethnicity[dat$race_ethnicity == 5] <- "Hispanic"
dat$race_ethnicity[dat$race_ethnicity == 6] <- "Other/Multiple"

dat$tanner_stage[dat$tanner_stage___1 == 1] <- 1
dat$tanner_stage[dat$tanner_stage_male_testicul___1 == 1] <- 1
dat$tanner_stage[dat$tanner_stage_female_pubic___1 == 1] <- 1
dat$tanner_stage[dat$tanner_stage_female_breast___1 == 1] <- 1

dat$tanner_stage[dat$tanner_stage___2 == 1] <- 2
dat$tanner_stage[dat$tanner_stage_male_testicul___2 == 1] <- 2
dat$tanner_stage[dat$tanner_stage_female_pubic___2 == 1] <- 2
dat$tanner_stage[dat$tanner_stage_female_breast___2 == 1] <- 2

dat$tanner_stage[dat$tanner_stage___3 == 1] <- 3
dat$tanner_stage[dat$tanner_stage_male_testicul___3 == 1] <- 3
dat$tanner_stage[dat$tanner_stage_female_pubic___3 == 1] <- 3
dat$tanner_stage[dat$tanner_stage_female_breast___3 == 1] <- 3

dat$tanner_stage[dat$tanner_stage___4 == 1] <- 4
dat$tanner_stage[dat$tanner_stage_male_testicul___4 == 1] <- 4
dat$tanner_stage[dat$tanner_stage_female_pubic___4 == 1] <- 4
dat$tanner_stage[dat$tanner_stage_female_breast___4 == 1] <- 4

dat$tanner_stage[dat$tanner_stage___5 == 1] <- 5
dat$tanner_stage[dat$tanner_stage_male_testicul___5 == 1] <- 5
dat$tanner_stage[dat$tanner_stage_female_pubic___5 == 1] <- 5
dat$tanner_stage[dat$tanner_stage_female_breast___5 == 1] <- 5

dat$g_tube_feeds[dat$g_tube_feeds == 1] <- "Yes"
dat$g_tube_feeds[dat$g_tube_feeds == 0] <- "No"

gluc <- dat %>% select(
  subject_id,
  date_of_visit,
  age,
  gender,
  race_ethnicity,
  starts_with("min"),
  -starts_with("min_spent"),
  -contains("sensor"),
  -contains("active"),
  -contains("glp"),
  -contains("gip")
)
gluc <- merge(gluc, dat %>% select(subject_id, date_of_visit, age, starts_with("ucd_")))
gluc <- unique(gluc)
gluc <- gluc[!is.na(gluc$min_30_glucose), ]
gluc[] <- lapply(gluc, gsub, pattern = "<", replacement = "")

gluc_names <- names(gluc)
gluc_names <- gsub("min_", "", gluc_names)
gluc_names <- gsub("minute_", "", gluc_names)
gluc_names <- gsub("minus_", "-", gluc_names)
gluc_names <- gsub("ucd_", "", gluc_names)
gluc_names <- gsub("_total", "", gluc_names)
gluc_names <- gsub("glp_1", "_glp1", gluc_names)
gluc_names <- gsub("c_peptide", "cpeptide", gluc_names)
names(gluc) <- gluc_names
gluc[gluc == -999.99] <- NA
gluc <- unique(gluc)
new_gluc <- gluc[0, ]

for (sub in unique(gluc$subject_id)) {
  df <- gluc[gluc$subject_id == sub, ]
  if (nrow(df) > 1) {
    df <- df %>% arrange(`0_gip`, desc(date_of_visit))
    df <- df[1, ]
  }
  new_gluc <- rbind(new_gluc, df)
}


gluc_wide <- new_gluc %>% select(-date_of_visit)
gluc <- gluc_wide %>% pivot_longer(
  cols = -c(subject_id),
  names_to = c("time", ".value"), names_sep = "_"
)

# gluc <- gluc[!is.na(gluc$glucose) | !is.na(gluc$glucagon) | !is.na(gluc$insulin)| !is.na(gluc$cpeptide),]
# gluc <- gluc[!is.na(gluc$glucagon),]
gluc <- gluc[gluc$time != "-10", ]
gluc$time <- as.factor(as.numeric(gluc$time))

gluc$glucose <- as.numeric(gluc$glucose)
gluc$insulin <- as.numeric(gluc$insulin)
gluc$glucagon <- as.numeric(gluc$glucagon)
gluc$cpeptide <- as.numeric(gluc$cpeptide)
gluc$gip <- as.numeric(gluc$gip)
gluc$glp1 <- as.numeric(gluc$glp1)

gluc$gip[gluc$gip < 0] <- NA
gluc$glp1[gluc$glp1 < 0] <- NA

gluc <- gluc %>% filter(!is.na(time))

aucs <- getAUCs(gluc, "subject_id", "time", c("glucose", "glucagon", "insulin", "cpeptide", "gip", "glp1"))

df_aucs <- merge(aucs, unique(gluc %>% select(subject_id)))
df_aucs$glucagon_auc <- as.numeric(df_aucs$glucagon_auc)
df_aucs$glucose_auc <- as.numeric(df_aucs$glucose_auc)
df_aucs$insulin_auc <- as.numeric(df_aucs$insulin_auc)
df_aucs$cpeptide_auc <- as.numeric(df_aucs$cpeptide_auc)
df_aucs$gip_auc <- as.numeric(df_aucs$gip_auc)
df_aucs$glp1_auc <- as.numeric(df_aucs$glp1_auc)

df_aucs <- df_aucs[!is.na(df_aucs$glucagon_auc) | !is.na(df_aucs$glucose_auc) | !is.na(df_aucs$insulin_auc) | !is.na(df_aucs$cpeptide_auc), ]

df_aucs <- unique(df_aucs)

gluc_aucs <- df_aucs

gluc$hypo_60 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])), "Yes", "No")
gluc$hypo_70 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])), "Yes", "No")


gluc_aucs <- merge(gluc_aucs, unique(gluc %>% select(subject_id, hypo_60, hypo_70)))

glucose <- gluc_wide %>% select(subject_id, contains("glucose"), -starts_with("-10"))
glucose <- glucose[complete.cases(glucose), ]
glucose[, 2:10] <- sapply(glucose[, 2:10], as.numeric)
glucose$glucose_AUC <- rowSums(glucose[, 2:10])
glucose$glucose_mean <- glucose$glucose_AUC / 9

insulin <- gluc_wide %>% select(subject_id, contains("insulin"), -starts_with("-10"))
insulin <- insulin[complete.cases(insulin), ]
insulin[, 2:10] <- sapply(insulin[, 2:10], as.numeric)
insulin$insulin_AUC <- rowSums(insulin[, 2:10])
insulin$insulin_mean <- insulin$insulin_AUC / 9

cpeptide <- gluc_wide %>% select(subject_id, contains("cpeptide"), -starts_with("-10"))
cpeptide <- cpeptide[complete.cases(cpeptide), ]
cpeptide[, 2:10] <- sapply(cpeptide[, 2:10], as.numeric)
cpeptide$cpeptide_AUC <- rowSums(cpeptide[, 2:10])
cpeptide$cpeptide_mean <- cpeptide$cpeptide_AUC / 9

glucagon <- gluc_wide %>% select(subject_id, contains("glucagon"), -starts_with("-10"))
glucagon <- glucagon[complete.cases(glucagon), ]
glucagon[, 2:10] <- sapply(glucagon[, 2:10], as.numeric)
glucagon$glucagon_AUC <- rowSums(glucagon[, 2:10])
glucagon$glucagon_mean <- glucagon$glucagon_AUC / 9

gip <- gluc_wide %>% select(subject_id, contains("gip"), -starts_with("minus"))
gip <- gip[complete.cases(gip), ]
gip[, 2:10] <- sapply(gip[, 2:10], as.numeric)
gip$gip_AUC <- rowSums(gip[, 2:10])
gip$gip_mean <- gip$gip_AUC / 9

glp1 <- gluc_wide %>% select(subject_id, contains("glp"), -starts_with("minus"))
glp1 <- glp1[complete.cases(glp1), ]
glp1[, 2:10] <- sapply(glp1[, 2:10], as.numeric)
glp1$glp1_AUC <- rowSums(glp1[, 2:10])
glp1$glp1_mean <- glp1$glp1_AUC / 9

auc_df <- merge(glucose, insulin, all = T)
auc_df <- merge(auc_df, cpeptide, all = T)
auc_df <- merge(auc_df, glucagon, all = T)
auc_df <- merge(auc_df, gip, all = T)
auc_df <- merge(auc_df, glp1, all = T)

auc_df$hypo_60 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])), "Yes", "No")
auc_df$hypo_70 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])), "Yes", "No")

auc_df$matsuda_ins <- 10000 / sqrt(auc_df$`0_glucose` * auc_df$`0_insulin` * auc_df$glucose_mean * auc_df$insulin_mean)
auc_df$matsuda_cpep <- 20 / (auc_df$`0_cpeptide` * auc_df$`0_glucose`)
auc_df$homa_ir <- auc_df$`0_glucose` * auc_df$`0_insulin` / 405
auc_df$igi <- (auc_df$`30_insulin` - auc_df$`0_insulin`) / (auc_df$`30_glucose` - auc_df$`0_glucose`)
auc_df$igi[is.infinite(auc_df$igi)] <- NA
auc_df$cpep_index <- 100 * auc_df$`0_cpeptide` / auc_df$`0_glucose`
auc_df$odi_ins <- auc_df$igi * auc_df$matsuda_ins
auc_df$odi_cpep <- auc_df$igi * auc_df$matsuda_cpep
auc_df$cpep_ins0 <- auc_df$`0_cpeptide` / auc_df$`0_insulin`
auc_df$cpep_auc_30 <- (auc_df$`0_cpeptide` + auc_df$`10_cpeptide` + auc_df$`20_cpeptide` + auc_df$`30_cpeptide`) / 4
auc_df$ins_auc_30 <- (auc_df$`0_insulin` + auc_df$`10_insulin` + auc_df$`20_insulin` + auc_df$`30_insulin`) / 4
auc_df$cpep_ins_auc_30 <- auc_df$cpep_auc_30 / auc_df$ins_auc_30
auc_df$cpep_ins_auc <- auc_df$cpeptide_AUC / auc_df$insulin_AUC

gluc_peak <- NULL
gluc_nadir <- NULL
gluc_peak_time <- NULL
gluc_nadir_time <- NULL

ins_peak <- NULL
ins_nadir <- NULL
ins_peak_time <- NULL
ins_nadir_time <- NULL

cpep_peak <- NULL
cpep_nadir <- NULL
cpep_peak_time <- NULL
cpep_nadir_time <- NULL

gluca_peak <- NULL
gluca_nadir <- NULL
gluca_peak_time <- NULL
gluca_nadir_time <- NULL

gip_peak <- NULL
gip_nadir <- NULL
gip_peak_time <- NULL
gip_nadir_time <- NULL

glp1_peak <- NULL
glp1_nadir <- NULL
glp1_peak_time <- NULL
glp1_nadir_time <- NULL

for (subject in auc_df$subject_id) {
  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("glucose"))
  gluc_peak <- c(gluc_peak, max(sub_df))
  gluc_nadir <- c(gluc_nadir, min(sub_df))
  gluc_peak_time <- c(
    gluc_peak_time,
    strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1]
  )
  gluc_nadir_time <- c(
    gluc_nadir_time,
    strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1]
  )
  
  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("insulin"))
  ins_peak <- c(ins_peak, max(sub_df))
  ins_nadir <- c(ins_nadir, min(sub_df))
  ins_peak_time <- c(
    ins_peak_time,
    strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1]
  )
  ins_nadir_time <- c(
    ins_nadir_time,
    strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1]
  )
  
  
  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("cpeptide"))
  cpep_peak <- c(cpep_peak, max(sub_df))
  cpep_nadir <- c(cpep_nadir, min(sub_df))
  
  cppt_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  cpnt_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(cppt_temp) > 0) {
    cpep_peak_time <- c(
      cpep_peak_time,
      cppt_temp[[1]][1]
    )
  } else {
    cpep_peak_time <- c(
      cpep_peak_time,
      NA
    )
  }
  if (length(cpnt_temp) > 0) {
    cpep_nadir_time <- c(
      cpep_nadir_time,
      cpnt_temp[[1]][1]
    )
  } else {
    cpep_nadir_time <- c(
      cpep_nadir_time,
      NA
    )
  }
  
  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("glucagon"))
  gluca_peak <- c(gluca_peak, max(sub_df))
  gluca_nadir <- c(gluca_nadir, min(sub_df))
  glucap_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glucan_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(glucap_temp) > 0) {
    gluca_peak_time <- c(
      gluca_peak_time,
      glucap_temp[[1]][1]
    )
  } else {
    gluca_peak_time <- c(
      gluca_peak_time,
      NA
    )
  }
  if (length(glucan_temp) > 0) {
    gluca_nadir_time <- c(
      gluca_nadir_time,
      glucan_temp[[1]][1]
    )
  } else {
    gluca_nadir_time <- c(
      gluca_nadir_time,
      NA
    )
  }
  
  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("gip"))
  gip_peak <- c(gip_peak, max(sub_df))
  gip_nadir <- c(gip_nadir, min(sub_df))
  gipp_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  gipn_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(gipp_temp) > 0) {
    gip_peak_time <- c(
      gip_peak_time,
      gipp_temp[[1]][1]
    )
  } else {
    gip_peak_time <- c(
      gip_peak_time,
      NA
    )
  }
  if (length(gipn_temp) > 0) {
    gip_nadir_time <- c(
      gip_nadir_time,
      gipn_temp[[1]][1]
    )
  } else {
    gip_nadir_time <- c(
      gip_nadir_time,
      NA
    )
  }
  
  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("glp1"))
  glp1_peak <- c(glp1_peak, max(sub_df))
  glp1_nadir <- c(glp1_nadir, min(sub_df))
  glp1p_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glp1n_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(gipp_temp) > 0) {
    glp1_peak_time <- c(
      glp1_peak_time,
      glp1p_temp[[1]][1]
    )
  } else {
    glp1_peak_time <- c(
      glp1_peak_time,
      NA
    )
  }
  if (length(glp1n_temp) > 0) {
    glp1_nadir_time <- c(
      glp1_nadir_time,
      glp1n_temp[[1]][1]
    )
  } else {
    glp1_nadir_time <- c(
      glp1_nadir_time,
      NA
    )
  }
}

auc_df$gluc_peak <- gluc_peak
auc_df$gluc_nadir <- gluc_nadir
auc_df$gluc_peak_time <- gluc_peak_time
auc_df$gluc_nadir_time <- gluc_nadir_time
auc_df$ins_peak <- ins_peak
auc_df$ins_nadir <- ins_nadir
auc_df$ins_peak_time <- ins_peak_time
auc_df$ins_nadir_time <- ins_nadir_time
auc_df$cpep_peak <- cpep_peak
auc_df$cpep_nadir <- cpep_nadir
auc_df$cpep_peak_time <- cpep_peak_time
auc_df$cpep_nadir_time <- cpep_nadir_time
auc_df$gluca_peak <- gluca_peak
auc_df$gluca_nadir <- gluca_nadir
auc_df$gluca_peak_time <- gluca_peak_time
auc_df$gluca_nadir_time <- gluca_nadir_time

auc_df$gip_peak <- gip_peak
auc_df$gip_nadir <- gip_nadir
auc_df$gip_peak_time <- gip_peak_time
auc_df$gip_nadir_time <- gip_nadir_time

auc_df$glp1_peak <- glp1_peak
auc_df$glp1_nadir <- glp1_nadir
auc_df$glp1_peak_time <- glp1_peak_time
auc_df$glp1_nadir_time <- glp1_nadir_time

keep_ids <- gluc_aucs$subject_id


table_dat <- merge(new_gluc, gluc_aucs)
table_dat <- merge(table_dat, auc_df %>% select(
  subject_id,
  matsuda_ins,
  matsuda_cpep,
  homa_ir,
  igi,
  cpep_index,
  odi_ins,
  odi_cpep,
  cpep_ins0,
  cpep_ins_auc_30,
  cpep_ins_auc,
  `0_glucose`,
  `0_insulin`,
  `0_cpeptide`,
  `0_glucagon`,
  `0_gip`,
  `0_glp1`,
  gluc_peak,
  gluc_nadir,
  gluc_peak_time,
  gluc_nadir_time,
  ins_peak,
  ins_nadir,
  ins_peak_time,
  ins_nadir_time,
  cpep_peak,
  cpep_nadir,
  cpep_peak_time,
  cpep_nadir_time,
  gluca_peak,
  gluca_nadir,
  gluca_peak_time,
  gluca_nadir_time,
  gip_peak,
  gip_nadir,
  gip_peak_time,
  gip_nadir_time,
  glp1_peak,
  glp1_nadir,
  glp1_peak_time,
  glp1_nadir_time
), all = T)

cgm <- dat %>%
  select(
    subject_id,
    date_of_visit,
    num_days_good_data:interquartile_range
  ) %>%
  filter(subject_id %in% keep_ids) %>%
  filter(!is.na(num_days_good_data)) %>%
  unique()

cgm$daytime_avg_sensor_glucose <- round(as.numeric(cgm$daytime_avg_sensor_glucose), 4)
cgm$percent_time_over_140 <- round(as.numeric(cgm$percent_time_over_140), 4)
cgm$percent_time_over_180 <- round(as.numeric(cgm$percent_time_over_180), 4)
cgm$daytime_sd <- round(as.numeric(cgm$daytime_sd), 4)
cgm$cov <- round(as.numeric(cgm$cov), 4)
cgm$mag <- round(as.numeric(cgm$mag), 4)
cgm <- unique(cgm)
cgm <- cgm[!duplicated(cgm[, 1:2]), ]
cgm <- cgm[!duplicated(cgm$subject_id), ]

table_dat <- merge(table_dat, cgm %>% select(-date_of_visit), all = T)


dat$mutation_severity[dat$cf_mutation_1_class <= 3 | dat$cf_mutation_2_class <= 3] <- "Severe"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 4 |
                                                        dat$cf_mutation_2_class == 4 |
                                                        dat$cf_mutation_1_class <= 5 |
                                                        dat$cf_mutation_2_class <= 5)] <- "Mild"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 6 |
                                                        dat$cf_mutation_2_class == 6)] <- "Other"



dat_demo <- dat %>% select(
  subject_id,
  date_of_visit,
  group_category,
  gender,
  race_ethnicity,
  tanner_stage,
  weight,
  height,
  bmi,
  bmi_z_score,
  mutation_severity,
  fev1_percent,
  fvc,
  poc_a1c,
  cf_pancreatic,
  cf_insulin_use,
  g_tube_feeds
)
dat_demo <- dat_demo %>% filter(subject_id %in% keep_ids)

new_demo <- dat_demo[0, ]
for (id in unique(dat_demo$subject_id)) {
  id_df <- dat_demo %>%
    filter(subject_id == id) %>%
    arrange(date_of_visit) %>%
    na.locf(na.rm = FALSE)
  id_df <- id_df %>%
    arrange(desc(date_of_visit)) %>%
    na.locf(na.rm = FALSE)
  new_demo <- rbind(new_demo, id_df)
}

dat_demo <- new_demo %>%
  select(-date_of_visit) %>%
  unique()

dat_demo <- dat_demo[!duplicated(dat_demo$subject_id), ]

table_dat <- merge(table_dat, dat_demo, all = T)


table_dat$`0_glucose` <- as.numeric(table_dat$`0_glucose`)
table_dat$`120_glucose` <- as.numeric(table_dat$`120_glucose`)
table_dat$`60_glucose` <- as.numeric(table_dat$`60_glucose`)

table_dat$`0_insulin` <- as.numeric(table_dat$`0_insulin`)
table_dat$`120_insulin` <- as.numeric(table_dat$`120_insulin`)
table_dat$`60_insulin` <- as.numeric(table_dat$`60_insulin`)

table_dat$`0_cpeptide` <- as.numeric(table_dat$`0_cpeptide`)
table_dat$`120_cpeptide` <- as.numeric(table_dat$`120_cpeptide`)
table_dat$`60_cpeptide` <- as.numeric(table_dat$`60_cpeptide`)

table_dat$`0_glucagon` <- as.numeric(table_dat$`0_glucagon`)
table_dat$`120_glucagon` <- as.numeric(table_dat$`120_glucagon`)
table_dat$`60_glucagon` <- as.numeric(table_dat$`60_glucagon`)

table_dat$`0_gip` <- as.numeric(table_dat$`0_gip`)
table_dat$`120_gip` <- as.numeric(table_dat$`120_gip`)
table_dat$`60_gip` <- as.numeric(table_dat$`60_gip`)

table_dat$`0_glp1` <- as.numeric(table_dat$`0_glp1`)
table_dat$`120_glp1` <- as.numeric(table_dat$`120_glp1`)
table_dat$`60_glp1` <- as.numeric(table_dat$`60_glp1`)

table_dat$ogtt_category[table_dat$`0_glucose` < 100 & table_dat$`60_glucose` < 155 & table_dat$`120_glucose` < 140] <- "NGT"
table_dat$ogtt_category[table_dat$`0_glucose` < 100 & table_dat$`60_glucose` > 155 &
                          table_dat$`120_glucose` < 140 & table_dat$`60_glucose` < 200] <- "AGT" # previously EGI
table_dat$ogtt_category[(table_dat$`0_glucose` >= 100 & table_dat$`0_glucose` < 125) |
                          (table_dat$`60_glucose` > 199) |
                          (table_dat$`120_glucose` > 140 & table_dat$`120_glucose` <= 200)] <- "AGT" # previously IFG/IGT/INDET
table_dat$ogtt_category[table_dat$`0_glucose` >= 125 | table_dat$`120_glucose` >= 200] <- "CFRD"

cf_groups <- dat %>%
  select(subject_id, group_category) %>%
  filter(subject_id %in% keep_ids) %>%
  na.omit() %>%
  unique()
cf_groups$CF <- ifelse(cf_groups$group_category == "Healthy", "No", "Yes")
cf_groups <- cf_groups %>%
  select(subject_id, CF) %>%
  unique()
table_dat <- merge(table_dat, cf_groups)

table_dat$CF_hypo_60[table_dat$CF == "Yes" & table_dat$hypo_60 == "Yes"] <- "CF/Hypo+"
table_dat$CF_hypo_60[table_dat$CF == "Yes" & table_dat$hypo_60 == "No"] <- "CF/Hypo-"
table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "Yes"] <- "HC/Hypo+"
table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "No"] <- "HC/Hypo-"

table_dat$CF_hypo_70[table_dat$CF == "Yes" & table_dat$hypo_70 == "Yes"] <- "CF/Hypo+"
table_dat$CF_hypo_70[table_dat$CF == "Yes" & table_dat$hypo_70 == "No"] <- "CF/Hypo-"
table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "Yes"] <- "HC/Hypo+"
table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "No"] <- "HC/Hypo-"

table_dat <- table_dat %>% filter(is.na(cf_pancreatic) | cf_pancreatic == 2)
table_dat$cf_insulin_use[table_dat$cf_insulin_use == 1] <- "Yes"
table_dat$cf_insulin_use[table_dat$cf_insulin_use == 0] <- "No"

table_dat$delta_glucagon <- as.numeric(table_dat$`180_glucagon`) - as.numeric(table_dat$`150_glucagon`)
table_dat$delta_glucose <- as.numeric(table_dat$`180_glucose`) - as.numeric(table_dat$`150_glucose`)
table_dat$delta_insulin <- as.numeric(table_dat$`180_insulin`) - as.numeric(table_dat$`150_insulin`)
table_dat$delta_cpeptide <- as.numeric(table_dat$`180_cpeptide`) - as.numeric(table_dat$`150_cpeptide`)
table_dat$delta_gip <- as.numeric(table_dat$`180_gip`) - as.numeric(table_dat$`150_gip`)
table_dat$delta_glp1 <- as.numeric(table_dat$`180_glp1`) - as.numeric(table_dat$`150_glp1`)

gluc <- merge(gluc, table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))

auc_df <- merge(auc_df, table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))

table_dat$age <- as.numeric(table_dat$age)

table_dat <- table_dat[!is.na(table_dat$date_of_visit), ]

write.csv(table_dat, "analysis_data_incl_cfinsulinuse.csv")

# Remove CFRD participants (for reviewer comments)
table_dat = table_dat %>% filter(group_category != "CFRD")
```

# Methods

Global tests by group are Kruskal-Wallis (numeric) or Pearson Chi-square tests (count/categorical)

Pairwise-tests are exact wilcoxon rank sum tests with continuity correction (numeric) or chi-square (count/categorical)

p-values have not been adjusted for multiple comparisons

## After removing CF Insulin Use

Results did not change too much after removing participants with CF Insulin use. Some p-values that were close to 0.05 shifted from one side to the other

After removing CF Insulin Use:

- matsuda now significant on ANOVA with Hypo < 60
- OGTT Category pairwise comparison CF/Hypo60 vs CF/No-Hypo60 now significant
- HOMA-IR pairwise comparison CF/Hypo60 vs HC/No-Hypo60 now significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs HC/Hypo60 no longer significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs HC/No-Hypo60 no longer significant
- CGM MAG CF/No-Hypo60 vs HC/Hypo60 now significant
- Matsuda and Homa-IR pairwise comparisons CF/Hypo70 vs HC/No-Hypo70 now significant
- Delta Glucagon pairwise comparison CF/No-Hypo70 vs HC/Hypo70 no longer significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs HC/Hypo70 no longer significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs HC/No-Hypo70 no longer significant
- %time over 140 pairwise comparison CF/Hypo70 vs CF/Hypo- no longer significant


# With CF Insulin

## Table 1 by CF/Hypo+ < 60

```{r, results='asis', warning = F, message = F}
control <- tableby.control(numeric.test = "kwt")

tab <- tableby(
  CF_hypo_60 ~
    gender +
    age +
    race_ethnicity +
    tanner_stage +
    weight +
    height +
    bmi +
    bmi_z_score +
    mutation_severity +
    fev1_percent +
    fvc +
    poc_a1c +
    ogtt_category +
    cf_insulin_use +
    g_tube_feeds +
    matsuda_ins +
    matsuda_cpep +
    homa_ir +
    igi +
    cpep_index +
    odi_ins +
    odi_cpep +
    cpep_ins0 +
    cpep_ins_auc_30 +
    cpep_ins_auc +
    `0_glucose` +
    gluc_peak +
    gluc_nadir +
    gluc_peak_time +
    gluc_nadir_time +
    glucose_auc +
    delta_glucose +
    `0_insulin` +
    ins_peak +
    ins_nadir +
    ins_peak_time +
    ins_nadir_time +
    insulin_auc +
    delta_insulin +
    `0_cpeptide` +
    cpep_peak +
    cpep_nadir +
    cpep_peak_time +
    cpep_nadir_time +
    cpeptide_auc +
    delta_cpeptide +
    `0_glucagon` +
    gluca_peak +
    gluca_nadir +
    gluca_peak_time +
    gluca_nadir_time +
    glucagon_auc +
    delta_glucagon +
    `0_gip` +
    gip_peak +
    gip_nadir +
    gip_peak_time +
    gip_nadir_time +
    gip_auc +
    delta_gip +
    `0_glp1` +
    glp1_peak +
    glp1_nadir +
    glp1_peak_time +
    glp1_nadir_time +
    glp1_auc +
    delta_glp1,
  data = table_dat,
  control = control
)
summary(tab)
```

```{r, warning = F, message = F}
pairwiseTableby <- function(tab) {
  tbcall <- tab$Call
  data <- eval(tbcall$data)
  group_term <- tab$tables[[1]]$y$term
  group_levels <- head(names(tab$tables[[1]]$y$stats), -1)
  comps <- combn(group_levels, 2)
  
  for (i in seq_len(ncol(comps))) {
    comp_name <- paste(comps[, i], collapse = " vs ")
    sub_df <- data %>% filter(get(group_term) %in% comps[, i])
    try(
      {
        sub_tab <- tableby(eval(tbcall$formula), sub_df, numeric.test = "wt", 
                           wilcox.correct = TRUE, wilcox.exact = TRUE)
        p_list <- 
          lapply(sub_tab$tables[[1]]$tables[[1]], function(x) x$test$p.value)
        if (exists("results_frame")) {
          results_frame[comp_name] <- unlist(p_list)
        } else {
          results_frame <- data.frame(row.names = names(unlist(p_list)))
          results_frame[comp_name] <- unlist(p_list)
        }
      },
      silent = TRUE
    )
  }
  results_frame
}
ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = 
                ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = 
                na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, 
                               "red", "white"), "white)")) %>%
  column_spec(4, background = 
                na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, 
                               "red", "white"), "white)")) %>%
  column_spec(5, background = 
                na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, 
                               "red", "white"), "white)")) %>%
  column_spec(6, background = 
                na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, 
                               "red", "white"), "white)"))
```

## CGM with Hypo < 60

```{r, results='asis', warning = F, message = F}
cgm_table_dat <- table_dat %>% filter(!is.na(daytime_avg_sensor_glucose))

tab <- tableby(
  formula(
    paste(
      "CF_hypo_60 ~",
      paste(
        names(cgm_table_dat %>%
                select(num_days_good_data:interquartile_range)),
        collapse = "+"
      )
    )
  ),
  data = cgm_table_dat,
  control = control
)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = 
                ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = 
                na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, 
                               "red", "white"), "white)")) %>%
  column_spec(4, background = 
                na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, 
                               "red", "white"), "white)")) %>%
  column_spec(5, background = 
                na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, 
                               "red", "white"), "white)")) %>%
  column_spec(6, background = 
                na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, 
                               "red", "white"), "white)"))
```

## OGTT with Hypo < 60

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_60.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
             correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_60_bw.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```


## Insulin Resistance with Hypo < 60

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(45, 50),
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-")),
    color = "black",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_60.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(45, 50),
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-")),
    color = "black",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_60_bw.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

## Table 1 by CF/Hypo+ < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(
  CF_hypo_70 ~
    gender +
    age +
    race_ethnicity +
    tanner_stage +
    weight +
    height +
    bmi +
    bmi_z_score +
    mutation_severity +
    fev1_percent +
    fvc +
    poc_a1c +
    ogtt_category +
    cf_insulin_use +
    g_tube_feeds +
    matsuda_ins +
    matsuda_cpep +
    homa_ir +
    igi +
    cpep_index +
    odi_ins +
    odi_cpep +
    cpep_ins0 +
    cpep_ins_auc_30 +
    cpep_ins_auc +
    `0_glucose` +
    gluc_peak +
    gluc_nadir +
    gluc_peak_time +
    gluc_nadir_time +
    glucose_auc +
    delta_glucose +
    `0_insulin` +
    ins_peak +
    ins_nadir +
    ins_peak_time +
    ins_nadir_time +
    insulin_auc +
    delta_insulin +
    `0_cpeptide` +
    cpep_peak +
    cpep_nadir +
    cpep_peak_time +
    cpep_nadir_time +
    cpeptide_auc +
    delta_cpeptide +
    `0_glucagon` +
    gluca_peak +
    gluca_nadir +
    gluca_peak_time +
    gluca_nadir_time +
    glucagon_auc +
    delta_glucagon +
    `0_gip` +
    gip_peak +
    gip_nadir +
    gip_peak_time +
    gip_nadir_time +
    gip_auc +
    delta_gip +
    `0_glp1` +
    glp1_peak +
    glp1_nadir +
    glp1_peak_time +
    glp1_nadir_time +
    glp1_auc +
    delta_glp1,
  data = table_dat,
  control = control
)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, "red", "white"), "white)")) %>%
  column_spec(5, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(6, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, "red", "white"), "white)"))
```

## CGM with Hypo < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(
  formula(
    paste(
      "CF_hypo_70 ~",
      paste(
        names(cgm_table_dat %>%
                select(num_days_good_data:interquartile_range)),
        collapse = "+"
      )
    )
  ),
  data = cgm_table_dat,
  control = control
)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, "red", "white"), "white)")) %>%
  column_spec(5, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(6, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, "red", "white"), "white)"))
```


## OGTT with Hypo < 70

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glucose ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(insulin ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(cpeptide ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(gip ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glp1 ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_70.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_70_bw.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```

## Insulin Resistance with Hypo < 70

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-")),
    color = "black",
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-")),
    color = "black",
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_70.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-")),
    color = "black",
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-")),
    color = "black",
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_70_bw.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

## CGM Figures

```{r fig.height=5, results="asis"}
outcomes <- c(
  "daytime_avg_sensor_glucose",
  "percent_time_under_60",
  "percent_time_under_70",
  "percent_time_over_140",
  "percent_time_over_180",
  "daytime_sd",
  "cov",
  "mag"
)
outcomes_labs <- c(
  "Avg Glucose (Daytime)",
  "% Time Under 60",
  "% Time Under 70",
  "% Time Over 140",
  "% Time Over 180",
  "SD",
  "COV",
  "MAG"
)

for (i in seq_len(length(outcomes))) {
  p70 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_70")) +
    labs(y = outcomes_labs[i], x = "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 70)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  
  
  p60 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_60")) +
    labs(y = outcomes_labs[i], x = "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 60)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  cat("\n")
  cat(paste("##", outcomes_labs[i]))
  cat("\n")
  print(p70 + p60)
  cat("\n")
}
```

## CGM Hypoglycemia

```{r,results="asis"}
tab <- tableby(group_category ~ percent_time_under_70 + percent_time_under_60, data = cgm_table_dat)
summary(tab)

under_70_cutoff <- 2 * 2.06
under_60_cutoff <- 2 * 0.687

cgm_table_dat$CGM_hypo_60 <- ifelse(cgm_table_dat$percent_time_under_60 > under_60_cutoff, "Yes", "No")
cgm_table_dat$CGM_hypo_70 <- ifelse(cgm_table_dat$percent_time_under_70 > under_70_cutoff, "Yes", "No")
```

Hypoglycemia by CGM is defined as a % time under 70 or 60 more than 2 SDs away from average among healthy controls 

<60 cutoff

```{r}
under_60_cutoff
```

<70 cutoff

```{r}
under_70_cutoff
```

## Hypoglycemia by CGM by group

```{r, results="asis"}
cgm_tableby <- cgm_table_dat
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "No"] <- "Not Hypo by OGTT"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "No"] <- "Not Hypo by OGTT"

tab <- tableby(group_category ~ CGM_hypo_60 + CGM_hypo_70, data = cgm_tableby)
summary(tab)
```

## Hypoglycemia Concordance

### 70/70

Hypoglycemia by OGTT with cutoff <70 and by CGM with cutoff < 70

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of 
-0.15 among participants with CF and 0.07 among those without CF


```{r, results = "asis"}
tab <- tableby(CGM_hypo_70 ~ hypo_70,
               strata = CF,
               data = cgm_tableby,
               test = F,
               total = F
)
summary(tab)
```
#### Concordance with CF
```{r}
mat <- matrix(data = c(4, 12, 16, 23), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix(data = c(1, 2, 3, 9), nrow = 2, ncol = 2)
Kappa(mat)
```

### 60/60

Hypoglycemia by OGTT with cutoff <60 and by CGM with cutoff < 60

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of
-0.17 among participants with CF. The concordance among those without CF was 
0.42, though due to a small sample size, this was not statistically significant. 


```{r, results = "asis"}
tab <- tableby(CGM_hypo_60 ~ hypo_60,
               strata = CF,
               data = cgm_tableby,
               test = F,
               total = F
)
summary(tab)
```
#### Concordance with CF
```{r}
mat <- matrix(data = c(2, 12, 13, 28), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix(data = c(1, 1, 1, 12), nrow = 2, ncol = 2)
Kappa(mat)
```

### 70/60

Hypoglycemia by OGTT with cutoff <70 and by CGM with cutoff < 60

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of -0.17


```{r, results = "asis"}
tab <- tableby(CGM_hypo_60 ~ hypo_70, data = cgm_table_dat, test = F, total = F)
summary(tab)
```
```{r}
mat <- matrix(data = c(33, 21, 13, 3), nrow = 2, ncol = 2)
Kappa(mat)
```

### 60/70

Hypoglycemia by OGTT with cutoff <60 and by CGM with cutoff < 70

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of 0.03


```{r, results = "asis"}
tab <- tableby(CGM_hypo_70 ~ hypo_60, data = cgm_table_dat, test = F, total = F)
summary(tab)
```
```{r}
mat <- matrix(data = c(39, 12, 14, 5), nrow = 2, ncol = 2)
Kappa(mat)
```


```{r}
table_dat <- table_dat %>% filter(cf_insulin_use == "No" | is.na(cf_insulin_use))
gluc <- gluc %>% filter(subject_id %in% table_dat$subject_id)

write.csv(table_dat, "analysis_data_no_cfinsulinuse.csv")
```

# Removed CF Insulin Use

## Table 1 by CF/Hypo+ < 60

```{r, results='asis', warning = F, message = F}
control <- tableby.control(numeric.test = "kwt")

tab <- tableby(
  CF_hypo_60 ~
    gender +
    age +
    race_ethnicity +
    tanner_stage +
    weight +
    height +
    bmi +
    bmi_z_score +
    mutation_severity +
    fev1_percent +
    fvc +
    poc_a1c +
    ogtt_category +
    cf_insulin_use +
    g_tube_feeds +
    matsuda_ins +
    matsuda_cpep +
    homa_ir +
    igi +
    cpep_index +
    odi_ins +
    odi_cpep +
    cpep_ins0 +
    cpep_ins_auc_30 +
    cpep_ins_auc +
    `0_glucose` +
    gluc_peak +
    gluc_nadir +
    gluc_peak_time +
    gluc_nadir_time +
    glucose_auc +
    delta_glucose +
    `0_insulin` +
    ins_peak +
    ins_nadir +
    ins_peak_time +
    ins_nadir_time +
    insulin_auc +
    delta_insulin +
    `0_cpeptide` +
    cpep_peak +
    cpep_nadir +
    cpep_peak_time +
    cpep_nadir_time +
    cpeptide_auc +
    delta_cpeptide +
    `0_glucagon` +
    gluca_peak +
    gluca_nadir +
    gluca_peak_time +
    gluca_nadir_time +
    glucagon_auc +
    delta_glucagon +
    `0_gip` +
    gip_peak +
    gip_nadir +
    gip_peak_time +
    gip_nadir_time +
    gip_auc +
    delta_gip +
    `0_glp1` +
    glp1_peak +
    glp1_nadir +
    glp1_peak_time +
    glp1_nadir_time +
    glp1_auc +
    delta_glp1,
  data = table_dat,
  control = control
)
summary(tab)
ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, "red", "white"), "white)")) %>%
  column_spec(5, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(6, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, "red", "white"), "white)"))
```


## CGM with Hypo < 60

```{r, results='asis', warning = F, message = F}
cgm_table_dat <- table_dat %>% filter(!is.na(daytime_avg_sensor_glucose))

tab <- tableby(
  formula(
    paste(
      "CF_hypo_60 ~",
      paste(
        names(cgm_table_dat %>%
                select(num_days_good_data:interquartile_range)),
        collapse = "+"
      )
    )
  ),
  data = cgm_table_dat,
  control = control
)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, "red", "white"), "white)")) %>%
  column_spec(5, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(6, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, "red", "white"), "white)"))
```

## OGTT with Hypo < 60

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_60_noinsulin.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_60_noinsulin_bw.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```

## Insulin Resistance with Hypo < 60

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(45, 50),
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(4.5, 5),
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_60_noinsulin.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(45, 50),
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(4.5, 5),
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_60_noinsulin_bw.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

## Table 1 by CF/Hypo+ < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(
  CF_hypo_70 ~
    gender +
    age +
    race_ethnicity +
    tanner_stage +
    weight +
    height +
    bmi +
    bmi_z_score +
    mutation_severity +
    fev1_percent +
    fvc +
    poc_a1c +
    ogtt_category +
    cf_insulin_use +
    g_tube_feeds +
    matsuda_ins +
    matsuda_cpep +
    homa_ir +
    igi +
    cpep_index +
    odi_ins +
    odi_cpep +
    cpep_ins0 +
    cpep_ins_auc_30 +
    cpep_ins_auc +
    `0_glucose` +
    gluc_peak +
    gluc_nadir +
    gluc_peak_time +
    gluc_nadir_time +
    glucose_auc +
    delta_glucose +
    `0_insulin` +
    ins_peak +
    ins_nadir +
    ins_peak_time +
    ins_nadir_time +
    insulin_auc +
    delta_insulin +
    `0_cpeptide` +
    cpep_peak +
    cpep_nadir +
    cpep_peak_time +
    cpep_nadir_time +
    cpeptide_auc +
    delta_cpeptide +
    `0_glucagon` +
    gluca_peak +
    gluca_nadir +
    gluca_peak_time +
    gluca_nadir_time +
    glucagon_auc +
    delta_glucagon +
    `0_gip` +
    gip_peak +
    gip_nadir +
    gip_peak_time +
    gip_nadir_time +
    gip_auc +
    delta_gip +
    `0_glp1` +
    glp1_peak +
    glp1_nadir +
    glp1_peak_time +
    glp1_nadir_time +
    glp1_auc +
    delta_glp1,
  data = table_dat,
  control = control
)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, "red", "white"), "white)")) %>%
  column_spec(5, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(6, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, "red", "white"), "white)"))
```

## CGM with Hypo < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(
  formula(
    paste(
      "CF_hypo_70 ~",
      paste(
        names(cgm_table_dat %>%
                select(num_days_good_data:interquartile_range)),
        collapse = "+"
      )
    )
  ),
  data = cgm_table_dat,
  control = control
)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>%
  column_spec(2, background = ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, "red", "white"), "white)")) %>%
  column_spec(5, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
  column_spec(6, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, "red", "white"), "white)"))
```


## OGTT with Hypo < 70

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glucose ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(insulin ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(cpeptide ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(gip ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

model <- lme(glp1 ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette = "Dark2")

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_70_noinsulin.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")

png("ogtt_70_noinsulin_bw.png", res = 300, width = 2000, height = 3000)
p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position = "bottom")
dev.off()
```

## Insulin Resistance with Hypo < 70

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(45, 50),
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(4.5, 5),
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_brewer(palette = "Dark2")


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_70_noinsulin.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(45, 50),
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(
    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
    color = "black",
    y_position = c(4.5, 5),
    test = "wilcox.test",
    map_signif_level = function(p) sprintf("p = %.2g", p)
  ) +
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")

png("resistance_70_noinsulin_bw.png", height = 2000, width = 3000, res = 300)
p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position = "bottom")
dev.off()
```

## CGM Figures

```{r fig.height=5, results="asis"}
outcomes <- c(
  "daytime_avg_sensor_glucose",
  "percent_time_under_60",
  "percent_time_under_70",
  "percent_time_over_140",
  "percent_time_over_180",
  "daytime_sd",
  "cov",
  "mag"
)
outcomes_labs <- c(
  "Avg Glucose (Daytime) [mg/dL]",
  "% Time Under 60 (mg/dL)",
  "% Time Under 70 (mg/dL)",
  "% Time Over 140 (mg/dL)",
  "% Time Over 180 (mg/dL)",
  "SD",
  "COV",
  "MAG"
)

for (i in seq_len(length(outcomes))) {
  p70 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_70")) +
    labs(y = outcomes_labs[i], x = "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 70)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  
  
  p60 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_60")) +
    labs(y = outcomes_labs[i], x = "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 60)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  cat("\n")
  cat(paste("###", outcomes_labs[i]))
  cat("\n")
  print(p70 + p60)
  cat("\n")
}
```

## CGM Hypoglycemia

```{r,results="asis"}
tab <- tableby(group_category ~ percent_time_under_70 + percent_time_under_60, data = cgm_table_dat)
summary(tab)

under_70_cutoff <- 2 * 2.06
under_60_cutoff <- 2 * 0.687

cgm_table_dat$CGM_hypo_60 <- ifelse(cgm_table_dat$percent_time_under_60 > under_60_cutoff, "Yes", "No")
cgm_table_dat$CGM_hypo_70 <- ifelse(cgm_table_dat$percent_time_under_70 > under_70_cutoff, "Yes", "No")
```

Hypoglycemia by CGM is defined as a % time under 70 or 60 more than 2 SDs away from average among healthy controls 

<60 cutoff

```{r}
under_60_cutoff
```

<70 cutoff

```{r}
under_70_cutoff
```

## Hypoglycemia by CGM by group

```{r, results="asis"}
cgm_tableby <- cgm_table_dat
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "No"] <- "Not Hypo by OGTT"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "No"] <- "Not Hypo by OGTT"

tab <- tableby(group_category ~ CGM_hypo_60 + CGM_hypo_70, data = cgm_tableby)
summary(tab)
```

## Hypoglycemia Concordance

### 70/70

Hypoglycemia by OGTT with cutoff <70 and by CGM with cutoff < 70

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of 
-0.22 among participants with CF and 0.07 among those without CF


```{r, results = "asis"}
tab <- tableby(CGM_hypo_70 ~ hypo_70,
               strata = CF,
               data = cgm_tableby,
               test = F,
               total = F
)
summary(tab)
```

#### Concordance with CF
```{r}
mat <- matrix(data = c(3, 10, 15, 16), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix(data = c(1, 2, 3, 9), nrow = 2, ncol = 2)
Kappa(mat)
```

### 60/60

Hypoglycemia by OGTT with cutoff <60 and by CGM with cutoff < 60

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of
-0.17 among participants with CF. The concordance among those without CF was 
0.42, though due to a small sample size, this was not statistically significant. 


```{r, results = "asis"}
tab <- tableby(CGM_hypo_60 ~ hypo_60,
               strata = CF,
               data = cgm_tableby,
               test = F,
               total = F
)
summary(tab)
```
#### Concordance with CF
```{r}
mat <- matrix(data = c(2, 9, 12, 21), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix(data = c(1, 1, 1, 12), nrow = 2, ncol = 2)
Kappa(mat)
```

# Tim's additional changes for reviewer 1

```{r}
# Plot
ggplot(data = cgm_table_dat, aes_string(y = "percent_time_under_70", x = "CF_hypo_70")) +
  labs(y = "% Time Under 70 (mg/dL)", x = "") +
  geom_jitter(alpha = 0.5, width = 0.1) +
  geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
  geom_hline(yintercept=4.1, linetype="dashed")+
  annotate("text",x=4.4, y=4.5, label="4.1 %") +
  theme_light() +
  ggtitle("Hypoglycemia by OGTT (< 70)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r results='asis'}
# Comparisons
cgm_table_dat$perc_time_70_thresh = cut(cgm_table_dat$percent_time_under_70,
                                        breaks = c(-Inf,4.1,Inf),
                                        labels = c("No","Yes"))
t = tableby(CF_hypo_70~fe(perc_time_70_thresh),data = cgm_table_dat)
summary(t,pfootnote = T,
        labelTranslations = 
          list(perc_time_70_thresh="Percent time under 70 mg/dL > 4.1%"))
```

```{r results='asis'}
cgm_table_dat$CF = factor(cgm_table_dat$CF,levels = c("Yes","No"),labels = c("CF","HC"))
t = tableby(CF~fe(perc_time_70_thresh),data = cgm_table_dat)
summary(t,pfootnote = T,
        labelTranslations = 
          list(perc_time_70_thresh="Percent time under 70 mg/dL > 4.1%"))
```
