---
title: "Racial and Ethnic Differences in ETI: Outcomes"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-height: 10
    fig-width: 10
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
# Libraries
library(tidyverse)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(splines)
library(ggeffects)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
```

# Data cleaning

```{r flowchart}
ggflowchart(flow_chart)
```

# Table 1

## By person

Continuous variables were averaged across all years for each person, and categorical variables are reported as the most common value.

```{r}
t1_participant %>%
  select(-eDWID) %>%
  tbl_summary(
    by = Race,
    type = list(`Years With Data` ~ "continuous"),
    missing_text = "Missing"
  ) %>%
  add_overall()
```

# Encounter-level

```{r}
encounter %>%
  select(Race, encounterlocation, bacterialculturedone:othermicroorganisms1) %>%
  tbl_summary(by = Race, missing_text = "Missing")
```

# LOESS-smoothed outcome plots

```{r functions}
# Plots
loess_plot <- function(data = encounter, outcome, facet = F) {
  p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
    geom_point(
      shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F,
      alpha = 0.1
    ) +
    geom_smooth(method = "loess", se = F) +
    theme_classic()
  if (facet) {
    p <- p + facet_wrap(~Race)
  }
  return(p)
}
```

```{r}
invisible(lapply(continuous_outcomes, function(o) {
  cat("\n")
  cat("\n")
  cat("##", o)
  cat("\n")
  cat("\n")
  print(loess_plot(outcome = o))
}))
```

# Segmented models

The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject.

```{r results='asis'}
invisible(lapply(continuous_outcomes, function(o) {
  # Formula
  f <-
    as.formula(paste0(
      o,
      "~ bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
    ))
  m <- lmer(f, data = encounter)
  # Print table
  m %>%
    tbl_regression() %>%
    print()
  # Plot
  predict_response(m, terms = c("Days[-1095,0,730]", "Race")) %>% 
    plot()
}))
```
