---
title: "Racial and Ethnic Differences in ETI: Outcomes"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-height: 10
    fig-width: 10
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(splines)
library(ggeffects)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
```

# Data cleaning

```{r flowchart}
ggflowchart(flow_chart)
```

# Table 1

## By person

Continuous variables are reported at the most recent visit prior to ETI start for each person, and categorical variables are reported as the most common value.

```{r}
t1_participant %>%
  select(-eDWID) %>%
  tbl_summary(
    by = Race,
    type = list(`Years With Data` ~ "continuous"),
    missing_text = "Missing"
  ) %>%
  add_overall()
```

# Encounter-level

```{r}
encounter %>%
  select(Race, encounterlocation, bacterialculturedone:othermicroorganisms1) %>%
  tbl_summary(by = Race, missing_text = "Missing")
```

# Segmented models

```{r}
# Create change in slope variable
encounter$slope_change_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_change_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_change_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_change_years) <- "Change in Slope"
```

The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject (random slope did not improve the model AIC).

## FEV1 % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# EMMs
emm = emmeans(change_mod,spec = ~Years+Race)

# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
pop_dat = mod_df %>% select(Years:Insurance)
pop_dat$pred <- predict(change_mod, re.form = NA,newdata = pop_dat)
pop_dat = pop_dat%>% distinct()
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(data = pop_dat,aes(y = pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## FEV1/FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## BMI %ile 

### Full cohort

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  bmipercentile ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = bmipercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  bmipercentile ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = bmipercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  bmipercentile ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = bmipercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## Height %ile 

### Full cohort

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  heightpercentile ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = heightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  heightpercentile ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = heightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  heightpercentile ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = heightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## Weight %ile 

### Full cohort

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  weightpercentile ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = weightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  weightpercentile ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = weightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  weightpercentile ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = weightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```
