---
title: "Racial and Ethnic Differences in ETI: Outcomes"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    embed-resources: true
    fig-cap-location: top
    fig-height: 10
    fig-width: 10
    page-layout: full
editor: source
---

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(multcomp)
library(tidyverse)
library(patchwork)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(nlme)
library(broom)
library(emmeans)
library(kableExtra)
library(GLMMadaptive)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI",
  "Windows" = "C:/Users/tvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI",
  "Linux" = "/home/tvigers/OneDrive/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI"
)
github_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tvigers/Documents/GitHub",
  "Windows" = "C:/Users/tvigers/Documents/GitHub",
  "Linux" = "/home/tvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
set.seed(1017)
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
setwd(home_dir)
load(
  "./Data_Cleaned/outcomes_dataset.RData"
)
# Change reference levels
encounter$age_group <- relevel(encounter$age_group, ref = "[18,30)")
```

# Data cleaning

## Questions
- Ask Olga at CFFPR:
  - Should we only include encounterlocation = 1? Should we also take the care episodes file and exclude visits within 2 weeks on either side?
  - Okay to just exclude lung transplant > 1?

```{r flowchart}
ggflowchart(flow_chart, text_size = 3, y_nudge = 0.4, x_nudge = 0.45)
```

# Table 1

## By person

Continuous variables are reported at the most recent visit prior to ETI start for each person, and categorical variables are reported as the most common value.

```{r}
t1_participant %>%
  select(-eDWID) %>%
  tbl_summary(
    by = Race,
    type = list(`Years With Data` ~ "continuous"),
    statistic = list(
      c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"
    ),
    missing_text = "Missing"
  ) %>%
  add_overall()
```

# Encounter-level

```{r}
encounter %>%
  select(
    Race,
    encounterlocation,
    bacterialculturedone:othermicroorganisms1
  ) %>%
  tbl_summary(by = Race, missing_text = "Missing")
```

# New analyses

For the analyses below, we categorized encounter-level data by time from ETI start (e.g. "(-1095.75,-730.5]" includes all encounters from -3 years to -2 years relative to ETI start). Within each time bucket we averaged each outcome (race neutral for lung function measures), then fit an unadjusted model with a time by race interaction and random intercept for subject. Model results are presented as estimated marginal means at each "timepoint."

For BMI, we took each person's age at their last encounter within each time bucket (i.e. their maximum age). We used BMI raw values for those with age >=20 at ETI start, and BMI percentile for those age < 20. 

Mixed model structure was chosen based on FEV1 in the full cohort. We compared models with random intercept and slope to models with random intercept only and three correlation structures: compound symmetry, AR1, and heterogeneous AR1. Based on AIC, the best model was one with random intercept and heterogeneous AR1 (variance differing by race). However, this model structure was relatively unstable and estimates differed when run on different computers, which suggests some weirdness in the likelihood. Because of this, we elected to use a simpler AR1 correlation structure that did not vary over time or between groups. Also, this simplified model worked for all outcomes and data subsets, so using it across the board makes sense for ease of interpretation. We experimented with increasing the number of iterations used by `nlminb` and switching to the `optim` optimizer, but estimates remained unstable or failed to converge entirely. 

```{r annualize data}
annualized <- encounter |>
  filter(Days >= -3 * 365.25 & Days <= 2 * 365.25) |>
  mutate(
    Days_cat = cut(
      Days,
      breaks = c(c(-3, -2, -1, 0, 1, 2) * 365.25),
      labels = c("-3 to -2", "-2 to -1", "-1 to 0", "0 to 1", "1 to 2"),
      dig.lab = 6
    )
  ) |>
  group_by(eDWID, Days_cat) |>
  summarise(
    across(c(Race, encounterage, age_eti_start, age_eti_group), ~ last(.x)),
    across(
      c(gli_fev1_ppred_rn, gli_fvc_ppred_rn, bmivalue, bmipercentile),
      ~ mean(.x, na.rm = T)
    ),
    num_pex = sum(pe_assessment %in% 2:4),
    n_encounters = n(),
    .groups = "drop"
  )
# ID as a factor
annualized$eDWID <- factor(annualized$eDWID)
# Make an integer time column for easier lme() coding
annualized$days_cat_int <- as.integer(annualized$Days_cat)
```

## ppFEV1

```{r ppfev1 selection}
#| eval: false
# Random intercept with correlation structure
ann_mod_comp_sym <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corCompSymm(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
ann_mod_ar1 <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# AR1 changing over time does not converge
ann_mod_het_ar1_time <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Days_cat),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# This model doesn't converge on the Studio but does on the laptop, so we aren't
# using it. Something is up with the likelihood.
ann_mod_het_ar1_race <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Unstructured does not converge
ann_mod_unst <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corSymm(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Days_cat),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Random slope generally doesn't do better than AR1
ann_mod_rs <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ days_cat_int | eDWID,
  data = annualized,
  na.action = na.exclude,
  control = lmeControl(opt = "optim"),
  method = "ML"
)
```

### Full cohort

```{r ppfev1 full}
# AR1 varying by race is the winner! Refit with REML
ann_mod <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
# Compare pre to post
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
# Plot
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("FEV1 % Predicted") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Full Cohort") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/fev1_full.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(emmean, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

### Full cohort adjusted for age at ETI initiation

```{r ppfev1 full adjusted}
# AR1 varying by race is the winner! Refit with REML
ann_mod <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race + age_eti_start,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
# Plot
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("FEV1 % Predicted") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Full Cohort") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/fev1_full_adjusted.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
```

### Age < 18 at ETI initiation

```{r ppfev1 u18}
ann_mod <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized |> filter(age_eti_start < 18),
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("FEV1 % Predicted") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Age < 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/fev1_peds.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(emmean, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

### Age >= 18 at ETI initiation

```{r ppfev1 o18}
ann_mod <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized |> filter(age_eti_start >= 18),
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("FEV1 % Predicted") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Age >= 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/fev1_adults.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(emmean, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

## ppFVC

```{r ppfvc selection}
#| eval: false
# Random intercept with correlation structure
ann_mod_comp_sym <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corCompSymm(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
ann_mod_ar1 <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# AR1 changing over time does not converge
ann_mod_het_ar1_time <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Days_cat),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# See above
ann_mod_het_ar1_race <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Unstructured does not converge
ann_mod_unst <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corSymm(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Days_cat),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Random slope generally doesn't do better than AR1
ann_mod_rs <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ days_cat_int | eDWID,
  data = annualized,
  na.action = na.exclude,
  control = lmeControl(opt = "optim"),
  method = "ML"
)
```

### Full cohort

```{r ppfvc full}
ann_mod <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("FVC % Predicted") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Full Cohort") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/fvc_full.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(emmean, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

### Age < 18 at ETI initiation

```{r ppfvc u18}
ann_mod <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized |> filter(age_eti_start < 18),
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("FVC % Predicted") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Age < 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/fvc_peds.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(emmean, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

### Age >= 18 at ETI initiation

```{r ppfvc o18}
ann_mod <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized |> filter(age_eti_start >= 18),
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("FVC % Predicted") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Age >= 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/fvc_adults.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(emmean, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

## BMI 

### BMI %ile (age < 20 at ETI start)

Because BMI percentile is not a normally distributed variable and is bounded at 0 and 100, we used beta regression (with Smithson–Verkuilen squeeze for 0 values). We fit a beta regression generalized linear mixed model with a logit link on the mean $\mu$. The mean submodel included fixed effects for time bin, race, and their interaction, plus a random intercept per participant and a first-order autoregressive (AR1) correlation across time bins within participant. BMI %ile is presented on the (0, 1) scale.

Similar to the approach used in Miranda Afonso P, Rizopoulos D, Palipana AK, et al. A Joint Model for (Un)Bounded Longitudinal Markers, Competing Risks, and Recurrent Events Using Patient Registry Data. Stat Med. 2025;44(8-9):e70057. doi:10.1002/sim.70057.

```{r bmi u20 Beta}
# Beta regression with random intercept + AR1 within subject
library(glmmTMB)
library(DHARMa)
sub <- annualized |> filter(age_eti_start < 20)
n <- sum(!is.na(sub$bmipercentile))
sub$bmi_p <- (sub$bmipercentile / 100 * (n - 1) + 0.5) / n
ann_mod <- glmmTMB(
  bmi_p ~ Days_cat * Race + ar1(Days_cat + 0 | eDWID),
  data = sub,
  family = beta_family(link = "logit")
)
# sim_res <- simulateResiduals(ann_mod)
# plot(sim_res)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race
) |>
  regrid()
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = response, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(
    aes(ymin = asymp.LCL, ymax = asymp.UCL),
    width = 0.1
  ) +
  ylab("BMI Percentile") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Age < 20 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/bmi_perc.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(response * 100, 2),
      " (",
      round(asymp.LCL * 100, 2),
      ", ",
      round(asymp.UCL * 100, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate * 100, 2),
      " (",
      round(asymp.LCL * 100, 2),
      ", ",
      round(asymp.UCL * 100, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate * 100, 2),
      " (",
      round(asymp.LCL * 100, 2),
      ", ",
      round(asymp.UCL * 100, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate * 100, 2),
      " (",
      round(asymp.LCL * 100, 2),
      ", ",
      round(asymp.UCL * 100, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate * 100, 2),
      " (",
      round(asymp.LCL * 100, 2),
      ", ",
      round(asymp.UCL * 100, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

### BMI raw value (age >= 20 at ETI start)

```{r bmivalue selection}
#| eval: false
# Random intercept with correlation structure
ann_mod_comp_sym <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corCompSymm(form = ~ days_cat_int | eDWID),
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude,
  method = "ML"
)
ann_mod_ar1 <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude,
  method = "ML"
)
# AR1 changing over time does not converge
ann_mod_het_ar1_time <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Days_cat),
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude,
  method = "ML"
)
# See above
ann_mod_het_ar1_race <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude,
  method = "ML"
)
# Unstructured does not converge
ann_mod_unst <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corSymm(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Days_cat),
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude,
  method = "ML"
)
# Random slope generally doesn't do better than AR1
ann_mod_rs <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ days_cat_int | eDWID,
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude,
  control = lmeControl(opt = "optim"),
  method = "ML"
)
```

```{r bmi o20}
ann_mod <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude
)
# performance::check_model(ann_mod)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race,
  pbkrtest.limit = 1e6
)
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  labs(
    y = expression(BMI ~ (kg / m^2)),
    x = "Years prior to/after ETI initiation"
  ) +
  ggtitle("Age >= 20 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/bmi_value.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(emmean, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(lower.CL, 2),
      ", ",
      round(upper.CL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

## Pulmonary exacerbations

Pulmonary exacerbations are count data, so we modeled them using a Poisson mixed effect model with AR1 correlation structure across time and a participant-level random effect. The estimates below are rates of exacerbation, so a rate of 0.5 indicates that you'd expect an exacerbation every 2 years. 

### Full cohort

```{r}
ann_mod <- glmmTMB(
  num_pex ~ Days_cat * Race + ar1(Days_cat + 0 | eDWID),
  data = annualized,
  family = poisson
)
# sim_res <- simulateResiduals(ann_mod)
# plot(sim_res)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race
) |>
  regrid()
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = rate, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(
    aes(ymin = asymp.LCL, ymax = asymp.UCL),
    width = 0.1
  ) +
  ylab("PEx Rate") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Full Cohort") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/pex_full.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(rate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

### Age < 18 at ETI initiation

```{r}
ann_mod <- glmmTMB(
  num_pex ~ Days_cat * Race + ar1(Days_cat + 0 | eDWID),
  data = annualized |> filter(age_eti_start < 18),
  family = poisson
)
# sim_res <- simulateResiduals(ann_mod)
# plot(sim_res)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race
) |>
  regrid()
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = rate, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(
    aes(ymin = asymp.LCL, ymax = asymp.UCL),
    width = 0.1
  ) +
  ylab("PEx Rate") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Age < 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/pex_peds.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(rate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```

### Age >= 18 at ETI initiation

```{r}
ann_mod <- glmmTMB(
  num_pex ~ Days_cat * Race + ar1(Days_cat + 0 | eDWID),
  data = annualized |> filter(age_eti_start >= 18),
  family = poisson
)
# sim_res <- simulateResiduals(ann_mod)
# plot(sim_res)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat + Race
) |>
  regrid()
emm_change <- contrast(
  emm,
  list("Pre- to Post- Change" = c(0, 0, -1, 1, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
p <- ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = rate, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(
    aes(ymin = asymp.LCL, ymax = asymp.UCL),
    width = 0.1
  ) +
  ylab("PEx Rate") +
  xlab("Years prior to/after ETI initiation") +
  ggtitle("Age >= 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
p
ggsave(
  "./Dissemination/Figures/pex_adults.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Organize tables per Kam
emm |>
  data.frame() |>
  mutate(
    combined = paste0(
      round(rate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    )
  ) |>
  select(Race, Days_cat, combined) |>
  pivot_wider(
    names_from = Days_cat,
    values_from = combined
  ) |>
  kable(caption = "Time to and following ETI Initiation (Time 0)")
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 1 Year Pre-ETI to 1 year Post-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
# Same again for change pre-ETI
emm_change <- contrast(
  emm,
  list("Pre-ETI Change" = c(-1, 0, 1, 0, 0)),
  by = "Race"
)
# Now compare those changes pairwise across races
emm_change_comp <- contrast(
  emm_change,
  method = "trt.vs.ctrl",
  by = NULL,
  ref = 1
)
summary(emm_change, infer = T) |>
  data.frame() |>
  mutate(
    Change = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(Race, Change, p.value) |>
  kable(
    caption = "Average Change from 3 Year Pre-ETI to 1 year Pre-ETI"
  )
summary(emm_change_comp, infer = T) |>
  data.frame() |>
  mutate(
    Difference = paste0(
      round(estimate, 2),
      " (",
      round(asymp.LCL, 2),
      ", ",
      round(asymp.UCL, 2),
      ")"
    ),
    p.value = format.pval(p.value, eps = 0.001)
  ) |>
  select(contrast, Difference, p.value) |>
  rename(Comparison = contrast) |>
  kable(
    caption = "Differences in Change by Race"
  )
```
