---
title: "Racial and Ethnic Differences in ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
# Libraries
library(segmented)
library(Hmisc)
library(tidyverse)
library(gtsummary)
library(lmerTest)
library(nlme)
library(broom.mixed)
library(emmeans)
library(gt)
library(performance)
library(survival)
library(ggsurvfit)
library(ggstats)
library(naniar)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
```

# Race and ethnicity breakdown

The table below shows the race and ethnicity proportions for the entire cohort included in the CFFPR dataset. I decided to combine race and ethnicity with Hispanic "overriding" other racial identities. So if a participant was reported as both Black and Hispanic, they were consider Hispanic for the following analyses. Those with missing data for Hispanic ethnicity were moved to the "Other" category.

```{r}
demo %>%
  select(race) %>%
  tbl_summary(missing_text = "Missing") %>%
  bold_labels()
```

## Race by year

The table below shows the proportions of race and ethnicities included in the dataset in a given year. Most people are seen at least once a year, but this table serves as a sanity check to make sure that groups were generally seen equal amounts each year.

```{r}
annual %>%
  select(race, ReviewYear) %>%
  tbl_summary(by = ReviewYear, missing_text = "Missing") %>%
  bold_labels()
```

# Participant characteristics

```{r}
t1_vars <- c(
  "Gender", "F508_num", "mean_ppFEV", "Age_YrEnd",
  "mean_ppFEV", "weight_last", "weight_perc_last", "bmi_last",
  "bmi_perc_last", "Vx445comb"
)
```

## Characteristics by race

```{r}
demo %>%
  select(Gender, F508_num, race) %>%
  tbl_summary(by = race, missing_text = "Missing") %>%
  bold_labels()
```

## Characteristics by year

```{r results='asis'}
#| message: false
set_gtsummary_theme(theme_gtsummary_compact(), quiet = T)
years <- split.data.frame(annual, annual$ReviewYear)
invisible(lapply(names(years), function(n) {
  cat("\n")
  cat("###", n)
  cat("\n")
  d <- years[[n]]
  t <- d %>%
    select(all_of(t1_vars), race) %>%
    tbl_summary(by = race, missing_text = "Missing") %>%
    bold_labels()
  print(t)
  cat("\n")
}))
reset_gtsummary_theme()
```

# ETI eligibility

Trikafta originally approved October 21, 2019 for patients 12+ years of age. Approved for ages 6 - 11 on June 9, 2021.

The following tables have been aggregated by year, and therefore may miss some information regarding eligibility and ETI start. For example, if someone turned 12 on 12/30/2019, they would technically be considered eligible for ETI, even though realistically we probably wouldn't expect them to start that year.

## Proportions

### Overall by year

#### All patients

```{r}
annual %>%
  select(eti_elig, contains("Vx"), ReviewYear) %>%
  tbl_summary(
    by = ReviewYear, missing_text = "Missing",
    type = all_dichotomous() ~ "categorical"
  ) %>%
  bold_labels()
```

#### Eligible and prescribed off-label only

```{r}
annual %>%
  filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
  select(eti_elig, contains("Vx"), ReviewYear) %>%
  tbl_summary(
    by = ReviewYear, missing_text = "Missing",
    type = all_dichotomous() ~ "categorical"
  ) %>%
  add_ci() %>%
  add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
  separate_p_footnotes() %>%
  bold_labels()
```

### By race and year

```{r results='asis'}
invisible(lapply(c("2020", "2021", "2022"), function(n) {
  cat("\n")
  cat("###", n)
  cat("\n")
  cat("\n")
  cat("#### All patients")
  cat("\n")
  d <- years[[n]]
  t <- d %>%
    select(eti_elig, contains("Vx"), race) %>%
    tbl_summary(
      by = race, missing_text = "Missing",
      type = all_dichotomous() ~ "categorical"
    ) %>%
    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
    separate_p_footnotes() %>%
    bold_labels()
  print(t)
  cat("\n")
  cat("\n")
  cat("#### Eligible and prescribed off-label only")
  cat("\n")
  t <- d %>%
    filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
    select(eti_elig, contains("Vx"), race) %>%
    tbl_summary(
      by = race, missing_text = "Missing",
      type = all_dichotomous() ~ "categorical"
    ) %>%
    add_ci() %>%
    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
    separate_p_footnotes() %>%
    bold_labels()
  print(t)
  cat("\n")
}))
```

### Figure

The figure below includes those eligible for ETI and anyone who may have been prescribed ETI off label. Error bars represent 95% confidence intervals.

```{r}
p <- annual %>%
  filter(eti_elig == "Yes" | Vx445comb == "Yes", ReviewYear > 2019) %>%
  select(eDWID, race, ReviewYear, contains("Vx")) %>%
  group_by(ReviewYear, race) %>%
  summarise(
    perc = mean(Vx445comb == "Yes", na.rm = T),
    se = sqrt(perc * (1 - perc) / n()),
    lower = perc - (1.96 * se), upper = perc + (1.96 * se),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = ReviewYear, y = perc * 100, fill = race)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_errorbar(aes(ymin = lower * 100, ymax = upper * 100),
    position = "dodge", linewidth = 0.7
  ) +
  xlab("Year") +
  ylab("Percentage Prescribed ETI") +
  theme_bw(base_size = 16) +
  theme(legend.title = element_blank(), legend.text = element_text(size = 14)) +
  scale_fill_manual(values = c(
    "#264653", "#2A9D8F", "#E9C46A",
    "#F4A261", "#E76F51"
  ))
p
```

```{r}
#| warning: false
# Save image
ggsave("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Figures/bar_chart.png",
  plot = p, device = "png", width = 12, height = 8, units = "in"
)
```

## Survival analysis

For the following analysis, the start time ($t=0$) was the approximate date that each person became eligible for ETI. Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates, and then used the approximate DOB to determine date of eligibility. Encounters prior to ETI eligibility were excluded, and only time to first encounter on ETI was considered (i.e. we did not fit a multi-state model to account for those who started and then stopped the medication).

```{r}
# Create survival dataset
surv_dat <- encounter %>%
  filter(eti_elig == "Yes") %>%
  mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
  filter(stop > 0) %>%
  group_by(eDWID) %>%
  mutate(start = lag(stop, default = 0)) %>%
  select(
    eDWID, race, Insurance, family_income, patient_education, father_education,
    mother_education, encounterdate, encounterage, gli_fev1_ppred_rn,
    eti_elig_date, Vx445comb, start, stop
  )
surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1
# Fit
y <- "Surv(time = start, time2 = stop, event = Vx445comb)"
surv_f <- as.formula(paste0(y, "~race"))
fit <- survfit(surv_f, id = eDWID, data = surv_dat)
```

### Days to percentile started on ETI

```{r}
# Quantiles
tbl_survfit(fit, probs = c(0.25, 0.5, 0.75, 0.9))
```

### Plots

```{r}
#| warning: false
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
p <- fit %>%
  ggsurvfit(type = "risk") +
  xlim(c(0, 365 * 2)) +
  add_confidence_interval() +
  add_quantile() +
  ylab("Proportion on ETI") + xlab("Days From Eligibility") +
  theme_bw(base_size = 16) +
  theme(legend.title = element_blank(), legend.text = element_text(size = 14))
p
```

```{r}
#| warning: false
ggsave("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Figures/survival.png",
  plot = p, device = "png", width = 12, height = 8, units = "in"
)
```

Without Asian subgroup for clearer plot:

```{r}
#| warning: false
fit <- survfit(surv_f, id = eDWID, data = surv_dat[surv_dat$race != "Asian", ])
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
p <- fit %>%
  ggsurvfit(type = "risk") +
  xlim(c(0, 365 * 2)) +
  add_confidence_interval() +
  add_quantile() +
  ylab("Proportion on ETI") + xlab("Days From Eligibility") +
  theme_bw(base_size = 16) +
  theme(legend.title = element_blank(), legend.text = element_text(size = 14))
p
```

```{r}
#| warning: false
ggsave("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Figures/survival_no_asian.png",
  plot = p, device = "png", width = 12, height = 8, units = "in"
)
```

### Cox PH models

#### Base model (unadjusted)

```{r}
cox_res <- coxph(surv_f, data = surv_dat, id = eDWID)
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F, exponentiate = T
)
```

##### Missing data

```{r}
gg_miss_var(surv_dat, show_pct = T, facet = race)
```

#### Adjusted for combined insurance status only

Health Insurance Policy (e.g. Private Insurance) = Private

Indian Health Service = Public

Medicaid = Public

Medicare = Public

More than one = More than one

No Insurance = No Insurance

Other = No Insurance (per Liz Cromwell)

State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc. = Public

TriCare or other military health plan = Tricare

```{r}
surv_dat$Insurance <- factor(surv_dat$Insurance,
  levels = c(
    "Health Insurance Policy (e.g. Private Insurance)",
    "Indian Health Service",
    "Medicaid",
    "Medicare",
    "More than one",
    "No Insurance",
    "Other",
    "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
    "TriCare or other military health plan"
  ),
  labels = c(
    "Private", "Public", "Public", "Public", "More Than One", "No Insurance",
    "No Insurance", "Public", "Tricare"
  )
)
# FEV1 to categorical
surv_dat$fev1_cat <- cut(surv_dat$gli_fev1_ppred_rn,
  breaks = c(0, 40, 80, 100, Inf),right = F
)
label(surv_dat$fev1_cat)=label(surv_dat$gli_fev1_ppred_rn)
# New fit
new_surv_f <- update(surv_f, . ~ . + Insurance)
cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F, exponentiate = T
)
```

#### Adjusted for insurance, age group, and FEV1

```{r}
surv_dat$`Age Group` <- cut(surv_dat$encounterage,
  breaks = c(6, 12, 18, 30, Inf), right = F
)
new_surv_f <- update(surv_f, . ~ . + Insurance + `Age Group` +
  gli_fev1_ppred_rn)
cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F, exponentiate = T
)
```

# Number of clinic visits

For each participant, we divided the total number of visits by the number of unique years that they were included in the registry data. Visits per year are presented by race.

```{r}
encounter %>%
  group_by(eDWID, race) %>%
  summarise(
    `Visits per Year` = n() / length(unique(reviewyear)),
    .groups = "drop"
  ) %>%
  select(-eDWID) %>%
  tbl_summary(by = race)
```
