---
title: "Racial and Ethnic Differences in ETI: Outcomes"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: false
    self-contained: true
    fig-cap-location: top
    fig-height: 10
    fig-width: 10
    page-layout: full
editor: source
---

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(multcomp)
library(tidyverse)
library(patchwork)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(nlme)
library(broom)
library(emmeans)
library(kableExtra)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI"
)
github_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
setwd(home_dir)
load(
  "./Data_Cleaned/outcomes_dataset.RData"
)
# Change reference levels
encounter$age_group <- relevel(encounter$age_group, ref = "[18,30)")
```

# Data cleaning

## Questions
- Ask Olga at CFFPR:
  - Should we only include encounterlocation = 1? Should we also take the care episodes file and exclude visits within 2 weeks on either side?
  - Okay to just exclude lung transplant > 1?

```{r flowchart}
ggflowchart(flow_chart, text_size = 3, y_nudge = 0.4, x_nudge = 0.45)
```

# Table 1

## By person

Continuous variables are reported at the most recent visit prior to ETI start for each person, and categorical variables are reported as the most common value.

```{r}
t1_participant %>%
  select(-eDWID) %>%
  tbl_summary(
    by = Race,
    type = list(`Years With Data` ~ "continuous"),
    statistic = list(
      c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"
    ),
    missing_text = "Missing"
  ) %>%
  add_overall()
```

# Encounter-level

```{r}
encounter %>%
  select(
    Race,
    encounterlocation,
    bacterialculturedone:othermicroorganisms1
  ) %>%
  tbl_summary(by = Race, missing_text = "Missing")
```

# Unadjusted analyses

For the analyses below, we categorized encounter-level data by time from ETI start (e.g. "(-1095.75,-730.5]" includes all encounters from -3 years to -2 years relative to ETI start). Within each time bucket we averaged each outcome (race neutral for lung function measures), then fit an unadjusted model with a time by race interaction and random intercept for subject. Model results are presented as estimated marginal means at each "timepoint."

For BMI, we took each person's age at their last encounter within each time bucket (i.e. their maximum age). We used BMI raw values for those with age >=20 at ETI start, and BMI percentile for those age < 20. 

```{r annualize data}
annualized <- encounter |>
  filter(Days >= -3 * 365.25 & Days <= 2 * 365.25) |>
  mutate(
    Days_cat = cut(
      Days,
      breaks = c(c(-3, -2, -1, 0, 1, 2) * 365.25),
      dig.lab = 6
    )
  ) |>
  group_by(eDWID, Days_cat) |>
  summarise(
    across(c(Race, encounterage, age_eti_start, age_eti_group), ~ last(.x)),
    across(
      c(gli_fev1_ppred_rn, gli_fvc_ppred_rn, bmivalue, bmipercentile),
      ~ mean(.x, na.rm = T)
    ),
    .groups = "drop"
  )
# ID as a factor
annualized$eDWID = factor(annualized$eDWID)
# Make an integer time column for easier lme() coding
annualized$days_cat_int = as.integer(annualized$Days_cat)
```

## ppFEV1

Mixed model structure was chosen based on FEV1 in the full cohort. We compared models with random intercept and slope to models with random intercept only and three correlation structures: compound symmetry, AR1, and heterogenous AR1. Based on AIC, the best model with one with random intercept and heterogenous AR1. Diagnostic plots for this model looked beautiful and pristine. 

```{r ppfev1 selection}
# Random intercept with correlation structure
ann_mod_comp_sym <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corCompSymm(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
ann_mod_ar1 <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# AR1 changing over time does not converge
# ann_mod_het_ar1_time <- lme(
#   gli_fev1_ppred_rn ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corAR1(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_het_ar1_race <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Unstructured does not converge
# ann_mod_unst <- lme(
#   gli_fev1_ppred_rn ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corSymm(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_rs <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ days_cat_int | eDWID,
  data = annualized,
  na.action = na.exclude,
  control = lmeControl(opt = "optim"),
  method = "ML"
)
kable(performance::compare_performance(
  ann_mod_comp_sym,
  ann_mod_ar1,
  ann_mod_het_ar1_race,
  ann_mod_rs
))
```

### Full cohort

```{r ppfev1 full}
# AR1 varying by race is the winner! Refit with REML
ann_mod = lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("ppFEV1") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Full Cohort") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

### Age < 18 at ETI initiation

```{r ppfev1 u18}
ann_mod <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized |> filter(age_eti_start < 18),
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("ppFEV1") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Age < 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

### Age >= 18 at ETI initiation

```{r ppfev1 o18}
ann_mod <- lme(
  gli_fev1_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized |> filter(age_eti_start >= 18),
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("ppFEV1") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Age >= 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

## ppFVC

```{r ppfvc selection}
# Random intercept with correlation structure
ann_mod_comp_sym <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corCompSymm(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
ann_mod_ar1 <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# AR1 changing over time does not converge
# ann_mod_het_ar1_time <- lme(
#   gli_fvc_ppred_rn ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corAR1(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_het_ar1_race <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Unstructured does not converge
# ann_mod_unst <- lme(
#   gli_fvc_ppred_rn ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corSymm(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_rs <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ days_cat_int | eDWID,
  data = annualized,
  na.action = na.exclude,
  control = lmeControl(opt = "optim"),
  method = "ML"
)
kable(performance::compare_performance(
  ann_mod_comp_sym,
  ann_mod_ar1,
  ann_mod_het_ar1_race,
  ann_mod_rs
))
```

### Full cohort

```{r ppfvc full}
ann_mod = lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("ppFVC") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Full Cohort") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

### Age < 18 at ETI initiation

```{r ppfvc u18}
ann_mod <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized |> filter(age_eti_start < 18),
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("ppFVC") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Age < 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

### Age >= 18 at ETI initiation

```{r ppfvc o18}
ann_mod <- lme(
  gli_fvc_ppred_rn ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized |> filter(age_eti_start >= 18),
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("ppFVC") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Age >= 18 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

## BMI 

### BMI %ile (age < 20 at ETI start)

```{r bmipercentile selection}
# Random intercept with correlation structure
ann_mod_comp_sym <- lme(
  bmipercentile ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corCompSymm(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
ann_mod_ar1 <- lme(
  bmipercentile ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# AR1 changing over time does not converge
# ann_mod_het_ar1_time <- lme(
#   bmipercentile ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corAR1(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_het_ar1_race <- lme(
  bmipercentile ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Unstructured does not converge
# ann_mod_unst <- lme(
#   bmipercentile ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corSymm(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_rs <- lme(
  bmipercentile ~ Days_cat * Race,
  random = ~ days_cat_int | eDWID,
  data = annualized,
  na.action = na.exclude,
  control = lmeControl(opt = "optim"),
  method = "ML"
)
kable(performance::compare_performance(
  ann_mod_comp_sym,
  ann_mod_ar1,
  ann_mod_het_ar1_race,
  ann_mod_rs
))
```

```{r bmi u20}
ann_mod <- lme(
  bmipercentile ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized |> filter(age_eti_start < 20),
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("BMI %ile") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Age < 20 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

### BMI raw value (age >= 20 at ETI start)

```{r bmivalue selection}
# Random intercept with correlation structure
ann_mod_comp_sym <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corCompSymm(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
ann_mod_ar1 <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# AR1 changing over time does not converge
# ann_mod_het_ar1_time <- lme(
#   bmivalue ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corAR1(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_het_ar1_race <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized,
  na.action = na.exclude,
  method = "ML"
)
# Unstructured does not converge
# ann_mod_unst <- lme(
#   bmivalue ~ Days_cat * Race,
#   random = ~ 1 | eDWID,
#   correlation = corSymm(form = ~ days_cat_int | eDWID),
#   weights = varIdent(form = ~ 1 | Days_cat),
#   data = annualized,
#   na.action = na.exclude,
#   method = "ML"
# )
ann_mod_rs <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ days_cat_int | eDWID,
  data = annualized,
  na.action = na.exclude,
  control = lmeControl(opt = "optim"),
  method = "ML"
)
kable(performance::compare_performance(
  ann_mod_comp_sym,
  ann_mod_ar1,
  ann_mod_het_ar1_race,
  ann_mod_rs
))
```

```{r bmi o20}
ann_mod <- lme(
  bmivalue ~ Days_cat * Race,
  random = ~ 1 | eDWID,
  correlation = corAR1(form = ~ days_cat_int | eDWID),
  weights = varIdent(form = ~ 1 | Race),
  data = annualized |> filter(age_eti_start >= 20),
  na.action = na.exclude
)
# Summarize
emm <- emmeans(
  ann_mod,
  specs = ~ Days_cat | Race,
  pbkrtest.limit = 1e6
)
ggplot(
  data.frame(emm),
  aes(x = Days_cat, y = emmean, group = Race, color = Race)
) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  ylab("BMI") +
  xlab("Days prior to/after ETI initiation") +
  ggtitle("Age >= 20 at ETI initiation") +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  ) +
  theme_bw()
kable(emm)
```

# Segmented models

```{r}
# Create change in slope variable
encounter$slope_c_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_c_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_c_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_c_years) <- "Change in Slope"
# Prior modulator use
prior_mods <- encounter %>%
  select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
  filter(Days < 0) %>%
  group_by(eDWID) %>%
  summarise(
    prior_modulator = any(vx661comb == 1) |
      any(vx445comb == 1) |
      any(Vx770 == 1) |
      any(Vx809comb)
  )
prior_mods$prior_modulator <-
  fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
prior_mods$prior_modulator <- factor(
  prior_mods$prior_modulator,
  levels = c(T, "No"),
  labels = c("Yes", "No")
)
encounter <- left_join(encounter, prior_mods, by = join_by(eDWID))
encounter$prior_modulator <- relevel(encounter$prior_modulator, ref = "No")
# On enzymes for pancreatic insufficiency
encounter$isOnEnzymes <-
  factor(encounter$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
# Pseudomonas
encounter$pseudomonasaeruginosa[is.na(encounter$pseudomonasaeruginosa)] <- 0
encounter$pseudomonasaeruginosa <- factor(
  encounter$pseudomonasaeruginosa,
  levels = 0:1,
  labels = c("No", "Yes")
)
```

The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject (random slope did not improve the model AIC).

## FEV1 % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    age_group +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
  Years = -3:2,
  Race = unique(encounter$Race),
  age_group = "[18,30)",
  Insurance = "Private",
  prior_modulator = "Yes",
  isOnEnzymes = "Yes",
  pseudomonasaeruginosa = "No"
))
pop_dat$slope_c_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
p <- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("FEV1 % Predicted") +
  theme_classic(base_size = 15) +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  )
p
ggsave(
  "./Dissemination/Figures/fig2.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt(caption = "Predicted Values at Various Timepoints") %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    age_group +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
p <- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("FEV1 % Predicted (2012 Equation)") +
  theme_classic() +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  )
p
ggsave(
  "./Dissemination/Figures/fev1_2012.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 2),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

## FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    age_group +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
p <- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("FVC % Predicted") +
  theme_classic(base_size = 15) +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  )
p
ggsave(
  "./Dissemination/Figures/figS1.png",
  plot = p,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    age_group +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

## FEV1/FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    age_group +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    age_group +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

## Number of Pulmonary Exacerbations

For some reason, exacerbation models with the encounter-level data are just not working. I think it may be because there are often several pulmonary exacerbation visits very close together, so including the "Days" variable is causing problems. In the end I used a very simple logistic regression model on annualized PEx data. See `/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R` for details on making the exacerbation dataset.

### Full cohort

```{r}
# Zero-inflated models
# mod_p <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.poisson(), zi_fixed = ~1
# )
# mod_nb <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.negative.binomial(), zi_fixed = ~1
# )
pex_logistic <- glmer(
  any_pex ~ ETI * Race + (1 | eDWID),
  data = pex_data,
  control = glmerControl(optimizer = "bobyqa"),
  family = "binomial"
)
pex_logistic |> tbl_regression()
means <- data.frame(emmeans(pex_logistic, ~ Race + ETI, type = "response"))
means %>%
  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0)
aa_post <- c(0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 0, 1, 0, 0, 0, 0, 0)
hl_post <- c(0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(1, 0, 0, 0, 0, 0, 0, 0)
nhw_post <- c(0, 0, 0, 0, 1, 0, 0, 0)
other_pre <- c(0, 0, 0, 1, 0, 0, 0, 0)
other_post <- c(0, 0, 0, 0, 0, 0, 0, 1)
```

The table above shows the predicted probability of a pulmonary exacerbation pre- and post-ETI. So for example, NHW pwCF had a PEx probability of `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)`) post-ETI. Conversely, Black pwCF had a PEx probability of `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)`) post-ETI. 

### Age < 18 at ETI initiation

```{r}
pex_logistic <- glmer(
  any_pex ~ ETI * Race + (1 | eDWID),
  data = pex_data[pex_data$age_eti_group == "[-Inf,18)", ],
  control = glmerControl(optimizer = "bobyqa"),
  family = "binomial"
)
pex_logistic |> tbl_regression()
means <- emmeans(pex_logistic, ~ Race + ETI, type = "response")
data.frame(means) %>%
  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

### Age >= 18 at ETI initiation

```{r}
pex_logistic <- glmer(
  any_pex ~ ETI * Race + (1 | eDWID),
  data = pex_data[pex_data$age_eti_group == "[18, Inf)", ],
  control = glmerControl(optimizer = "bobyqa"),
  family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

## BMI

### BMI %ile (age < 20 at ETI start)

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  bmipercentile ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_start < 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
p3a <- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("BMI Percentile") +
  theme_classic(base_size = 15) +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  )
p3a
ggsave(
  "./Dissemination/Figures/fig3a.png",
  plot = p3a,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(bmipercentile = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### BMI raw value (age >= 20 at ETI start)

```{r}
change_mod <- lmer(
  bmivalue ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_start >= 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
p3b <- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("BMI (kg/m2)") +
  theme_classic(base_size = 15) +
  scale_color_manual(
    values = c("#1e212b", "#4d8b31", "#ffc800", "#ff8427"),
    labels = c("White", "Black", "Hispanic", "Other")
  )
p3b
ggsave(
  "./Dissemination/Figures/fig3b.png",
  plot = p3b,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
p <- p3a / p3b + plot_layout(guides = "collect")
ggsave(
  "./Dissemination/Figures/fig3ab.png",
  plot = p,
  width = 9,
  height = 7,
  units = "in",
  dpi = 600
)
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(bmivalue = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

## Height

### Height %ile (age < 20 at ETI start)

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  heightpercentile ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_start < 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("heightpercentile") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(heightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Height raw value (age >= 20 at ETI start)

```{r}
change_mod <- lmer(
  height ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_start >= 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("height") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(height = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

## Weight

### Weight %ile (age < 20 at ETI start)

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  weightpercentile ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_start < 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("weightpercentile") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(weightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```

### Weight raw value (age >= 20 at ETI start)

```{r}
change_mod <- lmer(
  weight ~
    Years *
    Race +
    slope_c_years * Race +
    Insurance +
    prior_modulator +
    isOnEnzymes +
    pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_start >= 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("weight") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(
    intercept = T,
    estimate_fun = ~ style_number(.x, digits = 1),
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  )
# Table of predicted values
pop_dat %>%
  select(-slope_c_years) %>%
  rename(weight = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_c,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_c,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_c,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 1)
```
