---
title: "Catecholamines and Hypoglycemia"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gt)
library(gtsummary)
library(lmerTest)
library(emmeans)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import cleaned data
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
  na.strings = ""
)
# For this analysis, we only want those with catecholamines
df <- df %>%
  group_by(study_id, Date) %>%
  mutate(catecholamines = any(!is.na(Epinephrine) | !is.na(Norepinephrine))) %>%
  filter(catecholamines) %>%
  ungroup()
```

# Demographics at first visit

BMI percentiles were calculated using the `childsds` R package, assuming that male = 1 and female = 2 for the `Sex` variable. Those with age > 20 were considered age = 20 for the purposes of percentile calculations.

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_visit:ethnicity, cftr_mutation_1,
    cftr_mutation_2
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

# Hypoglycemia defined as < 70 mg/dL at any point

## Demographics

IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.

```{r}
#| label: tbl-70
#| tbl-cap: BG < 70 at Any Point
#| warning: false
df %>%
  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
  pivot_wider(
    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
    id_cols = c(study_id:cftr_mutation_2, Hypo70)
  ) %>%
  select(Hypo70, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

# Hypoglycemia defined as < 60 mg/dL at any point

## Demographics

IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.

```{r}
#| label: tbl-60
#| tbl-cap: BG < 60 at Any Point
#| warning: false
df %>%
  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
  pivot_wider(
    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
    id_cols = c(study_id:cftr_mutation_2, Hypo60)
  ) %>%
  select(Hypo60, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
  tbl_summary(by = Hypo60, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

<!-- ## LS mean glucose by hypoglycemia status -->

<!-- ```{r mixed models 70} -->
<!-- # Time as a factor -->
<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
<!-- # Fit -->
<!-- mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df) -->
<!-- means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Long insulins -->
<!-- insulin_long_df <- df %>% -->
<!--   pivot_longer(all_of(insulin), -->
<!--     names_to = "Time", values_to = "Insulin", -->
<!--     names_pattern = "insulin_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
<!-- ) -->
<!-- # Fit -->
<!-- ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df) -->
<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Combine -->
<!-- means <- means %>% select(-SE, -df) -->
<!-- means$Variable <- "Glucose" -->
<!-- ins_means <- ins_means %>% select(-SE, -df) -->
<!-- ins_means$Variable <- "Insulin" -->
<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
<!-- levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-mixed-model70 -->
<!-- #| fig-cap: Model means by timepoint -->
<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
<!--   geom_line() + -->
<!--   ylab("EM Mean") + -->
<!--   theme_bw() + -->
<!--   facet_wrap(~Hypo70) -->
<!-- ``` -->

<!-- # Hypoglycemia defined as < 60 mg/dL at any point -->

<!-- ## Demographics -->

<!-- ```{r} -->
<!-- #| label: tbl-60 -->
<!-- #| tbl-cap: BG < 60 at Any Point -->
<!-- df %>% -->
<!--   select( -->
<!--     age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2, -->
<!--     all_of(glucose), Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno, -->
<!--     iAUC30gluc:iAUC180ins, homa_ir -->
<!--   ) %>% -->
<!--   tbl_summary(by = Hypo60, missing_text = "Missing") %>% -->
<!--   add_overall() %>% -->
<!--   add_p() %>% -->
<!--   separate_p_footnotes() -->
<!-- ``` -->

<!-- ## LS mean glucose by hypoglycemia status -->

<!-- ```{r mixed models 60} -->
<!-- # Long OGTTs -->
<!-- long_df <- df %>% -->
<!--   pivot_longer(timepoint_0_min:timepoint_180_min, -->
<!--     names_to = "Time", values_to = "Glucose", -->
<!--     names_pattern = "timepoint_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
<!-- # Fit -->
<!-- mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df) -->
<!-- means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Long insulins -->
<!-- insulin_long_df <- df %>% -->
<!--   pivot_longer(all_of(insulin), -->
<!--     names_to = "Time", values_to = "Insulin", -->
<!--     names_pattern = "insulin_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
<!-- ) -->
<!-- # Fit -->
<!-- ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df) -->
<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Combine -->
<!-- means <- means %>% select(-SE, -df) -->
<!-- means$Variable <- "Glucose" -->
<!-- ins_means <- ins_means %>% select(-SE, -df) -->
<!-- ins_means$Variable <- "Insulin" -->
<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
<!-- levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-mixed-model60 -->
<!-- #| fig-cap: Model means by timepoint -->
<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
<!--   geom_line() + -->
<!--   ylab("EM Mean") + -->
<!--   theme_bw() + -->
<!--   facet_wrap(~Hypo60) -->
<!-- ``` -->

<!-- # Hypoglycemia defined as <= 50 mg/dL at any point -->

<!-- ## Participant demographics -->

<!-- ```{r} -->
<!-- #| label: tbl-50 -->
<!-- #| tbl-cap: BG <= 50 at Any Point -->
<!-- df %>% -->
<!--   select( -->
<!--     age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2, -->
<!--     all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno, -->
<!--     iAUC30gluc:iAUC180ins, homa_ir -->
<!--   ) %>% -->
<!--   tbl_summary(by = Hypo50, missing_text = "Missing") %>% -->
<!--   add_overall() %>% -->
<!--   add_p() %>% -->
<!--   separate_p_footnotes() -->
<!-- ``` -->

<!-- ## LS mean glucose by hypoglycemia status -->

<!-- ```{r mixed models 50} -->
<!-- # Long OGTTs -->
<!-- long_df <- df %>% -->
<!--   pivot_longer(timepoint_0_min:timepoint_180_min, -->
<!--     names_to = "Time", values_to = "Glucose", -->
<!--     names_pattern = "timepoint_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
<!-- # Fit -->
<!-- mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df) -->
<!-- means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Long insulins -->
<!-- insulin_long_df <- df %>% -->
<!--   pivot_longer(all_of(insulin), -->
<!--     names_to = "Time", values_to = "Insulin", -->
<!--     names_pattern = "insulin_(.*)_min" -->
<!--   ) -->
<!-- # Time as a factor -->
<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
<!-- ) -->
<!-- # Fit -->
<!-- ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df) -->
<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4)) -->
<!-- # Combine -->
<!-- means <- means %>% select(-SE, -df) -->
<!-- means$Variable <- "Glucose" -->
<!-- ins_means <- ins_means %>% select(-SE, -df) -->
<!-- ins_means$Variable <- "Insulin" -->
<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
<!-- levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: fig-mixed-model50 -->
<!-- #| fig-cap: Model means by timepoint -->
<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
<!--   geom_point() + -->
<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
<!--   geom_line() + -->
<!--   ylab("EM Mean") + -->
<!--   theme_bw() + -->
<!--   facet_wrap(~Hypo50) -->
<!-- ``` -->
