---
title: "EnVision CF Analyses"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(arsenal)
library(pROC)
library(gt)
library(patchwork)
library(lmerTest)
library(emmeans)
library(plotly)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Set seed for bootstrap
set.seed(1017)
# Import data
setwd(home_dir)
# See GitHub/CF/Christine Chan/EnVision CF/clean_envision_data.R
load("./Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.RData")
```

# ROC Analyses

```{r}
# Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
# Remove CC0083 per Christine (weird CGM)
df <- final_df %>%
  filter(!is.na(mean_sensor), !is.na(Diagnosis), study_id != "CC0083")
# Change specific diagnoses per Christine
df$Diagnosis[df$study_id == "WU0019"] <- c("NGT", "NGT")
df$Diagnosis[df$study_id == "IA0080"] <- "IGT"
df$Dysglycemia[df$study_id == "WU0019"] <- c("NGT", "NGT")
df$Dysglycemia[df$study_id == "IA0080"] <- "AGT"
df$CFRD[df$study_id == "WU0019"] <- c("CFRD-", "CFRD-")
df$CFRD[df$study_id == "IA0080"] <- "CFRD-"
# Convert CV units to match Scully et al.
df$cv <- df$cv * 100
# First CGM and OGTT pair only
# df <- df %>%
#     group_by(study_id) %>%
#     slice_min(Date)
```

For the following analyses, we used ROC curves to assess various CGM metrics' ability to distinguish between CFRD- and CFRD+ people with CF. Because some people had multiple CGM wears and OGTTs, we decided to use each participant's first CGM and OGTT pair. In the cleaned EnVision dataset we have a total of `r nrow(df)` CGMs from `r length(unique(df$study_id))` unique individuals. Of these CGM wears, `r sum(is.na(df$Diagnosis))` were missing a diagnosis due to missing OGTT timepoints. Those without a diagnosis were excluded from all analyses.

## Participant Demographics by Diagnosis

```{r results='asis'}
#| tbl-cap: "Unique Participants"
t1 <- tableby(
  Dysglycemia ~
    fe(redcap_data_access_group) +
    age_visit +
    height +
    weight +
    bmi_perc +
    sex +
    fe(origin_race) +
    fe(ethnicity) +
    fe(pancreatic_status) +
    fe(CFTR) +
    corrector_yes_no +
    fev1 +
    fvc +
    a1c_result,
  data = df |> group_by(study_id) |> slice_min(Date)
)
summary(t1, pfootnote = T)
```

Pairwise HbA1c comparisons:

```{r}
df %>%
  group_by(study_id) %>%
  slice_min(Date) %>%
  aov(a1c_result ~ Dysglycemia, data = .) %>%
  TukeyHSD()
```

```{r results='asis'}
#| tbl-cap: "CGM Wears"
t2 <- tableby(
  Dysglycemia ~
    fe(redcap_data_access_group) +
    age_visit +
    height +
    weight +
    bmi_perc +
    sex +
    fe(origin_race) +
    fe(ethnicity) +
    fe(pancreatic_status) +
    fe(CFTR) +
    corrector_yes_no +
    fev1 +
    fvc +
    a1c_result,
  data = df
)
summary(t2, pfootnote = T)
```

Pairwise HbA1c comparisons:

```{r}
df %>%
  aov(a1c_result ~ Dysglycemia, data = .) %>%
  TukeyHSD()
```

## Demographics by CFRD status

```{r results='asis'}
#| tbl-cap: "Unique Participants"
t1 <- tableby(
  CFRD ~
    fe(redcap_data_access_group) +
    age_visit +
    height +
    weight +
    bmi_perc +
    sex +
    fe(origin_race) +
    fe(ethnicity) +
    fe(pancreatic_status) +
    fe(CFTR) +
    corrector_yes_no +
    fev1 +
    fvc +
    a1c_result,
  data = df |> group_by(study_id) |> slice_min(Date)
)
summary(t1, pfootnote = T)
```


```{r results='asis'}
#| tbl-cap: "CGM Wears"
t2 <- tableby(
  CFRD ~
    fe(redcap_data_access_group) +
    age_visit +
    height +
    weight +
    bmi_perc +
    sex +
    fe(origin_race) +
    fe(ethnicity) +
    fe(pancreatic_status) +
    fe(CFTR) +
    corrector_yes_no +
    fev1 +
    fvc +
    a1c_result,
  data = df
)
summary(t2, pfootnote = T)
```

## CGM Metrics by Diagnosis

```{r results='asis'}
#| tbl-cap: "Unique Participants"
t1 <- tableby(
  Dysglycemia ~
    Glucose_0 +
    Glucose_60 +
    Glucose_120 +
    mean_sensor +
    gmi +
    kwt(percent_time_over_140, "Nmiss", "medianq1q3") +
    kwt(percent_time_over_180, "Nmiss", "medianq1q3") +
    kwt(percent_time_over_250, "Nmiss", "medianq1q3") +
    kwt(percent_time_under_70, "Nmiss", "medianq1q3") +
    kwt(percent_time_under_60, "Nmiss", "medianq1q3") +
    kwt(standard_deviation, "Nmiss", "medianq1q3") +
    kwt(cv, "Nmiss", "medianq1q3") +
    kwt(r_mage, "Nmiss", "medianq1q3"),
  data = df |> group_by(study_id) |> slice_min(Date)
)
summary(t1, pfootnote = T)
```

```{r results='asis'}
#| tbl-cap: "CGM Wears"
t1 <- tableby(
  Dysglycemia ~
    Glucose_0 +
    Glucose_60 +
    Glucose_120 +
    mean_sensor +
    gmi +
    kwt(percent_time_over_140, "Nmiss", "medianq1q3") +
    kwt(percent_time_over_180, "Nmiss", "medianq1q3") +
    kwt(percent_time_over_250, "Nmiss", "medianq1q3") +
    kwt(percent_time_under_70, "Nmiss", "medianq1q3") +
    kwt(percent_time_under_60, "Nmiss", "medianq1q3") +
    kwt(standard_deviation, "Nmiss", "medianq1q3") +
    kwt(r_mage, "Nmiss", "medianq1q3"),
  data = df
)
summary(t1, pfootnote = T)
```

## Mean Sensor Glucose 

```{r}
r <- roc(CFRD ~ mean_sensor, data = df, quiet = T)
rmg = r
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
mg_best = coords(r, "best", ret = "all")
gt(mg_best) %>% fmt_number(decimals = 3)
mg_coords = data.frame(coords(
  r,
  "all",
  ret = c(c(
    "thresh",
    "sensitivity",
    "specificity",
    "fp",
    "tp",
    "fn",
    "tn",
    "npv",
    "ppv"
  ))
))
gt(mg_coords) %>%
  fmt_number(decimals = 3)
# Get thresholds for plotting
mg_high_sens = max(mg_coords$threshold[
  which(mg_coords$sensitivity >= 0.9)
])
mg_high_spec = min(mg_coords$threshold[
  which(mg_coords$specificity >= 0.9)
])
mg_youden = mg_best$threshold
```

## Percent Time >140 mg/dL 

```{r}
r <- roc(CFRD ~ percent_time_over_140, data = df, quiet = T)
rtar140 = r
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
tar140_best = coords(r, "best", ret = "all")
gt(tar140_best) %>% fmt_number(decimals = 3)
tar140_coords = data.frame(coords(
  r,
  "all",
  ret = c(c(
    "thresh",
    "sensitivity",
    "specificity",
    "fp",
    "tp",
    "fn",
    "tn",
    "npv",
    "ppv"
  ))
))
gt(tar140_coords) %>%
  fmt_number(decimals = 3)
# Get thresholds for plotting
tar140_high_sens = max(tar140_coords$threshold[
  which(tar140_coords$sensitivity >= 0.9)
])
tar140_high_spec = min(tar140_coords$threshold[
  which(tar140_coords$specificity >= 0.9)
])
tar140_youden = tar140_best$threshold
```

## Percent Time >180 mg/dL

```{r}
r <- roc(CFRD ~ percent_time_over_180, data = df, quiet = T)
rtar180 = r
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
tar180_best = coords(r, "best", ret = "all")
gt(tar180_best) %>% fmt_number(decimals = 3)
tar180_coords = data.frame(coords(
  r,
  "all",
  ret = c(c(
    "thresh",
    "sensitivity",
    "specificity",
    "fp",
    "tp",
    "fn",
    "tn",
    "npv",
    "ppv"
  ))
))
gt(tar180_coords) %>%
  fmt_number(decimals = 3)
# Get thresholds for plotting
tar180_high_sens = max(tar180_coords$threshold[which(
  tar180_coords$sensitivity >= 0.9
)])
tar180_high_spec = min(tar180_coords$threshold[
  which(tar180_coords$specificity >= 0.9)
])
tar180_youden = tar180_best$threshold
```

## GMI

```{r}
r <- roc(CFRD ~ gmi, data = df, quiet = T)
rgmi = r
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
gmi_best = coords(r, "best", ret = "all")
gt(gmi_best) %>% fmt_number(decimals = 3)
gmi_coords = data.frame(coords(
  r,
  "all",
  ret = c(c(
    "thresh",
    "sensitivity",
    "specificity",
    "fp",
    "tp",
    "fn",
    "tn",
    "npv",
    "ppv"
  ))
))
gt(gmi_coords) %>%
  fmt_number(decimals = 3)
# Get thresholds for plotting
gmi_high_sens = max(gmi_coords$threshold[
  which(gmi_coords$sensitivity >= 0.9)
])
gmi_high_spec = min(gmi_coords$threshold[
  which(gmi_coords$specificity >= 0.9)
])
gmi_youden = gmi_best$threshold
```

## SD

```{r}
r <- roc(CFRD ~ standard_deviation, data = df, quiet = T)
rsd = r
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
sd_best = coords(r, "best", ret = "all")
gt(sd_best) %>% fmt_number(decimals = 3)
sd_coords = data.frame(coords(
  r,
  "all",
  ret = c(c(
    "thresh",
    "sensitivity",
    "specificity",
    "fp",
    "tp",
    "fn",
    "tn",
    "npv",
    "ppv"
  ))
))
gt(sd_coords) %>%
  fmt_number(decimals = 3)
# Get thresholds for plotting
sd_high_sens = max(sd_coords$threshold[
  which(sd_coords$sensitivity >= 0.9)
])
sd_high_spec = min(sd_coords$threshold[
  which(sd_coords$specificity >= 0.9)
])
sd_youden = sd_best$threshold
```

## CV

```{r}
r <- roc(CFRD ~ cv, data = df, quiet = T)
rcv = r
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
cv_best = coords(r, "best", ret = "all")
gt(cv_best) %>% fmt_number(decimals = 3)
cv_coords = data.frame(coords(
  r,
  "all",
  ret = c(c(
    "thresh",
    "sensitivity",
    "specificity",
    "fp",
    "tp",
    "fn",
    "tn",
    "npv",
    "ppv"
  ))
))
gt(cv_coords) %>%
  fmt_number(decimals = 3)
# Get thresholds for plotting
cv_high_sens = max(cv_coords$threshold[
  which(cv_coords$sensitivity >= 0.9)
])
cv_high_spec = min(cv_coords$threshold[
  which(cv_coords$specificity >= 0.9)
])
cv_youden = cv_best$threshold
```

## HbA1c

```{r}
r <- roc(CFRD ~ a1c_result, data = df, quiet = T)
ra1c = r
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
a1c_best = coords(r, "best", ret = "all")
gt(a1c_best) %>% fmt_number(decimals = 3)
a1c_coords = data.frame(coords(
  r,
  "all",
  ret = c(c(
    "thresh",
    "sensitivity",
    "specificity",
    "fp",
    "tp",
    "fn",
    "tn",
    "npv",
    "ppv"
  ))
))
gt(a1c_coords) %>%
  fmt_number(decimals = 3)
# Get thresholds for plotting
a1c_high_sens = max(a1c_coords$threshold[
  which(a1c_coords$sensitivity >= 0.9)
])
a1c_high_spec = min(a1c_coords$threshold[
  which(a1c_coords$specificity >= 0.9)
])
a1c_youden = a1c_best$threshold
```

## ROC Comparisons

```{r}
cat("### MG vs. CV")
roc.test(rmg, rcv)
cat("### MG vs. SD")
roc.test(rmg, rsd)
cat("### TAR140 vs. CV")
roc.test(rtar140, rcv)
cat("### TAR140 vs. SD")
roc.test(rtar140, rsd)
cat("### TAR180 vs. CV")
roc.test(rtar180, rcv)
cat("### TAR180 vs. SD")
roc.test(rtar180, rsd)
```

## Pulmonary Function Comparisons by CGM

```{r}
df$scully_ag <- cut(df$mean_sensor, c(-Inf, 112, Inf))
df$our_ag <- cut(df$mean_sensor, c(-Inf, 101, Inf))
df$scully_140 <- cut(df$percent_time_over_140, c(-Inf, 17.5, Inf))
df$our_140 <- cut(df$percent_time_over_140, c(-Inf, 11.3, Inf))
df$scully_180 <- cut(df$percent_time_over_180, c(-Inf, 3.4, Inf))
df$our_180 <- cut(df$percent_time_over_180, c(-Inf, 3.1, Inf))
df$scully_sd <- cut(df$standard_deviation, c(-Inf, 30.8, Inf))
df$our_sd <- cut(df$standard_deviation, c(-Inf, 28.8, Inf))
df$scully_cv <- cut(df$cv, c(-Inf, 30.6, Inf))
df$our_cv <- cut(df$cv, c(-Inf, 27.7, Inf))
```

### Mean glucose

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_ag ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_ag ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### Percent time spent >140 mg/dL 

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_140 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_140 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### Percent time spent >180 mg/dL 

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_180 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_180 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### SD 

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_sd ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_sd ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### CV

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_cv ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_cv ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

## Figure

In the figure below, red lines indicate the cutoffs identified in Scully et al. and purple lines indicate the cutoffs identified in EnVision.

```{r}
#| message: false
#| warning: false
#| fig-height: 16
#| fig-width: 12
# Dotplot
box_tar140 <- ggplot(
  df,
  aes(
    x = Dysglycemia,
    y = percent_time_over_140,
    text = study_id
  )
) +
  geom_jitter(alpha = 0.2) +
  geom_hline(aes(yintercept = tar140_youden, color = "Youden")) +
  geom_hline(aes(yintercept = tar140_high_sens, color = "High Sensitivity")) +
  geom_hline(aes(yintercept = tar140_high_spec, color = "High Specificity")) +
  ggtitle("% Time Over 140 (mg/dL)") +
  theme_bw(base_size = 25) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = rel(1))
  ) +
  scale_color_manual(
    name = "Threshold",
    values = c("#32936f", "#2274a5", "#ffbf00")
  )
box_tar180 <- ggplot(
  df,
  aes(
    x = Dysglycemia,
    y = percent_time_over_180,
    text = study_id
  )
) +
  geom_jitter(alpha = 0.2) +
  geom_hline(aes(yintercept = tar180_youden, color = "Youden")) +
  geom_hline(aes(yintercept = tar180_high_sens, color = "High Sensitivity")) +
  geom_hline(aes(yintercept = tar180_high_spec, color = "High Specificity")) +
  ggtitle("% Time Over 180 (mg/dL)") +
  theme_bw(base_size = 25) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = rel(1))
  ) +
  scale_color_manual(
    name = "Threshold",
    values = c("#32936f", "#2274a5", "#ffbf00")
  )
box_mg <- ggplot(
  df,
  aes(
    x = Dysglycemia,
    y = mean_sensor,
    text = study_id
  )
) +
  geom_jitter(alpha = 0.2) +
  geom_hline(aes(yintercept = mg_youden, color = "Youden")) +
  geom_hline(aes(yintercept = mg_high_sens, color = "High Sensitivity")) +
  geom_hline(aes(yintercept = mg_high_spec, color = "High Specificity")) +
  ggtitle("Mean glucose (mg/dL)") +
  theme_bw(base_size = 25) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = rel(1))
  ) +
  scale_color_manual(
    name = "Threshold",
    values = c("#32936f", "#2274a5", "#ffbf00")
  )
box_cv <- ggplot(
  df,
  aes(
    x = Dysglycemia,
    y = cv,
    text = study_id
  )
) +
  geom_jitter(alpha = 0.2) +
  geom_hline(aes(yintercept = cv_youden, color = "Youden")) +
  geom_hline(aes(yintercept = cv_high_sens, color = "High Sensitivity")) +
  geom_hline(aes(yintercept = cv_high_spec, color = "High Specificity")) +
  ggtitle("Coefficient of Variation (%)") +
  theme_bw(base_size = 25) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = rel(1))
  ) +
  scale_color_manual(
    name = "Threshold",
    values = c("#32936f", "#2274a5", "#ffbf00")
  )
box_hba1c <- ggplot(
  df,
  aes(
    x = Dysglycemia,
    y = a1c_result,
    text = study_id
  )
) +
  geom_jitter(alpha = 0.2) +
  geom_hline(aes(yintercept = a1c_youden, color = "Youden")) +
  geom_hline(aes(yintercept = a1c_high_sens, color = "High Sensitivity")) +
  geom_hline(aes(yintercept = a1c_high_spec, color = "High Specificity")) +
  ggtitle("HbA1c (%)") +
  theme_bw(base_size = 25) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = rel(1))
  ) +
  scale_color_manual(
    name = "Threshold",
    values = c("#32936f", "#2274a5", "#ffbf00")
  )
box_sd <- ggplot(
  df,
  aes(
    x = Dysglycemia,
    y = standard_deviation,
    text = study_id
  )
) +
  geom_jitter(alpha = 0.2) +
  geom_hline(aes(yintercept = sd_youden, color = "Youden")) +
  geom_hline(aes(yintercept = sd_high_sens, color = "High Sensitivity")) +
  geom_hline(aes(yintercept = sd_high_spec, color = "High Specificity")) +
  ggtitle("SD glucose (mg/dL)") +
  theme_bw(base_size = 25) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = rel(1))
  ) +
  scale_color_manual(
    name = "Threshold",
    values = c("#32936f", "#2274a5", "#ffbf00")
  )
box_p <- (box_mg + box_hba1c) /
  (box_tar140 + box_tar180) /
  (box_cv + box_sd) +
  plot_layout(guides = "collect")
box_p
ggsave(
  "./Christine Chan/EnVision CF/Dissemination/ROC Curves/Figures/dotplot_all.png",
  height = 16,
  width = 12,
  units = "in"
)
```

# Hypoglycemia Comparisons

```{r}
# Timepoint and variables lists
timepoints <- c(0, 10, 30, 60, 90, 120, 150, 180)
limited_timepoints <- c(0, 120, 150, 180)
glucoses <- paste0("Glucose_", timepoints)
insulins <- paste0("Insulin_", timepoints)
cpeptides <- paste0("C.Peptide_", timepoints)
glucagons <- paste0("Glucagon_", timepoints)
glp1s <- paste0("GLP.1.Active_", timepoints)
gips <- paste0("GIP_", timepoints)
epinephrines <- paste0("Epinephrine_", timepoints)
norepinephrines <- paste0("Norepinephrine_", timepoints)
adren_survey <- paste0("adren_score_", limited_timepoints)
neuro_survey <- paste0("neuro_score_", limited_timepoints)
total_survey <- paste0("total_score_", limited_timepoints)
num_symptoms_survey <- paste0("num_symptoms_", limited_timepoints)
hypo_surveys <- c(adren_survey, neuro_survey, total_survey, num_symptoms_survey)
# First visit dataframe
first_visit_df <- final_df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup()
# Long data for mixed models
glmm_df <- final_df %>%
  select(
    study_id,
    Hypo60,
    Hypo70,
    ogtt_num,
    all_of(c(
      glucoses,
      insulins,
      cpeptides,
      glucagons,
      glp1s,
      gips,
      epinephrines,
      norepinephrines,
      hypo_surveys
    ))
  ) %>%
  pivot_longer(
    all_of(c(
      glucoses,
      insulins,
      cpeptides,
      glucagons,
      glp1s,
      gips,
      epinephrines,
      norepinephrines,
      hypo_surveys
    )),
    names_to = c(".value", "Timepoint"),
    names_pattern = "(.*)_([0-9]{1,3})"
  )
glmm_df$Timepoint <- factor(glmm_df$Timepoint, levels = timepoints)
```

## Demographics at First Visit

### By Hypoglycemia <70 mg/dL at Any Point

```{r results='asis'}
t1 <- tableby(
  Hypo70 ~
    fe(Diagnosis) +
    redcap_data_access_group +
    kwt(age_visit, "Nmiss", "medianq1q3", "range") +
    kwt(height, "Nmiss", "medianq1q3", "range") +
    kwt(weight, "Nmiss", "medianq1q3", "range") +
    bmi_perc +
    sex +
    fe(origin_race) +
    fe(ethnicity) +
    fe(pancreatic_status) +
    fe(CFTR) +
    corrector_yes_no +
    fev1 +
    fvc +
    a1c_result,
  data = first_visit_df,
  control = tableby.control(simulate.p.value = T)
)
summary(t1, pfootnote = T)
```

### By Hypoglycemia <60 mg/dL at Any Point

```{r results='asis'}
t2 <- tableby(
  Hypo60 ~
    fe(Diagnosis) +
    redcap_data_access_group +
    kwt(age_visit, "Nmiss", "medianq1q3", "range") +
    kwt(height, "Nmiss", "medianq1q3", "range") +
    kwt(weight, "Nmiss", "medianq1q3", "range") +
    bmi_perc +
    sex +
    fe(origin_race) +
    fe(ethnicity) +
    fe(pancreatic_status) +
    fe(CFTR) +
    corrector_yes_no +
    fev1 +
    fvc +
    a1c_result,
  data = first_visit_df,
  control = tableby.control(simulate.p.value = T)
)
summary(t2, pfootnote = T)
```

## OGTT Measures by Timepoint

Glucose, insulin, C-peptide, glucagon, epinephrine, and norepinephrine were modeled using a linear mixed effect model with random effects for OGTT within participant and time treated as a categorical variable. The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.

### Glucose

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  Glucose ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Glucose") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean glucose at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour glucose was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucose was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  Glucose ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Glucose") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean glucose at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour glucose was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucose was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### Insulin

Insulin values were log-transformed to improve model fit.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(Insulin) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Insulin") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean insulin at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour insulin was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour insulin was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(Insulin) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Insulin") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean insulin at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour insulin was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour insulin was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### C-peptide

C-peptide values were log-transformed to improve model fit.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(C.Peptide) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean C-peptide") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean C-peptide at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour C-peptide was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour C-peptide was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(C.Peptide) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean C-peptide") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean C-peptide at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour C-peptide was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour C-peptide was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### Glucagon

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  Glucagon ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Glucagon") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean glucagon at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour glucagon was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucagon was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  Glucagon ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Glucagon") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean glucagon at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour glucagon was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucagon was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### GIP

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  GIP ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean GIP") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean GIP at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  GIP ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean GIP") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean GIP at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### Epinephrine

Epinephrine values were log-transformed to improve model fit.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Epinephrine") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean epinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Epinephrine") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean epinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### Norepinephrine

Norepinephrine values were log-transformed to improve model fit.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Norepinephrine") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(
  log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean Norepinephrine") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### `adren_score`

The `adren_score` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  adren_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean adren_score") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  adren_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean adren_score") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### `neuro_score`

The `neuro_score` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  neuro_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean neuro_score") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  neuro_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean neuro_score") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### `total_score`

The `total_score` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  total_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean total_score") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  total_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean total_score") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### `num_symptoms`

The `num_symptoms` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  num_symptoms ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo70"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo70",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo70,
    group = Hypo70
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean num_symptoms") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
# Poisson model better than NB by AIC and BIC
mod <- glmer(
  num_symptoms ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
  data = glmm_df,
  family = "poisson",
  control = glmerControl(optimizer = "bobyqa")
)
# EM means
means <- data.frame(emmeans(
  mod,
  c("Timepoint", "Hypo60"),
  type = "response",
  pbkrtest.limit = 1e6
))
colnames(means) <- c(
  "Timepoint",
  "Hypo60",
  "Mean",
  "SE",
  "df",
  "CI Lower Limit",
  "CI Upper Limit"
)
# Plot
ggplot(
  means,
  aes(
    x = as.numeric(as.character(Timepoint)),
    y = Mean,
    colour = Hypo60,
    group = Hypo60
  )
) +
  geom_point() +
  geom_errorbar(
    aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
    width = 3
  ) +
  geom_line() +
  scale_x_continuous(breaks = timepoints) +
  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
  ylab("EM Mean num_symptoms") +
  xlab("Timepoint") +
  theme_bw(base_size = 14) +
  theme(legend.title = element_blank())
```

The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.

```{r}
means %>%
  gt() %>%
  fmt_number(decimals = 2)
```

The table above shows the estimated mean norepinephrine at each time point for the two groups. For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

## OGTT Summary Measures

OGTT summary measures (i.e., measures calculated using multiple OGTT timepoints) were modeled using a linear mixed effect model with random intercept for participant.

### Matsuda index

The Matsuda index was calculated across the 0, 30, 60, 90, and 120 minute timepoints using the formula: $$\frac{10000}{\sqrt{(\frac{\text{fasting glucose (mg/dL)}}{18} * \text{fasting insulin}) * (\frac{\text{mean glucose (mg/dL)}}{18} * \text{mean insulin})}}$$

Matsuda index values were log-transformed to improve model fit.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(log(matsuda) ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(log(matsuda) ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### HOMA-IR

HOMA-IR was calculated using the formula: $$\frac{\text{fasting glucose (mg/dL)}*\text{fasting insulin}}{405}$$

HOMA-IR values were log-transformed to improve model fit.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(log(homa_ir) ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the HOMA-IR was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(homa_ir ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the HOMA-IR was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 30 Minute Glucose

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC30gluc ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 30 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC30gluc ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 30 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 60 Minute Glucose

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC60gluc ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 60 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC60gluc ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 60 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 120 Minute Glucose

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC120gluc ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 120 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC120gluc ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 120 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 180 Minute Glucose

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC180gluc ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 180 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC180gluc ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 180 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 30 Minute Insulin

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC30ins ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 30 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC30ins ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 30 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 60 Minute Insulin

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC60ins ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 60 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC60ins ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 60 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 120 Minute Insulin

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC120ins ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 120 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC120ins ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 120 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### iAUC 180 Minute Insulin

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC180ins ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 180 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(iAUC180ins ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the iAUC 180 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### ODI

ODI was calculated as $\frac{\text{iAUC 30 Minute Insulin}}{\text{iAUC 30 Minute Glucose}}$.

#### Hypoglycemia <70 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(odi ~ Hypo70 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the ODI was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

#### Hypoglycemia <60 mg/dL at Any Point

```{r}
# Fit
mod <- lmer(odi ~ Hypo60 + (1 | study_id), data = final_df)
mod %>% gtsummary::tbl_regression()
```

The table above shows that the ODI was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
