---
title: "Title"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gtsummary)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import
df <- read.csv("/Users/timvigers/Desktop/EnVision/EnvisionCF_DATA_2024-03-14_1325.csv",
  na.strings = ""
)
# Fix missing access group
df$redcap_data_access_group[is.na(df$redcap_data_access_group)] <- "iowa"
# Trim columns
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
df <- df %>%
  select(
    study_id:redcap_data_access_group, age_at_current_date:origin_race,
    cftr_mutation_1, cftr_mutation_2, bmi_study_visit, bmi_stand_score,
    a1c_result, fev1, all_of(glucose)
  )
# Get diagnosis
df$Diagnosis <- NA
df$Diagnosis[df[, glucose[1]] < 100 &
  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) == rowSums(!is.na(df[, glucose[2:5]]))) &
  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) == rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
df$Diagnosis[df[, glucose[1]] >= 100] <- "IFG"
df$Diagnosis[rowSums(df[, glucose[2:5]] >= 200, na.rm = T) > 0] <- "INDET"
df$Diagnosis[rowSums(df[, glucose[6:8]] >= 140, na.rm = T) > 0] <- "IGF"
df$Diagnosis[df$timepoint_0_min >= 126 | df$timepoint_120_min >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
# Exclude those missing <= 3 missing OGTT values
df$Keep <- rowSums(is.na(df[, glucose])) <= 3
df <- df %>% filter(Keep)
# Write for checking
write.csv(df, file = "~/df.csv", row.names = F, na = "")
```

# Demographics at first visit

`r length(unique(df$study_id))` participants were included because they were missing <= 3 OGTT timepoints.

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(redcap_data_access_group:cftr_mutation_2) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

# Diagnoses

## All visits

```{r}
#| label: tbl-2
#| tbl-cap: Participant diagnoses based on OGTT
df %>%
  select(redcap_data_access_group, Diagnosis) %>%
  tbl_summary(by = redcap_data_access_group) %>%
  add_overall()
```

## Visit 1 only

```{r}
#| label: tbl-3
#| tbl-cap: Participant diagnoses based on OGTT (visit 1 only)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(redcap_data_access_group, Diagnosis) %>%
  tbl_summary(by = redcap_data_access_group) %>%
  add_overall()
```

# Figure

```{r}
#| label: fig-1
#| fig-cap: Diagnoses over time
#| fig-width: 12
#| fig-height: 8
df %>%
  group_by(study_id) %>%
  mutate(`Visit Num.` = row_number()) %>%
  group_by(`Visit Num.`) %>%
  count(Diagnosis) %>%
  mutate(total = sum(n), perc = n / total) %>%
  ggplot(., aes(x = `Visit Num.`, y = perc, fill = Diagnosis)) +
  geom_bar(position = "stack", stat = "identity") +
  geom_text(aes(label = paste0(scales::percent(perc), " (n = ", n, ")")),
    position = position_stack(vjust = 0.5)
  ) +
  theme_bw() +
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```
