---
title: "CF Registry Analysis - Lung Function"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    embed-resources: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(ggplot2)
library(gtsummary)
library(gt)
library(pROC)
library(ggrepel)
library(childsds)
library(labelled)
library(Hmisc)
library(nlme)
library(emmeans)
library(multcomp)
library(tidyverse)
library(parallel)
library(knitr)
base_dir <- switch(
    Sys.info()["nodename"],
    "togo" = "/home/tvigers/Documents/Data",
    "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
    "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
    "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers"
)
data_dir <- switch(
    Sys.info()["nodename"],
    "togo" = "/CF/Christine Chan/CF Registry",
    "Tims-MacBook-Air.local" = "/CF/Christine Chan/CF Registry",
    "Tims-Mac-mini.local" = "/CF/Christine Chan/CF Registry",
    "Mac" = "/CF/Christine Chan/CF Registry"
)
home_dir <- paste0(base_dir, data_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
#| warning: false
#| message: false
# Import data and codebook
setwd(home_dir)
annual <- read.csv(
    "./Data_Raw/CFF19_Annualized_Del1.csv",
    na.strings = c("", "N/A")
)
# Remove those with T1D or T2D
annual <- annual |> arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(
    annual$cfrd_status_annual,
    levels = 1:3,
    labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
races <- c(
    "White",
    "Black or African American",
    "American Indian or Alaska Native",
    "Asian",
    "Native Hawaiian or Other Pacific ",
    "Some other race"
)
demogs$Race <- apply(demogs, 1, function(r) {
    w <- which(r[paste0("Race", 1:6)] == 1)
    if (length(w) > 1) {
        return("More than one race")
    } else {
        return(races[w])
    }
})
demogs$Race <- factor(
    demogs$Race,
    levels = c(
        "White",
        "American Indian or Alaska Native",
        "Asian",
        "Black or African American",
        "More than one race",
        "Native Hawaiian or Other Pacific ",
        "Some other race"
    ),
    labels = c(
        "White",
        "Other race",
        "Other race",
        "Black or African American",
        "More than one race",
        "Other race",
        "Other race"
    )
)
demogs$Hispanicrace <- factor(
    demogs$Hispanicrace,
    levels = 1:2,
    labels = c("Yes", "No")
)
demogs <- demogs |> select(eDWID, Gender, MutClass, Race, Hispanicrace)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
    apply(annual[, liver_fields], 1, function(r) {
        if (all(is.na(r))) {
            return(NA)
        } else {
            return(max(as.numeric(r), na.rm = T))
        }
    })
# Get BMI categories
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
annual$bmi_perc_cdc <- sds(
    annual$bmi,
    age = annual$Age_YrEnd,
    sex = annual$Gender,
    male = "M",
    female = "F",
    item = "bmi",
    ref = cdc.ref,
    type = "perc"
) *
    100
annual$bmi_perc_who <- sds(
    annual$bmi,
    age = annual$Age_YrEnd,
    sex = annual$Gender,
    male = "M",
    female = "F",
    item = "bmi",
    ref = who.ref,
    type = "perc"
) *
    100
annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
annual$bmi_cat <- apply(annual, 1, function(r) {
    age <- as.numeric(r["Age_YrEnd"])
    gender <- as.character(r["Gender"])
    bmi <- as.numeric(r["bmi"])
    bmi_perc <- as.numeric(r["bmi_perc"])
    if (age <= 20) {
        return(cut(
            bmi_perc,
            c(-Inf, 10, 50, 85, Inf),
            right = F,
            labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
        ))
    } else if (age > 20) {
        if (gender == "M") {
            return(cut(
                bmi,
                c(-Inf, 18.5, 23, 25, Inf),
                right = F,
                labels = c(
                    "Undernourished",
                    "At Risk",
                    "Target/Goal",
                    "Overweight"
                )
            ))
        } else if (gender == "F") {
            return(cut(
                bmi,
                c(-Inf, 18.5, 22, 25, Inf),
                right = F,
                labels = c(
                    "Undernourished",
                    "At Risk",
                    "Target/Goal",
                    "Overweight"
                )
            ))
        }
    }
})
annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
# Raw values for adults, percentiles for peds
annual$bmi_adult <- annual$bmi
annual$bmi_adult[annual$Age_YrEnd <= 20] <- NA
annual$bmi_perc <- annual$bmi_perc
annual$bmi_perc[annual$Age_YrEnd > 20] <- NA
# Labels
var_label(annual$bmi) <- "BMI (all participants)"
var_label(annual$bmi_perc) <- "BMI percentile (age <= 20)"
var_label(annual$bmi_adult) <- "BMI value (age > 20)"
var_label(annual$bmi_cat) <- "BMI category"
# Format columns
annual$MutClass <- factor(
    annual$MutClass,
    levels = c("1-3", "4-5", "Oth", "N/A"),
    labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(
    annual$NumPulmExacerbation,
    labels = c("0", "1", "2", ">=3"),
    breaks = c(0, 1, 2, 3, Inf),
    right = F
)
# Bugs
bugs <- c(
    "A_pseudomonasaeruginosa",
    "A_fungalyeast1",
    "A_pulmonarycomplications1",
    "A_mycocultureresults3",
    "A_cultureresults3"
)
# Yes/No variables
annual$cfrd_treatment_options1[annual$cfrd_treatment_options1 == 2] <- NA
annual$cfrd_treatment_options4[annual$cfrd_treatment_options4 == 4] <- NA
yn_vars <- c(
    "A_corticosteroids1",
    "A_feeding2",
    "A_Vx770",
    "A_VX809comb",
    "liver_disease",
    bugs,
    "cfrd_treatment_options2",
    "cfrd_treatment_options3",
    "cfrd_treatment_options4"
)
annual[, yn_vars] <- lapply(
    annual[, yn_vars],
    factor,
    levels = 0:1,
    labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- factor(
    annual$A_IsOnEnzymes,
    levels = c(1, 0),
    labels = c("Yes", " No")
)
# Males to reference group
annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
# Hispanic no as reference
annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
# Get first year of CFRD
first_cfrd <- annual |>
    filter(cfrd_status_annual == "CFRD+") |>
    group_by(eDWID) |>
    summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
# years following are CFRD+
annual$CFRD <- NA
annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
# Check whether they had an OGTT and/or HbA1c the year of first CFRD or the year
# prior. This is super slow but idc idc.
cl <- makeCluster(type = "FORK", 8)
cfrd_only <- parApply(
    cl,
    first_cfrd,
    1,
    function(r) {
        id <- as.numeric(r["eDWID"])
        year <- as.numeric(r["first_cfrd"])
        d <- annual[
            annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1),
        ]
        a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
        ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
        ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
        d <- annual[
            annual$eDWID == id & annual$ReviewYear %in% c(year, year + 1),
        ]
        treatment <- any(
            d[, c(
                "cfrd_treatment_options2",
                "cfrd_treatment_options3",
                "cfrd_treatment_options4"
            )] ==
                "Yes",
            na.rm = TRUE
        )
        return(c(id, year, a1c, ogtt0hr, ogtt2hr, treatment))
    }
)
stopCluster(cl)
cfrd_only <- data.frame(t(cfrd_only))
colnames(cfrd_only) <- c(
    "eDWID",
    "dx_year",
    "a1c",
    "ogtt0hr",
    "ogtt2hr",
    "treatment"
)
# cfrd_only_store <- cfrd_only
cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
    not_missing <- paste(
        which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
        collapse = "_"
    )
    return(not_missing)
})
cfrd_only$a1c_ogtt <- factor(
    cfrd_only$a1c_ogtt,
    levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
    labels = c(
        "Neither",
        "HbA1c Only",
        "OGTT",
        "OGTT",
        "OGTT",
        "OGTT",
        "OGTT",
        "OGTT"
    )
)
var_label(cfrd_only$a1c_ogtt) <- "Diagnosis Method"
cfrd_only$treatment <- factor(
    cfrd_only$treatment,
    levels = 0:1,
    labels = c("No Treatment", "Treatment")
)
var_label(cfrd_only$treatment) <- "CFRD Treatment Started Within 1 Year?"
```

# Diagnosed with CFRD between 2010 - 2018

## Participant characteristics at diagnosis

```{r}
#| warning: false
#| message: false
# Create a model dataframe
mod_df <- inner_join(
    annual,
    cfrd_only |> filter(dx_year >= 2010),
    by = join_by(eDWID)
) |>
    drop_na(a1c_ogtt) |>
    mutate(years_to_cfrd = ReviewYear - first_cfrd) |>
    filter(years_to_cfrd <= 5 & years_to_cfrd >= -5) |>
    mutate(years_to_cfrd = factor(years_to_cfrd))
var_label(mod_df$years_to_cfrd) <- "Years From CFRD Dx"
# Write data for Christine to look at
mod_df |>
    select(
        eDWID,
        Race,
        Hispanicrace,
        MutClass,
        ReviewYear,
        Age_YrEnd,
        CFRD,
        first_cfrd,
        age_dx,
        A_weight,
        A_height,
        bmi_adult,
        bmi_perc,
        bmi_cat,
        A_FEV1_pct_predicted,
        A_FVC_pct_predicted,
        a1c_ogtt,
        A_hgba1c,
        A_ogttfast_bloodglucose,
        A_twohour_bloodglucose,
        A_IsOnEnzymes,
        liver_disease,
        A_pseudomonasaeruginosa,
        A_fungalyeast1,
        A_mycocultureresults3,
        A_cultureresults3,
        treatment,
        cfrd_treatment_options2,
        cfrd_treatment_options3,
        cfrd_treatment_options4,
        NumFEV1,
        NumFVC
    ) %>%
    write.csv(
        .,
        file = "./Data_Clean/fev1_model_data_2010.csv",
        row.names = FALSE,
        na = ""
    )
# Table 1 variable list
t1_vars <- c(
    "eDWID",
    "Race",
    "Hispanicrace",
    "MutClass",
    "age_dx",
    "A_weight",
    "A_height",
    "bmi_adult",
    "bmi_perc",
    "bmi_cat",
    "A_FEV1_pct_predicted",
    "A_FVC_pct_predicted",
    "a1c_ogtt",
    "A_hgba1c",
    "A_ogttfast_bloodglucose",
    "A_twohour_bloodglucose",
    "A_IsOnEnzymes",
    "liver_disease",
    "A_pseudomonasaeruginosa",
    "A_fungalyeast1",
    "A_mycocultureresults3",
    "A_cultureresults3",
    "treatment",
    "cfrd_treatment_options2",
    "cfrd_treatment_options3",
    "cfrd_treatment_options4",
    "NumFEV1",
    "NumFVC"
)
```

### All groups

```{r}
#| warning: false
#| message: false
mod_df |>
    filter(ReviewYear == first_cfrd) |>
    select(all_of(t1_vars)) |>
    tbl_summary(
        by = a1c_ogtt,
        missing_text = "Missing",
        statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) |>
    add_overall() |>
    add_p() |>
    modify_caption("**All Groups**")
```

### Pairwise comparisons

P values FDR-adjusted for multiple comparisons.

```{r}
#| warning: false
#| message: false
mod_df |>
    filter(ReviewYear == first_cfrd) |>
    filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
    mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
    select(all_of(t1_vars)) |>
    tbl_summary(
        by = a1c_ogtt,
        missing_text = "Missing",
        statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) |>
    add_overall() |>
    add_p(test = list(all_continuous() ~ "oneway.test")) |>
    add_q() |>
    modify_caption("**OGTT vs. HbA1c Only**")
mod_df |>
    filter(ReviewYear == first_cfrd) |>
    filter(a1c_ogtt %in% c("OGTT", "Neither")) |>
    mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
    select(all_of(t1_vars)) |>
    tbl_summary(
        by = a1c_ogtt,
        missing_text = "Missing",
        statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) |>
    add_overall() |>
    add_p(test = list(all_continuous() ~ "oneway.test")) |>
    add_q() |>
    modify_caption("**OGTT vs. Neither**")
mod_df |>
    filter(ReviewYear == first_cfrd) |>
    filter(a1c_ogtt %in% c("HbA1c Only", "Neither")) |>
    mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
    select(all_of(t1_vars)) |>
    tbl_summary(
        by = a1c_ogtt,
        missing_text = "Missing",
        statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) |>
    add_overall() |>
    add_p(test = list(all_continuous() ~ "oneway.test")) |>
    add_q() |>
    modify_caption("**HbA1c Only vs. Neither**")
```

## FEV1

Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, and pancreatic enzyme use. A model with random intercept for participant and continuous AR(1) correlation structure was selected based on AIC and visual inspection of diagnostic plots.

```{r}
#| warning: false
#| message: false
# Fit the model
# m1 <- lme(
#     A_FEV1_pct_predicted ~ years_to_cfrd *
#         a1c_ogtt +
#         years_to_cfrd * treatment +
#         age_dx +
#         NumFEV1 +
#         Gender +
#         MutClass +
#         A_IsOnEnzymes,
#     random = ~ 1 | eDWID,
#     data = mod_df,
#     na.action = "na.omit"
# )
m2 <- lme(
    A_FEV1_pct_predicted ~ years_to_cfrd *
        a1c_ogtt +
        years_to_cfrd * treatment +
        age_dx +
        NumFEV1 +
        Gender +
        MutClass +
        A_IsOnEnzymes,
    random = ~ 1 | eDWID,
    correlation = corCAR1(),
    data = mod_df,
    na.action = "na.exclude"
)
# m3 <- lme(
#     A_FEV1_pct_predicted ~ years_to_cfrd *
#         a1c_ogtt +
#         years_to_cfrd * treatment +
#         age_dx +
#         NumFEV1 +
#         Gender +
#         MutClass +
#         A_IsOnEnzymes,
#     random = ~ 1 | eDWID,
#     correlation = corARMA(p = 2, q = 0),
#     data = mod_df,
#     na.action = "na.omit"
# )
t3 <- car::Anova(m2, type = "III", contrasts = "contr.sum")
emm <- emmeans(
    m2,
    specs = ~ years_to_cfrd + a1c_ogtt + treatment,
    pbkrtest.limit = 1e5
)
ggplot(
    data.frame(emm),
    aes(
        x = years_to_cfrd,
        y = emmean,
        group = interaction(a1c_ogtt, treatment),
        color = a1c_ogtt,
        shape = treatment
    )
) +
    geom_point() +
    geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
    geom_line(aes(linetype = treatment)) +
    xlab("Years From CFRD Dx") +
    ylab("Estimated Mean ppFEV1") +
    labs(
        color = "Diagnosis Method",
        shape = "CFRD Treatment Within 1 Year",
        linetype = "CFRD Treatment Within 1 Year",
        caption = "Results are averaged over the levels of gender, mutation class, and pancreatic enzye use."
    ) +
    theme_bw() +
    facet_wrap(~a1c_ogtt)
# Model table
m2 |> tbl_regression()
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[which(rownames(t3)=="years_to_cfrd:a1c_ogtt"),"Pr(>Chisq)"], eps = 0.001)`) and time and treatment (p =`r format.pval(t3[which(rownames(t3)=="years_to_cfrd:treatment"),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
#| warning: false
#| message: false
# Create contrasts
emm_t <- data.frame(emm)
base <- rep(0, times = nrow(emm_t))
# Neither, no treatment
neither_nt_m_5 <- base
neither_nt_m_5[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == -5
)] <- 1
neither_nt_m_1 <- base
neither_nt_m_1[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == -1
)] <- 1
neither_nt_0 <- base
neither_nt_0[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 0
)] <- 1
neither_nt_1 <- base
neither_nt_1[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 1
)] <- 1
neither_nt_5 <- base
neither_nt_5[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 5
)] <- 1
# HbA1c, no treatment
hba1c_nt_m_5 <- base
hba1c_nt_m_5[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == -5
)] <- 1
hba1c_nt_m_1 <- base
hba1c_nt_m_1[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == -1
)] <- 1
hba1c_nt_0 <- base
hba1c_nt_0[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 0
)] <- 1
hba1c_nt_1 <- base
hba1c_nt_1[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 1
)] <- 1
hba1c_nt_5 <- base
hba1c_nt_5[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 5
)] <- 1
# OGTT, no treatment
ogtt_nt_m_5 <- base
ogtt_nt_m_5[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == -5
)] <- 1
ogtt_nt_m_1 <- base
ogtt_nt_m_1[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == -1
)] <- 1
ogtt_nt_0 <- base
ogtt_nt_0[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 0
)] <- 1
ogtt_nt_1 <- base
ogtt_nt_1[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 1
)] <- 1
ogtt_nt_5 <- base
ogtt_nt_5[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "No Treatment" &
        emm_t$years_to_cfrd == 5
)] <- 1
# Neither, treatment
neither_t_m_5 <- base
neither_t_m_5[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == -5
)] <- 1
neither_t_m_1 <- base
neither_t_m_1[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == -1
)] <- 1
neither_t_0 <- base
neither_t_0[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 0
)] <- 1
neither_t_1 <- base
neither_t_1[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 1
)] <- 1
neither_t_5 <- base
neither_t_5[which(
    emm_t$a1c_ogtt == "Neither" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 5
)] <- 1
# HbA1c, treatment
hba1c_t_m_5 <- base
hba1c_t_m_5[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == -5
)] <- 1
hba1c_t_m_1 <- base
hba1c_t_m_1[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == -1
)] <- 1
hba1c_t_0 <- base
hba1c_t_0[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 0
)] <- 1
hba1c_t_1 <- base
hba1c_t_1[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 1
)] <- 1
hba1c_t_5 <- base
hba1c_t_5[which(
    emm_t$a1c_ogtt == "HbA1c Only" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 5
)] <- 1
# OGTT, treatment
ogtt_t_m_5 <- base
ogtt_t_m_5[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == -5
)] <- 1
ogtt_t_m_1 <- base
ogtt_t_m_1[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == -1
)] <- 1
ogtt_t_0 <- base
ogtt_t_0[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 0
)] <- 1
ogtt_t_1 <- base
ogtt_t_1[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 1
)] <- 1
ogtt_t_5 <- base
ogtt_t_5[which(
    emm_t$a1c_ogtt == "OGTT" &
        emm_t$treatment == "Treatment" &
        emm_t$years_to_cfrd == 5
)] <- 1
```

```{r}
#| warning: false
#| message: false
kable(data.frame(emm), digits = 3, caption = "Estimated Marginal Means (EMM)")
```

The table above indicates the average FEV1 at a given timepoint within each group. For example, people who were diagnosed by OGTT but not started on CFRD treatment within the first year had an estimated mean FEV1 of 76.368 at diagnosis. The "lower.CL" column indicates the lower bound of the 95% CI, and the "upper.CL" column indicates the upper bound.

```{r}
#| warning: false
#| message: false
# Test contrasts
F <- contrast(
    emm,
    list(
        "Neither: -5 vs 0 years (No Treatment)" = neither_nt_0 - neither_nt_m_5,
        "Neither: 0 vs 5 years (No Treatment)" = neither_nt_5 - neither_nt_0,
        "Neither: -1 vs 0 years (No Treatment)" = neither_nt_0 - neither_nt_m_1,
        "Neither: 0 vs 1 years (No Treatment)" = neither_nt_1 - neither_nt_0,
        "HbA1c Only: -5 vs 0 years (No Treatment)" = hba1c_nt_0 - hba1c_nt_m_5,
        "HbA1c Only: 0 vs 5 years (No Treatment)" = hba1c_nt_5 - hba1c_nt_0,
        "HbA1c Only: -1 vs 0 years (No Treatment)" = hba1c_nt_0 - hba1c_nt_m_1,
        "HbA1c Only: 0 vs 1 years (No Treatment)" = hba1c_nt_1 - hba1c_nt_0,
        "OGTT: -5 vs 0 years (No Treatment)" = ogtt_nt_0 - ogtt_nt_m_5,
        "OGTT: 0 vs 5 years (No Treatment)" = ogtt_nt_5 - ogtt_nt_0,
        "OGTT: -1 vs 0 years (No Treatment)" = ogtt_nt_0 - ogtt_nt_m_1,
        "OGTT: 0 vs 1 years (No Treatment)" = ogtt_nt_1 - ogtt_nt_0,
        "Neither: -5 vs 0 years (Treatment)" = neither_t_0 - neither_t_m_5,
        "Neither: 0 vs 5 years (Treatment)" = neither_t_5 - neither_t_0,
        "Neither: -1 vs 0 years (Treatment)" = neither_t_0 - neither_t_m_1,
        "Neither: 0 vs 1 years (Treatment)" = neither_t_1 - neither_t_0,
        "HbA1c Only: -5 vs 0 years (Treatment)" = hba1c_t_0 - hba1c_t_m_5,
        "HbA1c Only: 0 vs 5 years (Treatment)" = hba1c_t_5 - hba1c_t_0,
        "HbA1c Only: -1 vs 0 years (Treatment)" = hba1c_t_0 - hba1c_t_m_1,
        "HbA1c Only: 0 vs 1 years (Treatment)" = hba1c_t_1 - hba1c_t_0,
        "OGTT: -5 vs 0 years (Treatment)" = ogtt_t_0 - ogtt_t_m_5,
        "OGTT: 0 vs 5 years (Treatment)" = ogtt_t_5 - ogtt_t_0,
        "OGTT: -1 vs 0 years (Treatment)" = ogtt_t_0 - ogtt_t_m_1,
        "OGTT: 0 vs 1 years (Treatment)" = ogtt_t_1 - ogtt_t_0
    ),
    adjust = "fdr"
)
F <- data.frame(F)
F$lower.CL <- F$estimate - 1.96 * F$SE
F$upper.CL <- F$estimate + 1.96 * F$SE
kable(F, digits = 3, caption = "EMM Contrasts")
```

The table above indicates the change in average FEV1 from a given timepoint to another within each group. For example, FEV1 in people who were diagnosed by OGTT but not started on CFRD treatment within the first year decreased by -8.612 on average over the 5 years prior to CFRD diagnosis. The "lower.CL" column indicates the lower bound of the 95% CI for this change, and the "upper.CL" column indicates the upper bound.

## FVC

Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, and pancreatic enzyme use. We again used a model with random intercept for participant and continuous AR(1) correlation structure, to be consistent with the FEV1 model.

```{r}
#| warning: false
#| message: false
# Fit the model
m2 <- lme(
    A_FVC_pct_predicted ~ years_to_cfrd *
        a1c_ogtt +
        years_to_cfrd * treatment +
        age_dx +
        NumFVC +
        Gender +
        MutClass +
        A_IsOnEnzymes,
    random = ~ 1 | eDWID,
    correlation = corCAR1(),
    data = mod_df,
    na.action = "na.exclude"
)
t3 <- car::Anova(m2, type = "III", contrasts = "contr.sum")
emm <- emmeans(
    m2,
    specs = ~ years_to_cfrd + a1c_ogtt + treatment,
    pbkrtest.limit = 1e5
)
ggplot(
    data.frame(emm),
    aes(
        x = years_to_cfrd,
        y = emmean,
        group = interaction(a1c_ogtt, treatment),
        color = a1c_ogtt,
        shape = treatment
    )
) +
    geom_point() +
    geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
    geom_line(aes(linetype = treatment)) +
    xlab("Years From CFRD Dx") +
    ylab("Estimated Mean ppFVC") +
    labs(
        color = "Diagnosis Method",
        shape = "CFRD Treatment Within 1 Year",
        linetype = "CFRD Treatment Within 1 Year",
        caption = "Results are averaged over the levels of gender, mutation class, and pancreatic enzye use."
    ) +
    theme_bw() +
    facet_wrap(~a1c_ogtt)
# Model table
m2 |> tbl_regression()
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[which(rownames(t3)=="years_to_cfrd:a1c_ogtt"),"Pr(>Chisq)"], eps = 0.001)`) and time and treatment (p =`r format.pval(t3[which(rownames(t3)=="years_to_cfrd:treatment"),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
#| warning: false
#| message: false
kable(data.frame(emm), digits = 3, caption = "Estimated Marginal Means (EMM)")
```

The table above indicates the average FVC at a given timepoint within each group. For example, people who were diagnosed by OGTT but not started on CFRD treatment within the first year had an estimated mean FVC of 86.267 at diagnosis. The "lower.CL" column indicates the lower bound of the 95% CI, and the "upper.CL" column indicates the upper bound.

```{r}
#| warning: false
#| message: false
# Test contrasts
F <- contrast(
    emm,
    list(
        "Neither: -5 vs 0 years (No Treatment)" = neither_nt_0 - neither_nt_m_5,
        "Neither: 0 vs 5 years (No Treatment)" = neither_nt_5 - neither_nt_0,
        "Neither: -1 vs 0 years (No Treatment)" = neither_nt_0 - neither_nt_m_1,
        "Neither: 0 vs 1 years (No Treatment)" = neither_nt_1 - neither_nt_0,
        "HbA1c Only: -5 vs 0 years (No Treatment)" = hba1c_nt_0 - hba1c_nt_m_5,
        "HbA1c Only: 0 vs 5 years (No Treatment)" = hba1c_nt_5 - hba1c_nt_0,
        "HbA1c Only: -1 vs 0 years (No Treatment)" = hba1c_nt_0 - hba1c_nt_m_1,
        "HbA1c Only: 0 vs 1 years (No Treatment)" = hba1c_nt_1 - hba1c_nt_0,
        "OGTT: -5 vs 0 years (No Treatment)" = ogtt_nt_0 - ogtt_nt_m_5,
        "OGTT: 0 vs 5 years (No Treatment)" = ogtt_nt_5 - ogtt_nt_0,
        "OGTT: -1 vs 0 years (No Treatment)" = ogtt_nt_0 - ogtt_nt_m_1,
        "OGTT: 0 vs 1 years (No Treatment)" = ogtt_nt_1 - ogtt_nt_0,
        "Neither: -5 vs 0 years (Treatment)" = neither_t_0 - neither_t_m_5,
        "Neither: 0 vs 5 years (Treatment)" = neither_t_5 - neither_t_0,
        "Neither: -1 vs 0 years (Treatment)" = neither_t_0 - neither_t_m_1,
        "Neither: 0 vs 1 years (Treatment)" = neither_t_1 - neither_t_0,
        "HbA1c Only: -5 vs 0 years (Treatment)" = hba1c_t_0 - hba1c_t_m_5,
        "HbA1c Only: 0 vs 5 years (Treatment)" = hba1c_t_5 - hba1c_t_0,
        "HbA1c Only: -1 vs 0 years (Treatment)" = hba1c_t_0 - hba1c_t_m_1,
        "HbA1c Only: 0 vs 1 years (Treatment)" = hba1c_t_1 - hba1c_t_0,
        "OGTT: -5 vs 0 years (Treatment)" = ogtt_t_0 - ogtt_t_m_5,
        "OGTT: 0 vs 5 years (Treatment)" = ogtt_t_5 - ogtt_t_0,
        "OGTT: -1 vs 0 years (Treatment)" = ogtt_t_0 - ogtt_t_m_1,
        "OGTT: 0 vs 1 years (Treatment)" = ogtt_t_1 - ogtt_t_0
    ),
    adjust = "fdr"
)
F <- data.frame(F)
F$lower.CL <- F$estimate - 1.96 * F$SE
F$upper.CL <- F$estimate + 1.96 * F$SE
kable(F, digits = 3, caption = "EMM Contrasts")
```

The table above indicates the change in average FVC from a given timepoint to another within each group. For example, FVC in people who were diagnosed by OGTT but not started on CFRD treatment within the first year decreased by -5.752 on average over the 5 years prior to CFRD diagnosis. The "lower.CL" column indicates the lower bound of the 95% CI for this change, and the "upper.CL" column indicates the upper bound.