---
title: "CF Registry Analysis - Lung Function"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    embed-resources: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(ggplot2)
library(gtsummary)
library(gt)
library(pROC)
library(ggrepel)
library(childsds)
library(labelled)
library(Hmisc)
library(nlme)
library(emmeans)
library(multcomp)
library(tidyverse)
library(parallel)
library(knitr)
base_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/home/tvigers/Documents/Data",
  "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
  "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
  "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
  "Tims-Mac-Studio.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers"
)
data_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/CF/Christine Chan/CF Registry",
  "Tims-MacBook-Air.local" = "/CF/Christine Chan/CF Registry",
  "Tims-Mac-mini.local" = "/CF/Christine Chan/CF Registry",
  "Mac" = "/CF/Christine Chan/CF Registry",
  "Tims-Mac-Studio.local" = "/CF/Christine Chan/CF Registry"
)
home_dir <- paste0(base_dir, data_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
#| warning: false
#| message: false
# Import data and codebook
setwd(home_dir)
annual <- read.csv(
  "./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
# Remove those with T1D or T2D
annual <- annual |> arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(
  annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations, etc.
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
# Combine race columns
races <- c(
  "White",
  "Black or African American",
  "American Indian or Alaska Native",
  "Asian",
  "Native Hawaiian or Other Pacific ",
  "Some other race"
)
demogs$Race <- apply(demogs, 1, function(r) {
  w <- which(r[paste0("Race", 1:6)] == 1)
  if (length(w) > 1) {
    return("More than one race")
  } else {
    return(races[w])
  }
})
demogs$Race <- factor(
  demogs$Race,
  levels = c(
    "White",
    "American Indian or Alaska Native",
    "Asian",
    "Black or African American",
    "More than one race",
    "Native Hawaiian or Other Pacific ",
    "Some other race"
  ),
  labels = c(
    "White",
    "Other race",
    "Other race",
    "Black or African American",
    "More than one race",
    "Other race",
    "Other race"
  )
)
demogs$Hispanicrace <- factor(
  demogs$Hispanicrace,
  levels = 1:2,
  labels = c("Yes", "No")
)
demogs <- demogs |> select(eDWID, Gender, MutClass, Race, Hispanicrace)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Combine insurance columns
ins_vars <- c(paste0("insurance", 1:7), "noinsurance")
ins <- c(
  "Health Insurance Policy (e.g. Private Insurance)",
  "Medicare",
  "Medicaid",
  "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
  "TriCare or other military health plan",
  "Indian Health Service",
  "Other",
  "No Insurance"
)
annual$Insurance <- apply(annual[, ins_vars], 1, function(r) {
  w <- which(r == 1)
  if (length(w) == 0) {
    return(NA)
  } else if (length(w) > 1) {
    return("More than one")
  } else {
    return(ins[w])
  }
})
annual$Insurance <- factor(
  annual$Insurance,
  levels = c(
    "Health Insurance Policy (e.g. Private Insurance)",
    "Indian Health Service",
    "Medicaid",
    "Medicare",
    "More than one",
    "No Insurance",
    "Other",
    "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
    "TriCare or other military health plan"
  ),
  labels = c(
    "Private",
    "Public",
    "Public",
    "Public",
    "More Than One",
    "Other",
    "Other",
    "Public",
    "Other"
  )
)
# Family income
annual$family_income = factor(
  annual$family_income,
  levels = 1:10,
  labels = c(
    "<$10,000",
    "$10,000 to $39,999",
    "$10,000 to $39,999",
    "$10,000 to $39,999",
    "$40,000 to $89,999",
    "$40,000 to $89,999",
    "$40,000 to $89,999",
    "$40,000 to $89,999",
    "$40,000 to $89,999",
    ">$90,000"
  )
)
# Parent education
annual$max_parent_educ = pmax(
  annual$mother_education,
  annual$father_education,
  na.rm = TRUE
)
annual$max_parent_educ = factor(
  annual$max_parent_educ,
  levels = 1:6,
  labels = c(
    "Less than High School",
    "High School diploma or equivalent",
    "Some College",
    "College Graduate",
    "Masters/Doctoral level degree",
    "Unknown/Not applicable"
  )
)
annual$patient_education = factor(
  annual$patient_education,
  levels = 1:6,
  labels = c(
    "Less than High School",
    "High School diploma or equivalent",
    "Some College",
    "College Graduate",
    "Masters/Doctoral level degree",
    "Unknown/Not applicable"
  )
)
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Get BMI categories
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
annual$bmi_perc_cdc <- sds(
  annual$bmi,
  age = annual$Age_YrEnd,
  sex = annual$Gender,
  male = "M",
  female = "F",
  item = "bmi",
  ref = cdc.ref,
  type = "perc"
) *
  100
annual$bmi_perc_who <- sds(
  annual$bmi,
  age = annual$Age_YrEnd,
  sex = annual$Gender,
  male = "M",
  female = "F",
  item = "bmi",
  ref = who.ref,
  type = "perc"
) *
  100
annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
annual$bmi_cat <- apply(annual, 1, function(r) {
  age <- as.numeric(r["Age_YrEnd"])
  gender <- as.character(r["Gender"])
  bmi <- as.numeric(r["bmi"])
  bmi_perc <- as.numeric(r["bmi_perc"])
  if (age <= 20) {
    return(cut(
      bmi_perc,
      c(-Inf, 10, 50, 85, Inf),
      right = F,
      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
    ))
  } else if (age > 20) {
    if (gender == "M") {
      return(cut(
        bmi,
        c(-Inf, 18.5, 23, 25, Inf),
        right = F,
        labels = c(
          "Undernourished",
          "At Risk",
          "Target/Goal",
          "Overweight"
        )
      ))
    } else if (gender == "F") {
      return(cut(
        bmi,
        c(-Inf, 18.5, 22, 25, Inf),
        right = F,
        labels = c(
          "Undernourished",
          "At Risk",
          "Target/Goal",
          "Overweight"
        )
      ))
    }
  }
})
annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
# Raw values for adults, percentiles for peds
annual$bmi_adult <- annual$bmi
annual$bmi_adult[annual$Age_YrEnd <= 20] <- NA
annual$bmi_perc <- annual$bmi_perc
annual$bmi_perc[annual$Age_YrEnd > 20] <- NA
# Labels
var_label(annual$bmi) <- "BMI (all participants)"
var_label(annual$bmi_perc) <- "BMI percentile (age <= 20)"
var_label(annual$bmi_adult) <- "BMI value (age > 20)"
var_label(annual$bmi_cat) <- "BMI category"
# Format columns
annual$MutClass <- factor(
  annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(
  annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf),
  right = F
)
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa",
  "A_fungalyeast1",
  "A_pulmonarycomplications1",
  "A_mycocultureresults3",
  "A_cultureresults3"
)
# Yes/No variables
annual$cfrd_treatment_options1[annual$cfrd_treatment_options1 == 2] <- NA
annual$cfrd_treatment_options4[annual$cfrd_treatment_options4 == 4] <- NA
yn_vars <- c(
  "A_corticosteroids1",
  "A_feeding2",
  "A_Vx770",
  "A_VX809comb",
  "liver_disease",
  bugs,
  "cfrd_treatment_options2",
  "cfrd_treatment_options3",
  "cfrd_treatment_options4"
)
annual[, yn_vars] <- lapply(
  annual[, yn_vars],
  factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- factor(
  annual$A_IsOnEnzymes,
  levels = c(1, 0),
  labels = c("Yes", " No")
)
# Males to reference group
annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
# Hispanic no as reference
annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
# Get first year of CFRD
first_cfrd <- annual |>
  filter(cfrd_status_annual == "CFRD+") |>
  group_by(eDWID) |>
  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
# years following are CFRD+
annual$CFRD <- NA
annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
# Check whether they had an OGTT and/or HbA1c the year of first CFRD or the year
# prior. This is super slow but idc idc.
cl <- makeCluster(type = "FORK", 8)
cfrd_only <- parApply(
  cl,
  first_cfrd,
  1,
  function(r) {
    id <- as.numeric(r["eDWID"])
    year <- as.numeric(r["first_cfrd"])
    d <- annual[
      annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1),
    ]
    a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
    ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
    ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
    d <- annual[
      annual$eDWID == id & annual$ReviewYear %in% c(year, year + 1),
    ]
    treatment <- any(
      d[, c(
        "cfrd_treatment_options2",
        "cfrd_treatment_options3",
        "cfrd_treatment_options4"
      )] ==
        "Yes",
      na.rm = TRUE
    )
    return(c(id, year, a1c, ogtt0hr, ogtt2hr, treatment))
  }
)
stopCluster(cl)
cfrd_only <- data.frame(t(cfrd_only))
colnames(cfrd_only) <- c(
  "eDWID",
  "dx_year",
  "a1c",
  "ogtt0hr",
  "ogtt2hr",
  "treatment"
)
# cfrd_only_store <- cfrd_only
cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
  not_missing <- paste(
    which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
    collapse = "_"
  )
  return(not_missing)
})
cfrd_only$a1c_ogtt <- factor(
  cfrd_only$a1c_ogtt,
  levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
  labels = c(
    "Neither",
    "HbA1c Only",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT"
  )
)
var_label(cfrd_only$a1c_ogtt) <- "Diagnosis Method"
cfrd_only$treatment <- factor(
  cfrd_only$treatment,
  levels = 0:1,
  labels = c("No Treatment", "Treatment")
)
var_label(cfrd_only$treatment) <- "CFRD Treatment Started Within 1 Year?"
```

# Diagnosed with CFRD between 2010 - 2018

## Participant characteristics at diagnosis

```{r}
#| warning: false
#| message: false
# Create a model dataframe
mod_df <- inner_join(
  annual,
  cfrd_only |> filter(dx_year >= 2010),
  by = join_by(eDWID)
) |>
  drop_na(a1c_ogtt) |>
  mutate(years_to_cfrd = ReviewYear - first_cfrd) |>
  filter(years_to_cfrd <= 5 & years_to_cfrd >= -5) |>
  mutate(years_to_cfrd = factor(years_to_cfrd))
var_label(mod_df$years_to_cfrd) <- "Years From CFRD Dx"
# Write data for Christine to look at
mod_df |>
  select(
    eDWID,
    Race,
    Hispanicrace,
    MutClass,
    ReviewYear,
    Age_YrEnd,
    CFRD,
    first_cfrd,
    age_dx,
    A_weight,
    A_height,
    bmi_adult,
    bmi_perc,
    bmi_cat,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    a1c_ogtt,
    A_hgba1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    treatment,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    NumFEV1,
    NumFVC,
    Insurance,
    family_income,
    max_parent_educ,
    patient_education
  ) %>%
  write.csv(
    .,
    file = "./Data_Clean/fev1_model_data_2010.csv",
    row.names = FALSE,
    na = ""
  )
# Table 1 variable list
t1_vars <- c(
  "eDWID",
  "Race",
  "Hispanicrace",
  "MutClass",
  "age_dx",
  "A_weight",
  "A_height",
  "bmi_adult",
  "bmi_perc",
  "bmi_cat",
  "A_FEV1_pct_predicted",
  "A_FVC_pct_predicted",
  "a1c_ogtt",
  "A_hgba1c",
  "A_ogttfast_bloodglucose",
  "A_twohour_bloodglucose",
  "A_IsOnEnzymes",
  "liver_disease",
  "A_pseudomonasaeruginosa",
  "A_fungalyeast1",
  "A_mycocultureresults3",
  "A_cultureresults3",
  "treatment",
  "cfrd_treatment_options2",
  "cfrd_treatment_options3",
  "cfrd_treatment_options4",
  "NumFEV1",
  "NumFVC",
  "Insurance",
  "family_income",
  "max_parent_educ",
  "patient_education"
)
```

### All groups

```{r}
#| warning: false
#| message: false
mod_df |>
  filter(ReviewYear == first_cfrd) |>
  select(all_of(t1_vars)) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)) |>
  modify_caption("**All Groups**")
```

### Pairwise comparisons

P values FDR-adjusted for multiple comparisons.

```{r}
#| warning: false
#| message: false
mod_df |>
  filter(ReviewYear == first_cfrd) |>
  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
  select(all_of(t1_vars)) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(
    test = list(all_continuous() ~ "oneway.test"),
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) |>
  add_q() |>
  modify_caption("**OGTT vs. HbA1c Only**")
mod_df |>
  filter(ReviewYear == first_cfrd) |>
  filter(a1c_ogtt %in% c("OGTT", "Neither")) |>
  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
  select(all_of(t1_vars)) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(
    test = list(all_continuous() ~ "oneway.test"),
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) |>
  add_q() |>
  modify_caption("**OGTT vs. Neither**")
mod_df |>
  filter(ReviewYear == first_cfrd) |>
  filter(a1c_ogtt %in% c("HbA1c Only", "Neither")) |>
  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
  select(all_of(t1_vars)) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(
    test = list(all_continuous() ~ "oneway.test"),
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)
  ) |>
  add_q() |>
  modify_caption("**HbA1c Only vs. Neither**")
```

## FEV1

Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, pancreatic enzyme use, and insurance status. A model with random intercept for participant and continuous AR(1) correlation structure was selected based on AIC and visual inspection of diagnostic plots.

```{r}
#| warning: false
#| message: false
# Fit the model
m <- lme(
  A_FEV1_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFEV1 +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    Insurance,
  random = ~ 1 | eDWID,
  correlation = corAR1(
    form = ~ as.numeric(as.character(years_to_cfrd)) | eDWID
  ),
  data = mod_df,
  na.action = "na.exclude"
)
t3 <- car::Anova(m, type = "III", contrasts = "contr.sum")
emm <- emmeans(
  m,
  specs = ~ years_to_cfrd | a1c_ogtt,
  pbkrtest.limit = 1e5
)
ggplot(
  data.frame(emm),
  aes(
    x = years_to_cfrd,
    y = emmean,
    color = a1c_ogtt
  )
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line(aes(group = a1c_ogtt)) +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFEV1") +
  labs(
    color = "Diagnosis Method",
    caption = "Results are averaged over the levels of age at diagnosis, number of FEV1 measures, sex, mutation class, pancreatic enzyme use, and insurance status."
  ) +
  theme_bw() +
  facet_wrap(~a1c_ogtt)
# Type 3 test
t3
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[which(rownames(t3)=="years_to_cfrd:a1c_ogtt"),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
#| warning: false
#| message: false
summary(emm, infer = TRUE, adjust = "bonferroni")
```

The table above indicates the average FEV1 at a given timepoint within each group. For example, people who were diagnosed by OGTT had an estimated mean FEV1 of 73.7 at diagnosis. The "lower.CL" column indicates the lower bound of the 95% CI, and the "upper.CL" column indicates the upper bound. P values are FDR-adjusted for multiple comparisons.

```{r}
# Create contrasts
base = rep(0, length(emm@levels$years_to_cfrd))
minus5 = base
minus5[which(emm@levels$years_to_cfrd == "-5")] = 1
minus1 = base
minus1[which(emm@levels$years_to_cfrd == "-1")] = 1
zero = base
zero[which(emm@levels$years_to_cfrd == "0")] = 1
m5 = base
m5[which(emm@levels$years_to_cfrd == "5")] = 1
m1 = base
m1[which(emm@levels$years_to_cfrd == "1")] = 1
# Apply them
cont_res = contrast(
  emm,
  method = list(
    "-5 to 0 years" = zero - minus5,
    "-1 to 0 years" = zero - minus1,
    "0 to 1 year" = m1 - zero,
    "0 to 5 years" = m5 - zero
  )
)
summary(cont_res, infer = TRUE, adjust = "bonferroni")
```

The table above indicates the average change from one timepoint to another within each group. For example, in the "neither" group, FEV1 changed by -13.46 from 5 years prior to diagnosis. P values are FDR-adjusted for multiple comparisons.

```{r}
cont_res2 = contrast(
  cont_res,
  method = list(
    "-5 to 0 years vs. 0 to 5 years" = c(-1, 0, 0, 1),
    "-1 to 0 years vs. 0 to 1 years" = c(0, -1, 1, 0)
  )
)
summary(cont_res2, infer = TRUE, adjust = "bonferroni")
```

This table shows a "difference of differences" that compares the changes pre-CFRD to post-CFRD diagnosis. For example, in the "neither" group, the 5 year change prior to CFRD was -13.46 and the 5 year change after diagnosis was -8.71, indicating that while lung function still decreased, the difference was 4.75 points better after diagnosis. P values are FDR-adjusted for multiple comparisons.

## FVC

Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, pancreatic enzyme use, and insurance status. A model with random intercept for participant and continuous AR(1) correlation structure was selected based on AIC and visual inspection of diagnostic plots.

```{r}
#| warning: false
#| message: false
# Fit the model
m <- lme(
  A_FVC_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFVC +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    Insurance,
  random = ~ 1 | eDWID,
  correlation = corAR1(
    form = ~ as.numeric(as.character(years_to_cfrd)) | eDWID
  ),
  data = mod_df,
  na.action = "na.exclude"
)
t3 <- car::Anova(m, type = "III", contrasts = "contr.sum")
emm <- emmeans(
  m,
  specs = ~ years_to_cfrd | a1c_ogtt,
  pbkrtest.limit = 1e5
)
ggplot(
  data.frame(emm),
  aes(
    x = years_to_cfrd,
    y = emmean,
    color = a1c_ogtt
  )
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line(aes(group = a1c_ogtt)) +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFVC") +
  labs(
    color = "Diagnosis Method",
    caption = "Results are averaged over the levels of age at diagnosis, number of FVC measures, sex, mutation class, pancreatic enzyme use, and insurance status."
  ) +
  theme_bw() +
  facet_wrap(~a1c_ogtt)
# Type 3
t3
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[which(rownames(t3)=="years_to_cfrd:a1c_ogtt"),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
#| warning: false
#| message: false
summary(emm, infer = TRUE, adjust = "bonferroni")
```

The table above indicates the average FVC at a given timepoint within each group. For example, people who were diagnosed by OGTT had an estimated mean FVC of 83.9 at diagnosis. P values are FDR-adjusted for multiple comparisons.

```{r}
# Create contrasts
base = rep(0, length(emm@levels$years_to_cfrd))
minus5 = base
minus5[which(emm@levels$years_to_cfrd == "-5")] = 1
minus1 = base
minus1[which(emm@levels$years_to_cfrd == "-1")] = 1
zero = base
zero[which(emm@levels$years_to_cfrd == "0")] = 1
m5 = base
m5[which(emm@levels$years_to_cfrd == "5")] = 1
m1 = base
m1[which(emm@levels$years_to_cfrd == "1")] = 1
# Apply them
cont_res = contrast(
  emm,
  method = list(
    "-5 to 0 years" = zero - minus5,
    "-1 to 0 years" = zero - minus1,
    "0 to 1 year" = m1 - zero,
    "0 to 5 years" = m5 - zero
  )
)
summary(cont_res, infer = TRUE, adjust = "bonferroni")
```

The table above indicates the average change from a given timepoint to another within each group. For example, in the "neither" group, FVC changed by -10.59 from 5 years prior to diagnosis. P values are FDR-adjusted for multiple comparisons.

```{r}
cont_res2 = contrast(
  cont_res,
  method = list(
    "-5 to 0 years vs. 0 to 5 years" = c(-1, 0, 0, 1),
    "-1 to 0 years vs. 0 to 1 years" = c(0, -1, 1, 0)
  )
)
summary(cont_res2, infer = TRUE, adjust = "bonferroni")
```

This table shows a "difference of differences" that compares the changes pre-CFRD to post-CFRD diagnosis. For example, in the "neither" group, the 5 year change prior to CFRD was -10.59 and the 5 year change after diagnosis was -6.19, indicating that while lung function still decreased, the difference was 4.4 points better after diagnosis. P values are FDR-adjusted for multiple comparisons.