---
title: "The Changing Landscape of Cystic Fibrosis Related Diabetes Risk Factors Between 2003-2018"
author: "Tim Vigers"
date: "June 17, 2025"
date-format: long
format: 
  revealjs:
    embed-resources: true
editor: visual
bibliography: wnar_references.bib
csl: american-medical-association.csl
---

```{r setup}
#| include: false
library(tidyverse)
library(survival)
library(survminer)
library(gtsummary)
knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
```

# Background

## The *CFTR* Gene

-   Cystic fibrosis (CF) is an autosomal recessive disease caused by mutations in the cystic fibrosis transmembrane conductance regulator (*CFTR*) gene.[@grasemannCysticFibrosis2023]
-   The CFTR protein is an ion channel that regulates electrolyte balance across cell membranes.[@stoltzOriginsCysticFibrosis2015]
-   As of 2023, over 2000 CFTR variants have been described. Of these, we are certain that over 700 are disease-causing.[@grasemannCysticFibrosis2023]
-   The *CFTR* gene is expressed in the upper and lower airways, **pancreas**, intestine, biliary tree, cervix, sweat glands, and other organ systems.[@grasemannCysticFibrosis2023]

## Cystic Fibrosis-Related Diabetes

-   Because *CFTR* is highly expressed in the pancreas, *CFTR* dysfunction commonly leads to pancreatic insufficiency.[@moheetNewConceptsPathogenesis2022]
-   The pathophysiology of CFRD is not fully understood, but we think that $\beta$-cell loss related to pancreas destruction is one of the major drivers of diabetes in people with CF.[@moheetNewConceptsPathogenesis2022]
-   CFRD is the most common non-pulmonary complication in CF. A diagnosis before the age of 10 is unusual, but 50% of adults over 40 have CFRD.[@moheetNewConceptsPathogenesis2022]
-   CFRD is associated with a decline in nutritional status, lung function, and early mortality. [@moranClinicalCareGuidelines2010]

## Advancements in Treatment

- Median predicted survival increased from 33 to 44 years between 2003-2018.[@cysticfibrosisfoundationCysticFibrosisFoundation2018]
- We expect to see even greater gains with the approval of new highly effective *CFTR* modulator therapies. [@Bessonova731; @middletonElexacaftorTezacaftorIvacaftor2019]
- However, despite these advances CFRD remains prevalent and there have been few large-scale studies examining changes in the epidemiology of CFRD over time.[@moranClinicalCareGuidelines2010; @marshallEpidemiologyCysticFibrosisrelated2005; @adlerHyperglycemiaDeathCystic2011; @chamnanDiabetesDeterminantMortality2010]

# How Have CFRD Risk Factors Changed With the Introduction of Highly Effective *CFTR* Modulators?

## Known Risk Factors {.smaller}

- Earlier CFRD studies [@marshallEpidemiologyCysticFibrosisrelated2005; @moranCysticFibrosisRelated2009] found positive associations between CFRD risk and:
  - Age
  - Female sex
  - Race/ethnicity from marginalized groups
  - BMI
  - Genotype
  - Pancreatic insufficiency
  - Nutritional status
  - Liver disease
  - Worse pulmonary disease 
  - Use of corticosteroids
  
## Known Risk Factors (cont.)
- Other studies have identified additional potential risk factors, but these are the most well-established.
- These variables were also readily available in the United States Cystic Fibrosis Foundation Patient Registry (CFFPR) and have been collected consistently since 2003.
- Very few model selection issues (yay!)

# The Fun Part!

## Participant Demographics in 2018 {.smaller}

```{r}
#| message: false
load("./Data_Clean/survival_dataset.RData")
annual %>%
  filter(ReviewYear == 2018) %>%
  select(
    CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_cat
  ) %>%
  tbl_summary(
    by = CFRD, missing_text = "Missing",
    statistic = list(c("Age_YrEnd", "bmi_perc") ~ "{mean} ({sd})")
  ) %>%
  add_p(test = list(c("Age_YrEnd", "bmi_perc") ~ "t.test")) %>%
  add_overall() %>%
  separate_p_footnotes()
```

## Participant Demographics (cont.) {.smaller}

```{r}
annual %>%
  filter(ReviewYear == 2018) %>%
  select(
    CFRD, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
  ) %>%
  tbl_summary(
    by = CFRD, missing_text = "Missing",
    type = list(A_corticosteroids1 ~ "categorical"),
    statistic = list(c("A_FEV1_pct_predicted") ~ "{mean} ({sd})")
  ) %>%
  add_p(test = list(c("A_FEV1_pct_predicted") ~ "t.test")) %>%
  add_overall() %>%
  separate_p_footnotes()
```

## Kaplan-Meier Plot

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
# KM plot
fit <- survfit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
  data = annual
)
# Visualize with survminer
ggsurvplot(fit, data = annual, risk.table = TRUE, palette = "simpsons")
```

## Special Thanks To

- The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.  
- The patients, care providers, and clinic coordinators at CF centers.
- Kristen Miller, MS

## References
