---
title: "CHEC Study Sweat Chloride"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 9
    fig-height: 7
    page-layout: full
editor: source
execute:
  warning: false
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(readxl)
library(arsenal)
library(ggpubr)
library(lmerTest)
library(performance)
library(gtsummary)
library(gt)
library(patchwork)
library(ggeffects)
library(ggstatsplot)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
setwd(home_dir)
# Import CHEC data
chec <- read.csv(
  "./Data_Raw/CHEC sweat sodium and chloride 2024-04-19.csv",
  na.strings = c("", "NA")
)
ethnicity <- read.csv(
  "./Data_Raw/CHEC sweat sodium and chloride 2025-07-30.csv",
  na.strings = c("", "NA")
)
chec <- left_join(
  chec,
  ethnicity %>% select(subject, crfethnicity_txt),
  by = join_by(subject)
)
# Only include people with either a pre-modulator or post-modulator sodium
chec <- chec %>% filter(!is.na(swsavg), site != "005 Buffalo, NY CC and Peds")
# Factor variable formatting
chec$race_txt <- factor(
  chec$race_txt,
  levels = c(
    "White",
    "Black or African American",
    "American Indian or Alaska Native",
    "Native Hawaiian or Other Pacific Islander",
    "More than One Race",
    "Unknown or Not Reported"
  )
)
chec$crfethnicity_txt <- factor(
  chec$crfethnicity_txt,
  levels = c("Not Hispanic or Latino", "Hispanic or Latino")
)
chec$modqual <- factor(
  chec$modqual,
  levels = c(
    "Ivacaftor",
    "Lumacaftor / Ivacaftor",
    "Tezacaftor / Ivacaftor",
    "Elexacaftor / Tezacaftor / Ivacaftor"
  ),
  labels = c("IVA", "LUM/IVA", "TEZ/IVA", "ELX/TEZ/IVA")
)
chec$agecat_txt <- factor(
  chec$agecat_txt,
  levels = c(
    "2-5 years",
    "6-11 years",
    "12-17 years",
    "18-25 years",
    ">=26 years"
  )
)
# Average the values for people with more than 1 using CFF ID
chec = chec |>
  group_by(cffid_first) |>
  mutate(across(
    c(swhsavgpm, swhcavgpm, swsavg, swcavg),
    ~ mean(.x, na.rm = TRUE)
  )) |>
  filter(row_number() == 1)
# Calculate ratios
chec$pre_ratio <- chec$swhsavgpm / chec$swhcavgpm
chec$post_ratio <- chec$swsavg / chec$swcavg
# Long dataset for easier plotting and paired analysis
chec_long <- chec %>%
  rename(
    pre_cl = swhcavgpm,
    pre_na = swhsavgpm,
    post_cl = swcavg,
    post_na = swsavg
  ) %>%
  pivot_longer(
    c(pre_cl, pre_na, post_cl, post_na, pre_ratio, post_ratio),
    names_to = c("pre_post", ".value"),
    names_pattern = "(.*)_(.*)",
    values_to = "count"
  )
chec_long$pre_post <- factor(
  chec_long$pre_post,
  levels = c("pre", "post"),
  labels = c("Pre-Modulator", "Post-Modulator")
)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianrange"),
    cat.stats = c("N", "Nmiss", "countpct"),
    digits = 1
  )
# Labels
label(chec_long$cl) <- "Sweat Cl"
label(chec_long$na) <- "Sweat Na"
label(chec_long$ratio) <- "Sweat Na:Cl"
```

# Tables

## Table 1: Participant demographics by CFTR modulator

```{r results='asis'}
t1 <- tableby(
  modqual ~ sex +
    kwt(age) +
    agecat_txt +
    race_txt +
    crfethnicity_txt +
    genotype_group +
    swcavg +
    swhcavgpm +
    swsavg +
    swhsavgpm +
    pre_ratio +
    post_ratio,
  data = chec,
  cat.test = "fe",
  simulate.p.value = TRUE,
  control = mycontrols
)
summary(t1, pfootnote = T)
```

## Table 2: Change in sweat electrolytes

```{r results='asis'}
pt1 <- paired(
  pre_post ~ cl + na + signed.rank(ratio, "N", "Nmiss2", "medianq1q3", "range"),
  data = chec_long,

  id = subject,
  control = paired.control(signed.rank.exact = F)
)
summary(pt1, pfootnote = T)
```

## Table 3: Change in sweat electrolytes by modulator

```{r results='asis'}
pt2 <- paired(
  pre_post ~
    signed.rank(cl, "N", "Nmiss2", "medianq1q3", "range") +
    signed.rank(na, "N", "Nmiss2", "medianq1q3", "range") +
    signed.rank(ratio, "N", "Nmiss2", "medianq1q3", "range"),
  data = chec_long,
  id = subject,
  strata = modqual,
  control = paired.control(signed.rank.exact = F)
)
summary(pt2, pfootnote = T)
```

# Histograms and density plots

## Sweat chloride (overall)

### Histogram

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

### Density

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

## Sweat sodium (overall)

### Histogram

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

### Density

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

## Sweat sodium:chloride ratio (overall)

### Histogram

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

### Density

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790"))
```

## Sweat chloride by modulator

### Histogram

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

### Density

```{r}
ggplot(chec_long, aes(x = cl, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

## Sweat sodium by modulator

### Histogram

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

### Density

```{r}
ggplot(chec_long, aes(x = na, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

## Sweat sodium:chloride ratio by modulator

### Histogram

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_histogram(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

### Density

```{r}
ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
  facet_wrap(~modqual)
```

# Boxplots

```{r}
ggplot(chec_long, aes(x = pre_post, y = ratio)) +
  geom_boxplot() +
  geom_signif(
    comparison = list(c("Pre-Modulator", "Post-Modulator")),
    test = "wilcox.test",
    test.args = list(paired = TRUE, exact = FALSE),
    y_position = 3.5,
    map_signif_level = function(p) {
      format.pval(p, digits = 3, eps = 0.001)
    }
  ) +
  theme_bw() +
  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
  ylab("Na:Cl Ratio") +
  facet_wrap(~modqual)
```

# Updates for John

## Boxplots

```{r}
library(robustrank)
robust_wilcox <- function(a, b) {
  pm <- pm.wilcox.test(a, b)
  return(list(p.value = pm$p.value))
}
fancy_boxplot <- function(outcome, df = chec_long) {
  ns <- df |>
    group_by(pre_post) |>
    summarise(non_na_count = sum(!is.na(.data[[outcome]])))
  levels(df$pre_post) <- paste0(
    levels(df$pre_post),
    "\n(n=",
    ns$non_na_count,
    ")"
  )
  p <- ggplot(df, aes(x = pre_post, y = .data[[outcome]])) +
    geom_boxplot() +
    geom_signif(
      comparison = list(levels(df$pre_post)),
      test = "robust_wilcox",
      y_position = max(df[, outcome], na.rm = T),
      map_signif_level = function(p) {
        format.pval(p, digits = 3, eps = 0.001)
      }
    ) +
    theme_bw() +
    theme(legend.title = element_blank(), axis.title.x = element_blank()) +
    ylim(
      0,
      max(df[, outcome], na.rm = T) + (max(df[, outcome], na.rm = T) * 0.1)
    ) +
    facet_wrap(~modqual)
  return(p)
}
```

```{r}
fancy_boxplot("cl")
fancy_boxplot("na")
fancy_boxplot("ratio")
```

```{r}
#| code-fold: false
pm.wilcox.test(chec$swhcavgpm, chec$swcavg)
pm.wilcox.test(chec$swhsavgpm, chec$swsavg)
pm.wilcox.test(chec$pre_ratio, chec$post_ratio)
```

## Scatterplots

### Overall Pre

```{r}
plot_df <- chec_long |> filter(pre_post == "Pre-Modulator")
ggscatterstats(
  plot_df,
  x = na,
  y = cl,
  type = "nonparametric",
  ylab = "Sweat Cl",
  xlab = "Sweat Na",
  digits = "signif2",
  marginal = FALSE
)
ggscatterstats(
  plot_df,
  x = ratio,
  y = cl,
  type = "nonparametric",
  ylab = "Sweat Cl",
  xlab = "Sweat Na:Cl",
  digits = "signif2",
  marginal = FALSE
)
```

### Overall Post

```{r}
plot_df <- chec_long |> filter(pre_post == "Post-Modulator")
ggscatterstats(
  plot_df,
  x = na,
  y = cl,
  type = "nonparametric",
  ylab = "Sweat Cl",
  xlab = "Sweat Na",
  digits = "signif2",
  marginal = FALSE
)
ggscatterstats(
  plot_df,
  x = ratio,
  y = cl,
  type = "nonparametric",
  ylab = "Sweat Cl",
  xlab = "Sweat Na:Cl",
  digits = "signif2",
  marginal = FALSE
)
```

### Overall Post by Modulator

```{r}
grouped_ggscatterstats(
  plot_df,
  x = na,
  y = cl,
  grouping.var = modqual,
  type = "nonparametric",
  ylab = "Sweat Cl",
  xlab = "Sweat Na",
  digits = "signif2",
  marginal = FALSE
)
grouped_ggscatterstats(
  plot_df,
  x = ratio,
  y = cl,
  grouping.var = modqual,
  type = "nonparametric",
  ylab = "Sweat Cl",
  xlab = "Sweat Na:Cl",
  digits = "signif2",
  marginal = FALSE
)
```

## Mixed model

```{r}
#| code-fold: false
mod <- lmer(cl ~ pre_post * na + (1 | subject), data = chec_long)
plot(predict_response(mod, terms = c("na", "pre_post")))
mod |>
  tbl_regression() |>
  add_n(location = "level")
```

## Post Modulator Table

```{r}
chec_long |>
  filter(pre_post == "Post-Modulator") |>
  group_by(subject, modqual) |>
  summarise(
    "SCC < 60" = cl < 60,
    "Na:Cl > 1.2" = ratio > 1.2,
    "SCC < 60 AND Na:Cl > 1.2" = cl < 60 & ratio > 1.2,
    "SCC < 60 OR Na:Cl > 1.2" = cl < 60 | ratio > 1.2,
    .groups = "drop"
  ) |>
  tbl_summary(by = modqual, include = -subject, missing_text = "Missing") |>
  add_overall() |>
  add_p(test = "SCC < 60 AND Na:Cl > 1.2" ~ "fisher.test") |>
  separate_p_footnotes() |>
  remove_footnote_header()
```