---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import CHCO, filter out those with no demo data, rename columns, etc.
chco <- read.csv("./Data_Cleaned/chco_reidentified.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing", "Please reorder testing and re",
    "The received specimen cannot"
  )
)
chco <- chco %>%
  group_by(PatientID) %>%
  filter(Demographics.Available == "TRUE") %>%
  select(
    PatientID, Name, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
    Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
  ) %>%
  rename(Age = Age.at.Test.in.Years) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
# Fix CHCO diagnoses per Angela
chco_diagnoses <- read.csv("./Data_Cleaned/chco_diagnoses.csv")
chco <- full_join(chco, chco_diagnoses, by = join_by(PatientID))
chco$PatientID <- NULL
# For "LESS THAN 10", replace with half the limit of detection
chco[chco == "LESS THAN 10"] <- "9"
chco[, c("Cl", "Cl_Left", "Cl_Right")] <-
  lapply(chco[, c("Cl", "Cl_Left", "Cl_Right")], as.numeric)
# Import NJH and add diagnoses from Katie
njh <- read.csv("./Data_Cleaned/njh_full_list.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing"
  )
)
njh_diagnoses <- read.csv("./Data_Cleaned/njh_diagnoses.csv", na.strings = "")
njh_diagnoses <- njh_diagnoses %>% filter(!is.na(Diagnosis))
njh <- full_join(njh, njh_diagnoses)
# Filter out those with no demo data, rename columns, etc.
njh <- njh %>%
  group_by(Patient.ID) %>%
  filter(Demographics.Available == "TRUE") %>%
  select(
    Patient.ID, Patient.Name, Genotype.1, Genotype.2, Sex, Age.at.test..in.years.,
    Sweat.sodium, Sweat.potassium, Sweat.chloride, Sweat.sodium.left.arm,
    Sweat.sodium.right.arm, Sweat.potassium.left.arm, Sweat.potassium.right.arm,
    Sweat.chloride.left.arm, Sweat.chloride.right.arm, Diagnosis
  ) %>%
  rename(
    Name = Patient.Name, Genotypes1 = Genotype.1,
    Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
    K = Sweat.potassium, Cl = Sweat.chloride, Na_Left = Sweat.sodium.left.arm,
    Na_Right = Sweat.sodium.right.arm, K_Left = Sweat.potassium.left.arm,
    K_Right = Sweat.potassium.right.arm, Cl_Left = Sweat.chloride.left.arm,
    Cl_Right = Sweat.chloride.right.arm
  ) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
njh$Patient.ID <- NULL
# Row bind, fill diagnosis, then remove dupes
df <- rbind(chco, njh)
df <- df %>%
  group_by(Name) %>%
  fill(Diagnosis) %>%
  distinct() %>%
  select(Name, Diagnosis, everything()) %>%
  arrange(Name, Age)
# Use each person's first sweat test
df <- df %>%
  group_by(Name) %>%
  slice_min(Age)
# If there are still duplicates, exclude them
dups <- df$Name[duplicated(df$Name)]
df <- df[!df$Name %in% dups, ]
# For those with left and right, average them
df$Na_mean <- rowMeans(df[, c("Na_Left", "Na_Right")])
df$Na <- coalesce(df$Na, df$Na_mean)
df$Cl_mean <- rowMeans(df[, c("Cl_Left", "Cl_Right")])
df$Cl <- coalesce(df$Cl, df$Cl_mean)
df$K_mean <- rowMeans(df[, c("K_Left", "K_Right")])
df$K <- coalesce(df$K, df$K_mean)
# Get rid of unnecessary columns, remove those with no diagnosis
df <- df %>%
  select(Name, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl) %>%
  filter(!is.na(Diagnosis))
# Combine CFTR-related disorder and CRMS
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("Cystic Fibrosis", "CFTR-related disorder", "CRMS", "Non-CF"),
  labels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "CFTR-RD or CRMS", "Non-CF")
)
# Calculate Na:Cl ratio
df$`Na:Cl` <- df$Na / df$Cl
# Set table options
mycontrols <- tableby.control(
  numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range")
)
# Save this for Angela et al. to check
write.csv(df,
  file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
  row.names = F, na = ""
)
```

# Data cleaning steps

- Imported re-identified CHCO data from Elin (5,046 rows and 4,748 unique IDs)
- Removed rows with `Demographics Available` not equal to `TRUE` and selected each person's earliest test by age.
  - This resulted in 3,123 rows with 3,084 unique IDs.
- Merged in the diagnoses from Angela.
- Values below the limit of detection were replaced with half the limit (e.g. "LESS THAN 10" was replaced with 9). Approximately 5% of the values were below the L.O.D.

- Imported NJH data (1,958 rows with 1,748 unique IDs) and merged in diagnoses from Katie.
- Removed 1,662 rows with `Demographics Available` not equal to `TRUE` (all of these were `FALSE`)
  - This resulted in 172 rows with unique IDs.
- Concatenated the CHCO and NJH data and removed duplicate rows, resulting in 3,178 rows with 3,162 unique IDs.
- For the remaining people with multiple sweat tests (i.e. tests from both CHCO and NJH), kept only the earliest test (based on age at test). This resulted in 3,169 rows with 3,162 unique IDs.
- The remaining 7 people with different values for the same test in the CHCO and NJH data were excluded. The final analysis dataset includes 3,155 unique participants, 3,111 of which have a diagnosis.
- For people with both left and right arm values, the two were averaged.

# Table 1: Participant Demographics by Diagnosis

```{r results='asis'}
summary(tableby(Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

# Boxplots

## Chloride

```{r}
#| warning: false
clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

## Sodium

```{r}
#| warning: false
nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

## Na:Cl

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

# Histograms and Density Plots

## Chloride

```{r}
#| warning: false
ggplot(df, aes(x = Cl, fill = Diagnosis)) +
  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = Cl, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

## Sodium

```{r}
#| warning: false
ggplot(df, aes(x = Na, fill = Diagnosis)) +
  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = Na, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

## Na:Cl

```{r}
#| warning: false
ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  xlim(0, 2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

# Na:Cl ROC Curves

```{r}
# Fit a multiclass ROC
mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
roc_list <- list(
  "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
)
ggroc(roc_list, legacy.axes = T) +
  theme_bw() +
  theme(legend.title = element_blank())
roc_table <- lapply(roc_list, function(r) {
  coords(r, "best",
    best.method = "youden",
    ret = c(
      "threshold", "sensitivity", "specificity", "accuracy", "tpr", "fpr",
      "tnr", "fnr"
    )
  )
})
roc_table <- do.call(rbind, roc_table)
kable(roc_table, digits = 3)
```

- The multiclass AUC for Na:Cl is `r round(as.numeric(auc(mc_roc)),3)`. 
- The AUC for using Na:Cl to distinguish CF from non-CF was `r round(as.numeric(auc(mc_roc$rocs[[2]])),3)`
- The best threshold for distinguishing CF from non-CF (based on the Youden index) was `r round(roc_table["Cystic Fibrosis vs. Non-CF","threshold"],3)`, and provided a sensitivity of `r round(roc_table["Cystic Fibrosis vs. Non-CF","sensitivity"],3)` and specificity of `r round(roc_table["Cystic Fibrosis vs. Non-CF","specificity"],3)`

- The AUC for using Na:Cl to distinguish CF from CFTR-RD or CRMS was `r round(as.numeric(auc(mc_roc$rocs[[1]])),3)`
- The best threshold for distinguishing CF from CFTR-RD or CRMS (based on the Youden index) was `r round(roc_table["CFTR-RD or CRMS vs. Cystic Fibrosis","threshold"],3)`, and provided a sensitivity of `r round(roc_table["CFTR-RD or CRMS vs. Cystic Fibrosis","sensitivity"],3)` and specificity of `r round(roc_table["CFTR-RD or CRMS vs. Cystic Fibrosis","specificity"],3)`

- The AUC for using Na:Cl to distinguish non-CF from CFTR-RD or CRMS was `r round(as.numeric(auc(mc_roc$rocs[[3]])),3)`
- The best threshold for distinguishing CF from CFTR-RD or CRMS (based on the Youden index) was `r round(roc_table["CFTR-RD or CRMS vs. Non-CF","threshold"],3)`, and provided a sensitivity of `r round(roc_table["CFTR-RD or CRMS vs. Non-CF","sensitivity"],3)` and specificity of `r round(roc_table["CFTR-RD or CRMS vs. Non-CF","specificity"],3)`
