---
title: "NTM CT Scoring - Initial Agreement"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r}
#| include: false
library(tidyverse)
library(irr)
library(redcapAPI)
library(Hmisc)
library(DT)
library(RColorBrewer)
library(knitr)
library(pheatmap)
library(arsenal)
# Import data
api <- read.table("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
                  header = F
)
rcon <- redcapConnection(
  url = "https://redcap.ucdenver.edu/api/",
  token = api[1, 1]
)
data <- exportRecords(rcon, labels = F)
data <- data.frame(lapply(data, as.character))
data$scan_id <- as.numeric(data$scan_id)
# Make scorer IDs by initial
data$scorer <- paste0(
  toupper(substr(data$reader_first_name, 1, 1)),
  toupper(substr(data$reader_last_name, 1, 1))
)
data$scorer[data$scorer == "GG"] <- "GS"
# Deidentify readers
old <- c("AF", "DL", "JW")
new <- c("A", "B", "C")
data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
# Revert duplicate IDs back to original to match with gold standard
training <- read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
dups <- read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
# Change duplicates to OG IDs
old <- dups$duplicated_ssid..dates.shifted..17.days.
new <- dups$original_ssid
data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
# Change training IDs to OG IDs
old <- training$Training_id
new <- training$Cffid.Scoring.Set
data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
# Plot colors
name <- "Accent"
colors <- brewer.pal(6, name)
names(colors) <- c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose", NA)
custom_colors <- scale_fill_manual(name = "", values = colors)
```

| Cohen's Kappa | Interpretation         |
|---------------|------------------------|
| 0             | No agreement           |
| 0.10 - 0.20   | Slight agreement       |
| 0.21 - 0.40   | Fair agreement         |
| 0.41 - 0.60   | Moderate agreement     |
| 0.61 - 0.80   | Substantial agreement  |
| 0.81 - 0.99   | Near perfect agreement |
| 1             | Perfect agreement      |

# Comparison to gold standards

```{r}
# Gold standard data
gs <- data %>%
  arrange(scan_id, scorer) %>%
  filter(scan_id %in% training$Cffid.Scoring.Set) %>%
  group_by(scorer, scan_id) %>%
  slice_max(scoresheet_timestamp) %>%
  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
  ungroup() %>%
  pivot_longer(bronchiectasis_rul:atelectasis___6,
               names_to = c("feature", "location"),
               names_pattern = "(.*)_(.*)$"
  ) %>%
  mutate(
    feature = gsub("__", "", feature),
    feature = str_to_title(gsub("_", " ", feature))
  )
gs$location <- factor(gs$location,
                      levels = c(
                        "1", "2", "3", "4", "5", "6", "ling", "lll",
                        "lul", "rll", "rml", "rul"
                      ),
                      labels = c(
                        "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
                        "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
                      )
)
# One column per scorer
gs <- gs %>%
  pivot_wider(names_from = scorer, values_from = value) %>%
  arrange(feature, location)
```

## Bronciechtasis

```{r}
gs_bronch <- gs %>% filter(feature == "Bronchiectasis")
fk <- kappam.fleiss(gs_bronch[, c("A", "B", "C", "GS")])
A <- kappa2(gs_bronch[, c("A", "GS")])
B <- kappa2(gs_bronch[, c("B", "GS")])
C <- kappa2(gs_bronch[, c("C", "GS")])
# Plot
gs_bronch %>%
  select(scan_id, feature, location, A:GS) %>%
  pivot_longer(A:GS) %>%
  ggplot(., aes(x = name, y = location, fill = value)) +
  geom_tile(color = "black") +
  theme_bw() +
  labs(fill = "", x = "Scorer", y = "Location") +
  custom_colors +
  facet_wrap(~scan_id)
```

Overall Fleiss' Kappa for all 5 gold standard scans was `r round(fk$value,3)` (p = `r format.pval(fk$p.value,eps = 0.001,digits = 3)`).

Overall Cohen's Kappa for A vs. the gold standard was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).

Overall Cohen's Kappa for B vs. the gold standard was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).

Overall Cohen's Kappa for C vs. the gold standard was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).

## Other features

```{r}
gs_other <- gs %>% filter(feature != "Bronchiectasis")
# Check agreement with GS column
gs_other$A <- factor(gs_other$A == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
gs_other$B <- factor(gs_other$B == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
gs_other$C <- factor(gs_other$C == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
# Plot
gs_other %>%
  select(scan_id, feature, location, A:GS) %>%
  pivot_longer(A:C) %>%
  ggplot(., aes(x = name, y = location, fill = value)) +
  geom_tile(color = "black") +
  theme_bw() +
  labs(fill = "", x = "Scorer", y = "Location") +
  facet_grid(rows = vars(feature), cols = vars(scan_id))
# Tables
gs_other %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with GS")
gs_other %>%
  group_by(scan_id) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with GS")
gs_other %>%
  group_by(feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with GS")
gs_other %>%
  group_by(scan_id, feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with GS")
# Heatmaps
gs_other %>%
  group_by(scan_id) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  column_to_rownames("scan_id") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
gs_other %>%
  group_by(feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  column_to_rownames("feature") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
gs_other %>%
  group_by(scan_id, feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
  column_to_rownames("ids") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
```

# Duplicates

```{r}
# Gold standard data
dup <- data %>%
  arrange(scan_id, scorer) %>%
  filter(scan_id %in% dups$original_ssid) %>%
  group_by(scorer, scan_id) %>%
  mutate(n = row_number()) %>%
  select(scorer, scan_id, n, bronchiectasis_rul:atelectasis___6) %>%
  ungroup() %>%
  pivot_longer(bronchiectasis_rul:atelectasis___6,
               names_to = c("feature", "location"),
               names_pattern = "(.*)_(.*)$"
  ) %>%
  mutate(
    feature = gsub("__", "", feature),
    feature = str_to_title(gsub("_", " ", feature))
  )
dup$location <- factor(dup$location,
                       levels = c(
                         "1", "2", "3", "4", "5", "6", "ling", "lll",
                         "lul", "rll", "rml", "rul"
                       ),
                       labels = c(
                         "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
                         "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
                       )
)
# One column per scorer
dup <- dup %>%
  pivot_wider(names_from = scorer, values_from = value) %>%
  pivot_wider(names_from = n, values_from = A:C) %>%
  arrange(feature, location)
```

## Bronciechtasis

```{r}
dup_bronch <- dup %>% filter(feature == "Bronchiectasis")
A <- kappa2(dup_bronch[, c("A_1", "A_2")])
B <- kappa2(dup_bronch[, c("B_1", "B_2")])
C <- kappa2(dup_bronch[, c("C_1", "C_2")])
# Plot
dup_bronch %>%
  select(scan_id, feature, location, A_1:C_2) %>%
  pivot_longer(A_1:C_2) %>%
  ggplot(., aes(x = name, y = location, fill = value)) +
  geom_tile(color = "black") +
  theme_bw() +
  labs(fill = "", x = "Scorer", y = "Location") +
  custom_colors +
  facet_wrap(~scan_id)
```

Overall Cohen's Kappa for A vs. themselves was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).

Overall Cohen's Kappa for B vs. themselves was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).

Overall Cohen's Kappa for C vs. themselves was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).

## Other features

```{r}
dup_other <- dup %>% filter(feature != "Bronchiectasis")
# Check agreement with GS column
dup_other$A <- factor(dup_other$A_1 == dup_other$A_2, levels = c(F, T), labels = c("Disagree", "Agree"))
dup_other$B <- factor(dup_other$B_1 == dup_other$B_2, levels = c(F, T), labels = c("Disagree", "Agree"))
dup_other$C <- factor(dup_other$C_1 == dup_other$C_2, levels = c(F, T), labels = c("Disagree", "Agree"))
# Plot
dup_other %>%
  select(scan_id, feature, location, A:C) %>%
  pivot_longer(A:C) %>%
  ggplot(., aes(x = name, y = location, fill = value)) +
  geom_tile(color = "black") +
  theme_bw() +
  labs(fill = "", x = "Scorer", y = "Location") +
  facet_grid(rows = vars(feature), cols = vars(scan_id))
# Tables
dup_other %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
dup_other %>%
  group_by(scan_id) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
dup_other %>%
  group_by(feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
dup_other %>%
  group_by(scan_id, feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
# Heatmaps
dup_other %>%
  group_by(scan_id) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  column_to_rownames("scan_id") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
dup_other %>%
  group_by(feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  column_to_rownames("feature") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
dup_other %>%
  group_by(scan_id, feature) %>%
  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
  column_to_rownames("ids") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
```

# Singleton scans

```{r}
# Gold standard data
df <- data %>%
  arrange(scan_id, scorer) %>%
  filter(!(scan_id %in% dups$original_ssid) &
           !(scan_id %in% training$Cffid.Scoring.Set)) %>%
  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
  pivot_longer(bronchiectasis_rul:atelectasis___6,
               names_to = c("feature", "location"),
               names_pattern = "(.*)_(.*)$"
  ) %>%
  mutate(
    feature = gsub("__", "", feature),
    feature = str_to_title(gsub("_", " ", feature))
  )
df$location <- factor(df$location,
                      levels = c(
                        "1", "2", "3", "4", "5", "6", "ling", "lll",
                        "lul", "rll", "rml", "rul"
                      ),
                      labels = c(
                        "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
                        "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
                      )
)
# One column per scorer
df <- df %>%
  pivot_wider(names_from = scorer, values_from = value) %>%
  arrange(feature, location)
```

## Bronciechtasis

```{r}
df_bronch <- df %>% filter(feature == "Bronchiectasis")
fk <- kappam.fleiss(df_bronch[, c("A", "B", "C")])
# Plot
df_bronch %>%
  select(scan_id, feature, location, A:C) %>%
  pivot_longer(A:C) %>%
  ggplot(., aes(x = name, y = location, fill = value)) +
  geom_tile(color = "black") +
  theme_bw() +
  labs(fill = "", x = "Scorer", y = "Location") +
  custom_colors +
  facet_wrap(~scan_id)
```

Overall Fleiss' Kappa for all 15 singleton scans was `r round(fk$value,3)` (p = `r format.pval(fk$p.value,eps = 0.001,digits = 3)`).

## Other features

```{r}
df_other <- df %>% filter(feature != "Bronchiectasis")
# Check agreement with GS column
df_other[, c("A", "B", "C")] <- lapply(df_other[, c("A", "B", "C")], factor,
                                       levels = c("Unchecked", "Checked"),
                                       labels = c("Absent", "Present")
)
# Three-way agreement
df_other$`A vs. B` <- factor(df_other$A == df_other$B, levels = c(F, T), labels = c("Disagree", "Agree"))
df_other$`A vs. C` <- factor(df_other$A == df_other$C, levels = c(F, T), labels = c("Disagree", "Agree"))
df_other$`B vs. C` <- factor(df_other$B == df_other$C, levels = c(F, T), labels = c("Disagree", "Agree"))
# Plot
df_other %>%
  select(scan_id, feature, location, `A vs. B`:`B vs. C`) %>%
  pivot_longer(`A vs. B`:`B vs. C`) %>%
  ggplot(., aes(x = name, y = location, fill = value)) +
  geom_tile(color = "black") +
  theme_bw() +
  labs(fill = "", x = "Scorer", y = "Location") +
  facet_grid(rows = vars(feature), cols = vars(scan_id))
# Tables
df_other %>%
  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
df_other %>%
  group_by(scan_id) %>%
  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
df_other %>%
  group_by(feature) %>%
  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
df_other %>%
  group_by(scan_id, feature) %>%
  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
  kable(., digits = 3, caption = "Proportion of agreement with themselves")
# Heatmaps
df_other %>%
  group_by(scan_id) %>%
  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  column_to_rownames("scan_id") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
df_other %>%
  group_by(feature) %>%
  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  column_to_rownames("feature") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
df_other %>%
  group_by(scan_id, feature) %>%
  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
  column_to_rownames("ids") %>%
  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
```

# Comparison of scores between groups

For scans that were duplicated, we looked at the first reading and excluded the second. Scores were calculated by considering bronchiectasis as absent = 0, cylindric = 1, varicose = 2, cystic = 3. For all other features absence was counted as 0 and presence as 1. 

```{r}
# Import disease state
outcomes = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
outcomes$NTM.PD..Y.N. = factor(outcomes$NTM.PD..Y.N.,
                               levels = c("No","Yes"),ordered = T)
# Remove training scans, get first reading of duplicates, convert to numeric
scores = data %>% 
  filter(scan_id %in% outcomes$cff_registry_id,scorer != "GS") %>%
  group_by(scorer,scan_id) %>%
  slice_min(scoresheet_timestamp) %>% 
  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
  mutate(across(contains("bronchiectasis"),
                ~as.numeric(factor(.x,levels=c("Absent","Cylindric",
                                               "Varicose","Cystic")))-1),
         across(mucoid_impaction___1:atelectasis___6,
                ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1))
# Add outcome info
scores = left_join(scores,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
                   by = join_by(scan_id == cff_registry_id)) %>%
  rename(`NTM-PD` = NTM.PD..Y.N.)
# Calculate scores
majority = scores %>%
  group_by(scan_id) %>%
  summarise(across(bronchiectasis_rul:`NTM-PD`,
                   ~tail(names(sort(table(.x))), 1)))
max = scores %>%
  group_by(scan_id) %>%
  summarise(across(bronchiectasis_rul:`NTM-PD`,
                   ~max(.x,na.rm = T)))
```

```{r}
comp_score = function(df){
  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
                           names_to = c("feature", "location"),
                           names_pattern = "(.*)_(.*)$"
  ) %>%
    mutate(
      feature = gsub("__", "", feature),
      feature = str_to_title(gsub("_", " ", feature))
    )
  df$location <- factor(df$location,
                        levels = c(
                          "1", "2", "3", "4", "5", "6", "ling", "lll",
                          "lul", "rll", "rml", "rul"
                        ),
                        labels = c(
                          "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
                          "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
                        )
  )
  df$value = as.numeric(df$value)
  # Tables
  # Total score
  sums = df %>% group_by(scan_id,`NTM-PD`) %>%
    summarise(Total = sum(value),.groups = "drop") 
  t1 = tableby(`NTM-PD`~Total,data = sums,numeric.test = "kwt",
               numeric.stats = c("Nmiss2","median","q1q3","range"))
  # By feature
  sums = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
    summarise(score = sum(value),.groups = "drop") %>%
    pivot_wider(names_from = feature,values_from = score)
  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$feature),"`"),
                                           collapse = "+")))
  t2 = tableby(f,data = sums,numeric.test = "kwt",
               numeric.stats = c("Nmiss2","median","q1q3","range"))
  # By location
  sums = df %>% group_by(scan_id,`NTM-PD`,location) %>%
    summarise(score = sum(value),.groups = "drop") %>%
    pivot_wider(names_from = location,values_from = score)
  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$location),"`"),
                                           collapse = "+")))
  t3 = tableby(f,data = sums,numeric.test = "kwt",
               numeric.stats = c("Nmiss2","median","q1q3","range"))
  # Plots
  total_plot = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
    summarise(score = sum(value),.groups = "drop") %>%
    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
    geom_violin() +
    theme_bw() + 
    theme(legend.position = "none")
  feature_plot = total_plot +
    facet_wrap(~feature,scales = "free") 
  location_plot = df %>% group_by(scan_id,`NTM-PD`,location) %>%
    summarise(score = sum(value),.groups = "drop") %>%
    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
    geom_violin() +
    theme_bw() + 
    theme(legend.position = "none") +
    facet_wrap(~location,scales = "free") 
  # Print
  cat("\n")
  cat("### Total Scores")
  cat("\n")
  cat("\n")
  print(summary(t1,pfootnote = T))
  cat("\n")
  cat("\n")
  print(total_plot)
  cat("\n")
  cat("\n")
  cat("### Scores by Feature")
  cat("\n")
  cat("\n")
  print(summary(t2,pfootnote = T))
  cat("\n")
  cat("\n")
  print(feature_plot)
  cat("\n")
  cat("\n")
  cat("### Scores by Location")
  cat("\n")
  cat("\n")
  print(summary(t3,pfootnote = T))
  cat("\n")
  cat("\n")
  print(location_plot)
  cat("\n")
  cat("\n")
}
```

## Using majority score for each feature

```{r results='asis'}
comp_score(majority)
```

## Using maximum score for each feature

```{r results='asis'}
comp_score(max)
```
